### 2022-05-05 talk/talk.qmd
start working on talk

---
title: "Developing Measures of Public Perceptions of Values in Science"
author: |
    | Dan Hicks 
    | Emilio Lobato
    | University of California, Merced
    | *[bitly]*
toc: true
format: 
    revealjs:
        toc-depth: 1
        slide-number: true
        center: false
        <!--margin: 0.0-->
link-citations: true
---

# Controversy and mismatch # {center="true"}

## Mismatch hypothesis

- A mismatch between **public understanding of philosophy of science** and **realities of scientific practice**
- This mismatch contributes to public scientific controversies
- In some cases, this mismatch has been deliberately exploited to manufacture doubt

## John: Climategate vs. transparency

:::: {.columns}
::: {.column}
![\n](img/John.jpg){fig-alt="Photo of Stephen John"}
:::
::: {.column}
*[E]pistemic trust in science may be 'fragile' because based on a false 'folk philosophy of science'.* [@JohnEpistemicTrustEthics2017 6-7]
:::
::::

- Inferring causation from correlation (observational data)
- Refusing (desk rejecting) papers by climate skeptics
- "[I]gnoring recalcitrant or strange data sets" [@JohnEpistemicTrustEthics2017 7]

## Kovaka: Pro-science science denial {.smaller}

:::: {.columns}
::: {.column}
*[P]eople have serious misconceptions about what scientific inquiry ought to look like. Their misconceptions interact with partisan biases to produce denial of human-caused climate change.* [@KovakaClimateChangeDenial2021 1]
:::
::: {.column}
![\n](img/Kovaka.jpg){fig-alt="Photo of Karen Kovaka"}
:::
::::

Misconceptions about

- Objectivity: "scientists as more objective than other people"; value-free ideal
- Scientific method: "a universal, step-by-step scientific method," based on experiments
- Uncertainty and fallibility: Galileo narratives and quest for certainty [@KovakaClimateChangeDenial2021 11-14]

## Nature of science

*[substantial area of research, esp. by education researchers; 
but no scale we found considered role of values in a sophisticated way]

"more than 25 NOS instruments have been developed over the past 50 years" [Neumann citing Lederman 2007]
LiangStudentUnderstandingScience2006 -> Lombrozo et al -> Weisberg et al.
Lombrozo et al version critiqued by Neumann
    - N had science educators map L items to categories based on a lit review of NOS education research (1377, 1384); these don't correspond to L's phil sci-informed categories (1384ff), and don't include issues such as VFI
    - psychometric critique

## *[other empirical research]*

*[Elliott et al; Safford et al]

# Developing a Values in Science Scale (VISS)

## VISS items ##

36 items, 3 each for 12 different topics in the \
science, values, and policy literature

:::: {.columns}
::: {.column}
Aims of science\
Conflicts of interest\
Consensus\
Fact-value distinction\
Fallibilism\
Inductive risk
:::
::: {.column}
Non-subjectivity\
Pluralism\
Scientism\
Standpoint theory\
Technocracy\
Value-free ideal
:::
::::

## Data and methods ##

- Online survey/marketing platform Prolific
- Representative sample of 980 US adults
- Replication of *[elliott et al.]* + VISS + demographics

</br>

- Exploratory factor analysis of VISS items
- Statistical analysis of several findings from *[elliott et al]*

## VISS factors ##

<!--:::: {.columns}
::: {.column width="30%"}
Scientism\
Values in science\
Cynicism\
Power\
Textbook science\
Value-free ideal
:::
::: {.column width="70%"}--->
![\n](../out/02_six_corr_matrix.png){fig-alt="Screenshot of the factor correlation matrix.  The factors are labeled scientism, VIS (values in science), cynicism, power, textbook (science), and VFI (value-free ideal)" height="650"}

### 2022-05-05 scripts/02_EFA.R

@@ -9,6 +9,7 @@ library(gghighlight)
 library(lavaan)
 
 library(visdat)
+library(grid)
 
 library(here)
 
@@ -163,6 +164,91 @@ fitmeasures(fit6, c('chisq','cfi','rmsea','rmsea.ci.upper','srmr','agfi'))
 #+ fig.height = 8, fig.width = 8
 score_grid(fit6, d_vis_cfa)
 
+## Correlation matrix
+fa_vars = c('scientism.1',
+            'fallible.3',
+            'ir.2',
+            'aims.1',
+            'technocracy.2',
+            'factvalue.1',
+            'ir.3',
+            'aims.2',
+            'aims.3',
+            'coi.1',
+            'consensus.3',
+            'factvalue.2',
+            'nonsubj.1',
+            'fallible.2',
+            'ir.1',
+            'coi.2',
+            'stdpt.1',
+            'stdpt.3',
+            'coi.3',
+            'stdpt.2',
+            'consensus.2',
+            'fallible.1',
+            'pluralism.3',
+            'pluralism.1',
+            'vfi.2',
+            'vfi.3',
+            'nonsubj.2',
+            'technocracy.1',
+            'factvalue.3')
+
+text_scientism = textGrob('scientism', gp=gpar(fontsize=13, fontface="bold"))
+text_vis = textGrob('VIS', gp=gpar(fontsize=13, fontface="bold"))
+text_cynicism = textGrob('cynicism', gp=gpar(fontsize=13, fontface="bold"))
+text_power = textGrob('power', gp=gpar(fontsize=13, fontface="bold"))
+text_textbook = textGrob('textbook', gp=gpar(fontsize=13, fontface="bold"))
+text_VFI = textGrob('VFI', gp=gpar(fontsize=13, fontface="bold"))
+
+six_corr_plot = d_vis_cfa |> 
+    cor() |> 
+    as_tibble(rownames = 'item1') |> 
+    pivot_longer(-item1, names_to = 'item2', values_to = 'cor') |> 
+    filter_at(vars(item1, item2), ~ .x %in% fa_vars) |> 
+    mutate(across(c(item1, item2), 
+                  ~ fct_relevel(.x, fa_vars))) |> 
+    ggplot(aes(item1, fct_rev(item2), fill = cor)) +
+    geom_raster() +
[…]

### 2022-05-16 talk/talk.qmd
VMST talk

@@
     | University of California, Merced
-    | *[bitly]*
-toc: true
+    | [`bit.ly/3w1RJSA`](https://bit.ly/3w1RJSA)
+toc: false
+toc-title: "Overview"
 format: 
@@
         slide-number: true
-        center: false
-        <!--margin: 0.0-->
+        self-contained: true
+        theme: [default, theming.scss]
 link-citations: true
+bibliography: ../paper/VISS.yaml
 ---
 
-# Controversy and mismatch # {center="true"}
+## Background & acknowledgment
 
-## Mismatch hypothesis
+- I'm a philosopher of science, STSer, data scientist
+    - Science, values, and policy
+    - Public scientific controversies
+- Emilio is a cognitive psychologist
+    - Conspiracy theories & other "epistemically unwarranted beliefs"
+    - Relationship between prejudice and EUBs
+</br>
+- This project was funded out of my faculty startup funds at UC Merced
 
-- A mismatch between **public understanding of philosophy of science** and **realities of scientific practice**
-- This mismatch contributes to public scientific controversies
-- In some cases, this mismatch has been deliberately exploited to manufacture doubt
+# Controversy and public understanding of science # {center="true"}
 
-## John: Climategate vs. transparency
+## Explaining public scientific controversies {.smaller}
 
-:::: {.columns}
-::: {.column}
-![\n](img/John.jpg){fig-alt="Photo of Stephen John"}
-:::
-::: {.column}
-*[E]pistemic trust in science may be 'fragile' because based on a false 'folk philosophy of science'.* [@JohnEpistemicTrustEthics2017 6-7]
-:::
-::::
+HPSTSers often explain controversies by appeal to public views on\ 
+science, values, and policy issues
 
-- Inferring causation from correlation (observational data)
-- Refusing (desk rejecting) papers by climate skeptics
-- "[I]gnoring recalcitrant or strange data sets" [@JohnEpistemicTrustEthics2017 7]
+<!-- - "The scientific method" [@MercerWhyPopperCan2016; @JohnEpistemicTrustEthics2017] -->
+- Objectivity & value-free ideal [VFI] [@FernandezPintoLegitimizingValuesRegulatory2019; @KovakaClimateChangeDenial2021]
+- Consensus and dissent [@OreskesScientificConsensusClimate2004; @CookQuantifyingConsensusAnthropogenic2013; @JohnEpistemicTrustEthics2017; @MyersConsensusRevisitedQuantifying2021]
+- Paternalism/technocracy/elitism \
+    [@WynneSheepfarmingChernobylCase1989; @JordanTrustworthinessResearchParadigm2011; @LargentVaccineDebateModern2012; @NavinValuesVaccineRefusal2015; @GoldenbergVaccineHesitancyPublic2021]
+- Conflicts of interest, manufacturing doubt or certainty \
+    [@ElliottSciencePolicyTransparency2014; @FernandezPintoCommercialInterestsErosion2020; @GoldenbergVaccineHesitancyPublic2021]
+- Science's historical and/or contemporary role in oppression \
+    [@SchemanEpistemologyResuscitatedObjectivity2001; @WashingtonMedicalApartheidDark2008; @McHughMoreSkinDeep2011; @NelsonBodySoulBlack2011; @GrasswickUnderstandingEpistemicTrust2018;  @GoldenbergVaccineHesitancyPublic2021]
 
-## Kovaka: Pro-science science denial {.smaller}
+## But do the public hold these views?
 
-:::: {.columns}
-::: {.column}
-*[P]eople have serious misconceptions about what scientific inquiry ought to look like.
-Their misconceptions interact with partisan biases to produce denial of human-caused climate change.* [@KovakaClimateChangeDenial2021 1]
-:::
-::: {.column}
-![\n](img/Kovaka.jpg){fig-alt="Photo of Karen Kovaka"}
-:::
-::::
+(As a gross generalization)
 
-Misconceptions about
+- Historians focus on actions of scientists or merchants of doubt
+- Sociologists look at very particular case studies that don't generalize to public at large
+- Philosophers don't offer any empirical evidence
 
-- Objectivity: "scientists as more objective than other people"; value-free ideal
-- Scientific method: "a universal, step-by-step scientific method," based on experiments
-- Uncertainty and fallibility: Galileo narratives and quest for certainty [@KovakaClimateChangeDenial2021 11-14]
+## Two notable exceptions {style="font-size:80%"}
 
-## Nature of science
+- <normsize>@WeisbergKnowledgeNatureScience2020</normsize>
+    - Evolution/creationism, climate, vaccines
+    - Using a "Nature of Science" instrument
+      [@LombrozoImportanceUnderstandingNature2008]
+    - But don't include any items from the science, values, and policy literature [[Table 1](https://journals.sagepub.com/na101/home/literatum/publisher/sage/journals/content/pusa/2021/pusa_30_2/0963662520977700/20210201/images/large/10.1177_0963662520977700-table1.jpeg){target="_blank"}]
 
-*[substantial area of research, esp. by education researchers; 
-but no scale we found considered role of values in a sophisticated way]
-
-"more than 25 NOS instruments have been developed over the past 50 years" [Neumann citing Lederman 2007]
-LiangStudentUnderstandingScience2006 -> Lombrozo et al -> Weisberg et al.
-Lombrozo et al version critiqued by Neumann
-    - N had science educators map L items to categories based on a lit review of NOS education research (1377, 1384); these don't correspond to L's phil sci-informed categories (1384ff), and don't include issues such as VFI
-    - psychometric critique
-
-## *[other empirical research]*
-
-*[Elliott et al; Safford et al]
+- <normsize>@ElliottValuesEnvironmentalResearch2017</normsize>
+    - Does transparency about values in science reduce trust? 
+    - Online survey experiment
+    - But don't ask about participants' views on, eg, VFI
 
@@
 
-36 items, 3 each for 12 different topics in the \
+[36 items](#prompts), 3 each for 12 different topics in the \
 science, values, and policy literature
@@
 
-- Online survey/marketing platform Prolific
+- Online survey platform Prolific
 - Representative sample of 980 US adults
-- Replication of *[elliott et al.]* + VISS + demographics
+- VISS + demographics + replication of <normsize>@ElliottValuesEnvironmentalResearch2017</normsize>
+- Approved by UC Merced IRB
 
@@
 
-- Exploratory factor analysis of VISS items
-- Statistical analysis of several findings from *[elliott et al]*
+- [Exploratory factor analysis](#factor-analysis-results) of VISS items
+- [Regression analysis](#replication-of-elliottvaluesenvironmentalresearch2017) of major findings from <normsize>@ElliottValuesEnvironmentalResearch2017</normsize>
 
-## VISS factors ##
+## Agreement/disagreement: </br> All VISS items ##
 
-<!--:::: {.columns}
-::: {.column width="30%"}
-Scientism\
-Values in science\
-Cynicism\
-Power\
-Textbook science\
-Value-free ideal
+```{r, fig.height=5.75}
+library(plotly)
+library(here)
+library(readr)
+
+barplotly = read_rds(here('out', '02_div_barplot.Rds'))
+barplotly
+```
+
+## Agreement/disagreement: </br>Top & bottom 5 ##
+
+```{r}
+library(gt)
+
+topbottom = read_rds(here('out', '02_topbottom.Rds'))
+
+topbottom |> 
+    group_by(group) |> 
+    gt() |> 
+    cols_label(agree_share = md('**(strongly)\nagree**'), 
+               item = md('**item**'), 
+               prompt = md('**prompt**')) |> 
+    fmt_percent(agree_share, decimals = 0) |> 
+    tab_style(cell_fill(color = 'blue', alpha = .25), 
+              cells_body(rows = group == 'top 5')) |> 
+    tab_style(cell_fill(color = 'red', alpha = .25), 
+                        cells_body(rows = group == 'bottom 5'))
+```
+
+## Ambiguous prompts {style="font-size:70%"}
+
+ir.1
+: Standards of scientific evidence are different in different situations.
+
+</br>
+Does this mean
+
+- *<blue>legitimate</blue>* standards of evidence can be *<blue>appropriately</blue>* different in different situations (eg, because of different downstream non-epistemic consequences); **or**
+- scientists *<red>illegitimately</red>* use different standards of evidence in different situations (eg, to reach their preferred conclusion)?
+
+</br>
+
+- Measurement error:  \
+    How many participants understood the prompt one way vs. the other? 
+- Construct validity:  \
+    Does the item measure the thing you intended to measure? 
+- Requires understanding how participants interpret prompts
+
+# Replication of <normsize>@ElliottValuesEnvironmentalResearch2017</normsize>
+
+## Dr. Riley Spence and BPA {.smaller}
+
+:::: columns
+::: {.column width="65%"}
+Slide attributed to (fictional) Dr. Riley Spence
+
+![](img/cond4.png){fig-alt="Example of a stimulus from Elliott et al.
+(2017).
+Slide reads: \"My conclusion.
+Protecting public health should be a top national priority.
+I examined the scientific evidence on potential health risks of BPA.
+I conclude that BPA in consumer products is not causing harm to people.\"" style="border: 1px solid black"}
 :::
-::: {.column width="70%"}--->
+::: {.column width="35%"}
+- 6 different slides
+    - No values disclosure; protecting public health; promoting economic growth
+    - BPA is/is not causing harm
+- Muenster Epistemic Trustworthiness Inventory [@HendriksMeasuringLaypeopleTrust2015]
+    - eg, competent-incompetent, 1-7 scale
+:::
+::::
+
+## Elliott et al. major findings
+
+Transparency penalty
+: If the scientist discloses values, participants tend to perceive them as less trustworthy
+
+Shared values effect
+: If the participant and scientist share the same values, participants tend to perceive the scientist as more trustworthy
+
+## Transparency penalty?
+
+![](../out/04_transparency.png){fig-alt="Dotplot of trust ratings for Elliott et al.
+(2017) and our data, broken out by whether or not the scientist disclosed values.
+Elliott et al.
+found an estimated effect of transparency of about -0.5 on a 1-7 scale; our data shows a negligible difference."}
+
+## Shared values effect?
+
+![](../out/04_shared_part.png){fig-alt="Dotplot of trust ratings for our data, broken out by participant values (economic growth or public health) and whether they share disclosed values with the scientist.
+The effect of shared values appears to be positive or negative depending on the participants' values."}
+
+## Scientist values effect
+
+![](../out/04_sci_part.png){fig-alt="Dotplot of trust ratings for our data, broken out by participant values (economic growth or public health) and the scientists' disclosed values.
+Scientist valuing public health appears to increase trust, about 0.7 on a 1-7 scale, for both public health and economic growth participants."}
+
+# Future work #
+
+## Future work {}
+
+- Qualitative designs to revise prompts
+    - Paraphrase the prompt; think aloud
+    - Likely will reduce measurement error
+- Other endpoints
+    - Dueling expert scenarios
+    - Trust in science-policy institutions (EPA) rather than individual scientists [@BrownTrustExpertiseScientific2022]
+    - Policy support
+
+# Fin # {.unlisted visibility='uncounted'}
+
+[`bit.ly/3w1RJSA`](https://bit.ly/3w1RJSA)
+
+# Extra slides # {.unlisted visibility='uncounted'}
+
+## Prompts {.scrollable visibility='uncounted'}
+
+aims-1
+: The knowledge produced by scientific research is valuable even when it has no practical applications.
+
+aims-2
+: A primary aim of science is to improve our understanding of threats to human health and the environment.
+
+aims-3
+: A primary aim of science is to develop new technology and stimulate economic growth.
+
+coi-1
+: Scientists will report conclusions that they think will get them more grant money even if the data does not fully support that conclusion.
+
+coi-2
+: Special interests can always find an scientist-for-hire who will support their point of view.
+
+coi-3
+: Research carried out by private companies is less trustworthy than research carried out by public instutitions.
+
+consensus-1
+: Progress in science happens when a lone dissenter challenges popular ideas.^K^
+
+consensus-2
+: Scientists rarely disagree with each other about the answers to scientific questions.
+
+consensus-3
+: The consensus of the scientific community is based on social status and prestige rather than evidence.
+
+factvalue-1
+: Moral and social values are the kinds of things that can be confirmed or challenged by scientific evidence.
+
+factvalue-2
+: If scientific findings conflict with important values that we hold, it's appropriate to be skeptical of the science.
+
+factvalue-3
+: Someone's moral and social values are more like their taste in music than scientific facts.
+
+fallible-1
+: Once a scientific theory has been established, it is never changed.^K^
+
+fallible-2
+: When a scientific theory changes or is revised, it means that the research that went into it initially was flawed.
+
+fallible-3
+: Scientific findings can be good enough to act on even when they're not entirely certain.
+
+ir-1
+: Standards of scientific evidence are different in different situations.
+
+ir-2
+: Scientists are not responsible for harmful effects of the technologies they develop.
+
+ir-3
+: When scientific research has potential consequences for society, scientists must take these consequences into account when drawing conclusions.
+
+nonsubj-1
+: The results of scientific studies are always ambiguous and require interpretation.
+
+nonsubj-2
+: Good scientific research is always free of assumptions and speculation.
+
+nonsubj-3
+: Scientific conclusions are never a matter of interpretation or judgment.
+
+pluralism-1
+: Scientific investigations always require laboratory experiments.^K^
+
+pluralism-2
+: Scientific research is always conducted in the following order: 1. Observation, 2. Hypothesis, 3. Experiment, 4. Conclusion^K^
+
+pluralism-3
+: All scientists use the same strict requirements for determining when empirical data confirm a tested hypothesis.
+
+scientism-1
+: Science is more reliable than any other way of producing knowledge.
+
+scientism-2
+: Everyone, whatever their worldview, should accept evidence-based scientific conclusions
+
+scientism-3
+: Science is the only source of justified belief or knowledge about ourselves and the world.
+
+stdpt-1
+: A scientific understanding of the social word depends on the background of the scientists carrying out the research.
+
+stdpt-2
+: People on the margins of society often have a unique understanding of that society that can benefit scientific research.
+
+stdpt-3
+: Science has a long history of ignoring the unique insights of women and people of color.
+
+technocracy-1
+: Public policy should wait until all relevant scientific questions have been settled.
+
+technocracy-2
+: When it comes to public policy, scientific evidence is more important than public opinion.
+
+technocrac-3
+: Scientists have too much influence over public policy today.
+
+vfi-1
+: Scientists are more objective than other people.^K^
+
+vfi-2
+: Scientists do not use imagination or creativity because doing so interferes with objectivity.^K^
+
+vfi-3
+: The evaluation and acceptance of scientific results must not be influenced by social and ethical values.^K^
+
+# Factor analysis results {visibility='uncounted'}
+
+## VISS factors ## {visibility='uncounted'}
+
 ![\n](../out/02_six_corr_matrix.png){fig-alt="Screenshot of the factor correlation matrix.
-The factors are labeled scientism, VIS (values in science), cynicism, power, textbook (science), and VFI (value-free ideal)" height="650"}
+The factors are labeled scientism, VIS (values in science), cynicism, power, textbook (science), and VFI (value-free ideal)"}
+
+::: {style="text-align:center"}
+<span style="color:red">⚠️ Psychometric properties of these factors aren't great ⚠️</span>
+:::
+
+## Factor analysis fit indices ## {visibility='uncounted' style="font-size:90%"}
+
+How well does the fitted FA model approximate the observed covariance structure? [@HuFitIndicesCovariance1998]
+
+</br>
+
+| statistic                                       | range   | want  | our model |
+|-------------------------------------------------|---------|-------|-----------|
+| Comparative fit index (CFI)                     | 0-1     | large | 0.6       |
+| Adjusted-goodness-of-fit index (AGFI)           | 0-1     | large | 0.8       |
+| Root-mean-square error of approximation (RMSEA) | >0      | small | 0.07      |
+| Standardized root-mean-square residual (SRMSR)  | >0      | small | 0.09      |
+
+## Scientism ## {visibility='uncounted'}
+
+:::: {.columns style="font-size:50%"}
+::: {.column}
+![\n](../out/02_scientism.png){fig-alt="Density plot for scientism factor scores.  The mode is just below 3.5, and values range from 1 to 5." width="100%"}
+
+- Science is more reliable than any other way of producing knowledge.
+- Scientific findings can be good enough to act on even when they're not entirely certain. 
+- The knowledge produced by scientific research is valuable even when it has no practical applications.
+
+:::
+::: {.column}
+![\n](../out/04_scientism_polid.png){fig-alt="Plot of scientism against political ideology.
+There's a weak negative association: on average conservatives are about 0.5 point less scientistic than liberals." width="100%"}
+
+- Scientists are not responsible for harmful effects of the technologies they develop.
+- When it comes to public policy, scientific evidence is more important than public opinion. 
+- Moral and social values are the kinds of things that can be confirmed or challenged by scientific evidence. 
+:::
+::::
+
+## Cynicism ## {visibility='uncounted'}
+
+:::: {.columns style="font-size:40%"}
+::: {.column}
+![\n](../out/02_cynicism.png){fig-alt="Density plot for cynicism factor scores.  The mode is about 2.5, and values range from 1 to about 4.5." width="100%"}
+
+- Scientists will report conclusions that they think will get them more grant money even if the data does not fully support that conclusion.^K^  
+- The consensus of the scientific community is based on social status and prestige rather than evidence.
+- If scientific findings conflict with important values that we hold, it's appropriate to be skeptical of the science.
+- The results of scientific studies are always ambiguous and require interpretation.
+
+:::
+::: {.column}
+![\n](../out/04_cynicism_polid.png){fig-alt="Plot of cynicism against political ideology.
+There's a positive association: on average conservatives are about 1 point more cynical than liberals." width="100%"}
+
+- When a scientific theory changes or is revised, it means that the research that went into it initially was flawed. 
+- Special interests can always find an scientist-for-hire who will support their point of view. 
+- <span style="color:red">Standards of scientific evidence are different in different situations.</span>
+- <span style="color:red">A scientific understanding of the social world depends on the background of the scientists carrying out the research.</span> 
+:::
+::::
+
+## Factors and trust ## {visibility='uncounted' .smaller}
+
+![](../out/04_fa_effects.png){fig-alt="Estimated effects plot for all six VISS factors.
+Power, VFI, and VIS factors are essentially flat.
+Scientism and textbook are associated with increased trust, while cynicism is associated with decreased trust.
+However, all effects are modest, highly unlikely to be greater than +/- 1 on the 1-7 trust scale."}
+
+# References # {.unlisted visibility='uncounted' style="font-size:50%"}

### 2022-05-16 scripts/04_dag.R

@@ -15,6 +15,7 @@
 library(tidyverse)
 theme_set(theme_minimal())
 library(broom)
+library(patchwork)
 library(ggbeeswarm)
 library(dagitty)
 library(ggdag)
@@ -158,15 +159,34 @@ plot_estimate(list(emad = model_b_emad,
 #' # C. Transparency penalty #
 #' *Scientists who disclose values are perceived as less trustworthy than scientists who do not.* 
 
-ggplot(emad_df, aes(disclosure, pa_mean)) +
-    geom_violin(draw_quantiles = .5) +
-    geom_beeswarm()
-
-ggplot(dataf, aes(disclosure, meti_mean)) +
-    geom_violin(draw_quantiles = .5) +
-    geom_beeswarm()
-
-
+{
+    trans_plot_emad = ggplot(emad_df, aes(disclosure, pa_mean)) +
+        # geom_violin(draw_quantiles = .5) +
+        geom_beeswarm(alpha = .25, size = .3) +
+        stat_summary(fun.data = 'mean_cl_boot', color = 'red', 
+                     size = 1.5, fatten = 1/2) +
+        stat_summary(geom = 'line', group = 1L, color = 'red') +
+        labs(y = 'trust')
+    trans_plot_emad
+    
+    
+    trans_plot_us = ggplot(dataf, aes(disclosure, meti_mean)) +
+        geom_beeswarm(alpha = .25, size = .3) +
+        stat_summary(fun.data = 'mean_cl_boot', color = 'red', 
+                     size = 1.5, fatten = 1/2) +
+        stat_summary(geom = 'line', group = 1L, color = 'red') +
+        labs(y = 'trust')
+    trans_plot_us
+    
+    trans_plot_emad + 
+        ggtitle('Elliott et al. (2017)') +
+        trans_plot_us +
+        ggtitle('Hicks and Lobato')
+    
+    ggsave(here(out_dir, '04_transparency.png'), 
+           height = 3, width = 6, scale = 1, 
+           bg = 'white')
+}
 #' Again, disclosure/transparency is experimentally controlled, so no adjustment is required. 
 dag |> 
     add_arrows('disclose -> METI') |> 
@@ -284,8 +304,18 @@ plot_estimate(list('emad' = model_emad_s,
 dataf |> 
     filter(disclosure, !is.na(part_values)) |> 
     ggplot(aes(sci_values, meti_mean)) +
[…]

### 2022-05-17 talk/talk.qmd
talk to GH pages

@@
     | University of California, Merced
-    | [`bit.ly/3w1RJSA`](https://bit.ly/3w1RJSA)
+    | [`values-in-science-scale.netlify.app`](https://values-in-science-scale.netlify.app)
 toc: false
@@
 
-[`bit.ly/3w1RJSA`](https://bit.ly/3w1RJSA)
+[`values-in-science-scale.netlify.app`](https://values-in-science-scale.netlify.app)
 

### 2022-05-17 scripts/04_dag.R

@@ -35,12 +35,18 @@ data_dir = here('data')
 ## Load data ----
 ## Elliott et al. data
 emad_df = read_rds(here(data_dir, 'emad.Rds'))
-## VISS six-factor model
+## VISS item labels and six-factor model
+source(here('R', 'vis_labels.R'))
+vis_labels_vec = vis_labels |> 
+    pull(item) |> 
+    set_names(vis_labels$tag)
+
 viss_df = read_csv(here(data_dir, 'fa_six.csv')) |> 
     select(pid, starts_with('fa_'))
 ## Our data
 dataf = read_rds(here(data_dir, 'data.Rds')) |> 
-    left_join(viss_df, by = 'pid')
+    left_join(viss_df, by = 'pid') |> 
+    rename(vis_labels_vec)
 
 ggplot() +
     geom_violin(aes(x = 'EMAD', pa_mean), 

### 2022-06-08 talk/talk.qmd
as presented at VMST

@@
         self-contained: true
-        theme: [default, theming.scss]
+        theme: [black, theming.scss]
 link-citations: true
@@
 - Emilio is a cognitive psychologist
-    - Conspiracy theories & other "epistemically unwarranted beliefs"
+    - Conspiracy theories & other \
+      "epistemically unwarranted beliefs"
     - Relationship between prejudice and EUBs
@@
     - Dueling expert scenarios
-    - Trust in science-policy institutions (EPA) rather than individual scientists [@BrownTrustExpertiseScientific2022]
+    - Trust in science-policy institutions (EPA) rather than individual scientists [@BrownTrustExpertiseScientific2022\; Biddle yesterday]
     - Policy support

### 2022-06-08 scripts/04_dag.R

@@ -447,7 +447,9 @@ dataf |>
     labs(x = 'shared values', 
          y = 'trust')
 
-ggsave(here(out_dir, '04_shared_part.png'), height = 3, width = 6, scale = 1)
+ggsave(here(out_dir, '04_shared_part.png'), 
+       height = 3, width = 6, scale = 1, 
+       bg = 'white')
 
 #' And political ideology appears to be either entirely upstream or independent to this
 dataf |> 
@@ -573,7 +575,8 @@ dataf |>
          y = 'cynicism')
 
 ggsave(here(out_dir, '04_cynicism_polid.png'), 
-       height = 3, width = 4, scale = 1)
+       height = 3, width = 4, scale = 1, 
+       bg = 'white')
 
 dataf |> 
     # filter(!disclosure) |> 
@@ -591,7 +594,8 @@ dataf |>
          y = 'scientism')
 
 ggsave(here(out_dir, '04_scientism_polid.png'), 
-       height = 3, width = 4, scale = 1)
+       height = 3, width = 4, scale = 1, 
+       bg = 'white')
 
 
 
@@ -668,7 +672,8 @@ ggplot() +
     coord_cartesian(ylim = c(3, 7))
 
 ggsave(here(out_dir, '04_scientism_trust.png'), 
-       height = 3, width = 4, scale = 1)
+       height = 3, width = 4, scale = 1, 
+       bg = 'white')
 
 ggplot() +
     geom_point(aes(fa_cynicism, meti_mean), 
@@ -690,7 +695,8 @@ ggplot() +
     coord_cartesian(ylim = c(3, 7))
 
 ggsave(here(out_dir, '04_cynicism_trust.png'), 
-       height = 3, width = 4, scale = 1)
+       height = 3, width = 4, scale = 1, 
+       bg = 'white')
 
 
 #' ## B (Consumer Risk) and C (Disclosure) ##
@@ -824,7 +830,8 @@ ggplot() +
     coord_cartesian(ylim = c(1, 7))
 
 ggsave(here(out_dir, '04_cynicism_scivalues.png'), 
-       height = 3, width = 5)
+       height = 3, width = 5, 
+       bg = 'white')
 
 ## Associations aren't as strong in the data plot, presum. bc confounding
 # dataf |> 

### 2022-06-08 scripts/02_EFA.R
quick cleaning out VISS

@@ -1,367 +1,0 @@
-library(car)
-library(tidyverse)
-theme_set(theme_minimal())
-library(psych)
-library(pander)
-library(gvlma)
-library(tables)
-library(gghighlight)
-library(lavaan)
-
-library(visdat)
-library(grid)
-library(plotly)
-
-library(here)
-
-data_dir = here("data")
-
-source(here('R', 'vis_labels.R'))
-source(here('R', 'loading_table.R'))
-source(here('R', 'scores.R'))
-
-## Should the script overwrite existing loadings tables? 
-overwrite_loading_tables = FALSE
-
-## Load data ----
-d_clean = read_rds(here(data_dir, 'data.Rds')) |> 
-    select(pid, ViS01:ViS36) |> 
-    mutate(across(-pid, as.numeric)) |> 
-    set_names('pid', vis_labels$tag)
-
-## 0.3% missing values
-vis_miss(d_clean)
-## Items with missing values don't appear to cluster
-vis_miss(d_clean, cluster = TRUE)
-
-## 906 complete cases; 
-## 69 missing 1, 11 missing 2, 2 missing 3
-d_clean |> 
-    rowwise() |> 
-    summarize(missing = sum(is.na(c_across(-pid)))) |> 
-    count(missing)
-
-
-#EFA for the ViS items, checking factor structure; need to check if N is large enough for a split to do CFA on
-d_vis <- d_clean |> 
-    column_to_rownames('pid') |> 
-    na.omit()
-
-#if N is big enough to reasonably split for EFA and CFA,
-set.seed(032585) #for reproducibility
-dummy_sep <- rbinom(nrow(d_vis), 1, 0.5) #create dummy indicator to randomly split sample
-d_vis_efa <- d_vis[dummy_sep == 0, ] #extract data where dummy == 0
-d_vis_cfa <- d_vis[dummy_sep == 1, ] #extract data where dummy == 1
-
-vis_cor(d_vis_efa)
-
-
-## Descriptive visualizations ----
-## Share of respondents (agreeing or strongly agreeing)
-d_vis |> 
-# d_vis_efa |> 
-    pivot_longer(everything(), names_to = 'item', values_to = 'value') |> 
-    count(item, value) |> 
-    group_by(item) |> 
-    mutate(share = n / sum(n)) |> 
-    ungroup() |> 
-    # filter(value >= 4) |> 
[…]

### 2022-06-08 scripts/04_dag.R

@@ -35,18 +35,9 @@ data_dir = here('data')
 ## Load data ----
 ## Elliott et al. data
 emad_df = read_rds(here(data_dir, 'emad.Rds'))
-## VISS item labels and six-factor model
-source(here('R', 'vis_labels.R'))
-vis_labels_vec = vis_labels |> 
-    pull(item) |> 
-    set_names(vis_labels$tag)
 
-viss_df = read_csv(here(data_dir, 'fa_six.csv')) |> 
-    select(pid, starts_with('fa_'))
 ## Our data
-dataf = read_rds(here(data_dir, 'data.Rds')) |> 
-    left_join(viss_df, by = 'pid') |> 
-    rename(vis_labels_vec)
+dataf = read_rds(here(data_dir, 'data.Rds'))
 
 ggplot() +
     geom_violin(aes(x = 'EMAD', pa_mean), 
@@ -507,337 +498,5 @@ dataf %>%
     # summary()
     plot_predictions(c('sci_values', 'part_values'), 
                  interaction_ci = TRUE)
-
-
-## F. VISS ----
-#' # F. VISS #
-#' *Effects c-e will vary depending the participant's views of the role of values in science (the latent variables from the six-factor model)*
-#' 
-#' Probably due to low psychometric qualities of VIS factors, seeing some inconsistency in *primary* effects (eg, primary effects estimates for fa_power across some of the models below).  
-#' Don't focus on outcomes of any particular model; at best these models provide very weak evidence
-#' Motivating future research to improve quality of prompts
-
-
-#' ## VISS and demographics ##
-#' Continuous and ordinal demographics
-#+ fig.width = 10, fig.height = 8
-dataf |> 
-    ggplot(aes(x = .panel_x, y = .panel_y)) +
-    ggpubr::stat_cor(label.x.npc = 'center', label.y.npc = 'center', 
-                     geom = 'rect', alpha = .8,
-                     mapping = aes(fill = after_stat(r), 
-                                   x = .panel_x, y = .panel_y,
-                                   xmin = -Inf, xmax = Inf, 
-                                   ymin = -Inf, ymax = Inf)) +
-    geom_jitter(alpha = .1) +
-    geom_smooth(method = 'lm') +
-    ggpubr::stat_cor(label.x.npc = 'right', label.y.npc = 'bottom', 
-                     geom = 'label', hjust = 'right', vjust = 'bottom',
-                     size = 5, 
-                     r.accuracy = .01,
-                     mapping = aes(label = after_stat(r.label), 
[…]

### 2022-06-20 paper/paper.Qmd
fill in tables

---
title: "Something something values disclosure replication"
author: "Daniel J. Hicks and Emilio Lobato"
abstract: |
  abstract
---

# Introduction #

Over the past 15 years, many philosophers of science have rejected the ideal of value-free science and related, traditional understandings of objectivity and political neutrality of science. *[Douglas and Elliott]*  According to the value-free ideal, social and political values — such as feminism, environmentalism, or the protection of human health — have no legitimate role to play in the evaluation of scientific hypotheses.  The value-free ideal is compatible with allowing social and political values to play important roles earlier and later in inquiry.  Specifically, these values may legitimately shape the content and framing of research questions — researchers might decide to investigate whether chemical X causes cancer out of a concern to protect human health — and will be essential when scientific findings are used to inform public policy — say, banning the use of chemical X.  But, according to the value-free ideal, these values must not influence the collection and analysis of data and the process of reaching an overall conclusion about whether or not chemical X causes cancer.

Challenges to the value-free ideal argue that at least some social and political values may, or even should, play a role in the evaluation of scientific hypotheses.  Keeping with the example, the question of whether chemical X causes cancer has much more significant social and political implications than the question of whether chemical X fluoresces under ultraviolet light.  So, in the interest of protecting human health, it would be appropriate to *scientifically* accept the hypothesis that chemical X is carcinogenic based on relatively provisional evidence *[cite]*.

While this kind of argument indicates that at least some social and political values may or should play at least some role in the evaluation of at least some scientific hypotheses, it doesn't provide us with much positive guidance:  which values, playing which roles, in the evaluation of which hypotheses?  *[New Direction]*  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research *[Elliott]*.  But some philosophers have objected to this transparency proposal, arguing that it might actually undermine trust in science.  According to this objection, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy *[John, Kovaka]*.

This objection is an empirical prediction: if scientists disclose their values, they will be perceived as less trustworthy.  @ElliottValuesEnvironmentalResearch2017 conducted an online survey study to evaluate this prediction.  However, these authors collected data from a somewhat small sample using Amazon Mechanical Turk, and adopted an analytical approach that significantly diluted their sample across different conditions.

In this paper, we report the results of a replication of @ElliottValuesEnvironmentalResearch2017 study 1, using a larger, more reliable sample and a more statistically efficient analytical approach.

# Methods and Materials #

## Experimental design ##

We use the same experimental stimulus and design as @ElliottValuesEnvironmentalResearch2017, embedded within a larger survey with other results not reported here.  In the replication component, participants are randomly assigned to one of $3 \times 2$ (values \times harm) conditions.  In each condition, participants are first shown a single presentation slide and the following explanatory text:

> For several decades, a scientist named Dr. Riley Spence has been doing research on chemicals used in consumer products. One chemical—Bisphenol A, popularly known as BPA—is found in a large number of consumer products and is suspected of posing risks to human health. Yet, scientists do not agree about these possible health risks. Dr. Spence recently gave a public talk in Washington, D.C., about BPA research. Here is the final slide from Dr. Spence's presentation.

| No other information about the (fictional) Dr. Spence is provided.  The content of the slide varies across the conditions.  Each slide has the header "My conclusion," followed by a list of 2 or 3 bullets.  The first bullet, if present, makes a values disclosure, stating that either economic growth or public health should be a top national priority.  This first bullet is not present in the "no disclosure" values condition.  The second bullet, identical across conditions, states "I examined the scientific evidence on potential health risks of BPA."  The third bullet concludes that BPA either does or does not cause harm to people, depending on whether the subject is in the harm or *[not-harm]* condition.

After viewing the slide, the participant is then asked to rate Dr. Spence's trustworthiness using a semantic differential scale.  @ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @ElliottValuesEnvironmentalResearch2017], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.  On the METI, participants rate the target scientist (the fictional Dr. Spence) on a 1-7 scale for 14 items, where each item is anchored at the ends by a pair of words, such as competent-incompetent or responsible-irresponsible.  In both @ElliottValuesEnvironmentalResearch2017 and our replication analysis, all items were averaged together to form a 1-7 composite measure of trustworthiness.  To aid interpretation, in the analysis the direction of the scale has been set so that increased values correspond to increasing perceived trustworthiness.

Ethicists make a key distinction between trust and trustworthiness *[Baier]*.  Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.  Trustworthiness is an assessment, of whether or not that trust is appropriate.  In the experiment, participants might trust Dr. Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa; though we would typically expect these two to go together.  As its name indicates, METI assesses perceived trustworthiness, not trust (such as accepting Dr. Spencer's conclusion about BPA).

@ElliottValuesEnvironmentalResearch2017 explain that they selected BPA as a complex, ongoing, public scientific controversy.  For our replication, we chose to keep BPA, rather than switching to a different public scientific controversy with a higher profile in 2021, such as climate change, police violence, voter fraud, or any of numerous aspects of the Covid-19 pandemic.  All of these controversies are highly politically charged, with prominent public experts and counterexperts *[cite]*.  While we anticipated some effects of political partisanship in the BPA case, we felt it would be less likely to swamp the values disclosure that was our primary interest.

After filling in the METI, participants provided demographic information and other sections of the survey that are not examined here.  Due to researcher error, a question about the participants' values (whether they prioritize economic growth or public health) was omitted in the first wave of data collection; this question was asked in a followup wave.

## Replication hypotheses and analytical approach ##

We identified 5 major findings from @ElliottValuesEnvironmentalResearch2017 for our replication attempt:

*[consistent labeling of hypotheses]*

a. Modest correlation between values and ideology:  
	(i) Political liberals are more likely to prioritize public health over economic growth, compared to political conservatives; but (ii) a majority of political conservatives prioritize public health.  
b. Consumer risk sensitivity:  Scientists who find that a chemical harms human health are perceived as more trustworthy than scientists who find that a chemical does not cause harm. 
c. Transparency penalty: Scientists who disclose values are perceived as less trustworthy than scientists who do not.  
d. Shared values:  Given that the scientist discloses values, if the participant and the scientist share the same values, the scientist is perceived as more trustworthy than if the participant and scientist have discordant values.  
e. Variation in effects:  The magnitude of the effects above vary depending on whether the participant prioritizes public health or economic growth.

Hypothesis c, the transparency penalty, corresponds to the objection to transparency:  disclosing values undermines trust in science.  However, the shared values effect works in the opposite direction, counteracting the transparency penalty.

*[hypothesis a]*

Hypotheses b-e were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed a priori to identify appropriate adjustments (covariates) for hypotheses d and e.

## Participants ##

Participants were recruited using the online survey platform Prolific, and the survey was administered in a web browser using Qualtrics.  Prolific has an option to draw samples that are balanced to be representative by age, binary gender, and a 5-category race variable (taking values Asian, Black, Mixed, Other, and White) for US adults *[https://researcher-help.prolific.co/hc/en-gb/articles/360019238413-Representative-samples-FAQ]*.  A recent analysis finds that Prolific produces substantially higher quality data than Amazon Mechanical Turk for online survey studies, though three of the five authors are affiliated with Prolific [@PeerDataQualityPlatforms2021].  Preliminary power analysis recommended a sample of approximately 1,000 participants to reliably detect non-interaction effects (hypotheses a-d).

The study was approved by the UC Merced IRB on August 17, 2021, and data collection ran October 18-20, 2021.  As mentioned above, due to researcher error a question about participants' values was not included in the original survey.  A followup survey asking this question of the same group of participants was conducted December 8, 2021 through March 5, 2022.

# Results #

After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample.  660 participants (67%) were randomly assigned to a values disclosure condition (scientist prioritizes either economic growth or public health); 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).  Consequently, subsamples for hypotheses d and e were substantially smaller than the full analysis sample.

*[participant demographics]*

Critically, our data are unlikely to be representative by education level and political ideology.  In 2021, about 9% of US adults 25 or over had a less than high school education, and 38% had a Bachelor's degree or higher.  Only 1% of our participants reported a less than high school education, and 57% reported a Bachelor's degree or higher.  *[<https://www.census.gov/library/visualizations/time-series/demo/cps-historical-time-series.html> fig. 2]*  For political ideology, the General Social Survey has consistently found over several decades that about 30% of US adults identify as liberal, about 30% identify as conservative, and about 40% as moderate *[https://gssdataexplorer.norc.org/variables/178/vshow]*.  Among our participants, liberals outnumber conservatives *[by over 2:1; hypothesis a fig]*.  Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling *[https://academic.oup.com/poq/article/82/1/1/4837043?login=false]*.  In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement]*.  This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  General Social Survey data shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public controversies align with liberal-conservative partisanship *[cite]*.  Insofar as some political conservatives are both less likely to participate in studies on Prolific (or at least in our particular study) and likely to perceive Spence as less trustworthy, this will lead to omitted variable bias for analyses that require adjustment by political ideology.  *[however, most of our analyses do not require this adjustment]*

We tested the hypothesis of a modest correlation between values and ideology in two ways. First, to test (H1a) whether political liberals are more likely to prioritize public health over economic growth compared to conservatives, we conducted a Spearman's rank order correlation. Results revealed a significant correlation in line with the hypothesis, Spearman's rho = -.47, p < .001. Political liberals were more likely than conservatives to value public health over economic growth. To test (H1b) that a majority of political conservatives priortize public health over economic growth we cross-tabulated the data. Results revealed that, contrary to the hypothesis, slightly more than half of the self-reported political conservatives in our sample reported valuing economic growth (51.7%) over public health (48.3%). See @fig-part-values.

![Participant values by political ideology. (A) Absolute counts, (B) shares within political ideology categories.  Panel (A) shows that our sample substantially over-represents political liberals.](../out/03_part_values.png){#fig-part-values}

Next, we tested the hypotheses that (H2) scientists who find a chemical harms human health are perceived as more trustworthy than scientist who find that a chemical does not cause harm and (H3) a scientist who discloses values are perceived as less trustworthy than a scientist who does not. For this analysis, we regressed participants' METI ratings onto both the Conclusions and Disclosure experimental conditions. The full model was significant, adj. R^2 = .148, F(2, 985) = 86.59, p < .001 (see @tbl-bc). Specifically, results revealed that the conclusions reported by the scientist predicted participants' perceived trustworthiness in line with our hypothesis. Participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm. By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values are perceived as less trustworthy than a scientist who does not disclose values.

```{r}
#| echo: false
#| label: tbl-bc
#| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
bc_tbl = readRDS('../out/03_bc_tbl.Rds')
bc_tbl
```

Next, we tested the hypothesis that (H4) if the participant and scientist share the same values, the scientist is perceived as more trustworthy than if the participant and scientist do not share the same values. For this and related analyses, we only included data from participants who were assigned to either of the Disclosure conditions and self-reported their own values, reducing the sample to 567. Following this, we created a new Shared Values variable as a composite of the participants' reported values and the scientist's values. For this analysis, an a priori directed acyclic graphic (DAG) indicated that a univariate regression of participants' METI rating onto Shared Values would produce a biased estimate, that adjustments were required for both the participant's and scientist's values, and that including demographic variables would not effect this estimate [@fig-shared-dag].  While the univariate model was significant, adj. R^2 = .029, F(1, 565) = 18.0, p < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.  Rather, only the scientist's disclosed values significantly predicted how trustworthy participants rated the scientist, such that a scientist who disclosed valuing public health was rated as more trustworthy than a scientist who disclosed valuing economic growth.  This scientist values effect was robust across alternative model specifications, with an estimated effect of 0.6 (95% CI 0.4-0.8) [@tbl-shared models 2-5], again consistent with the DAG.

![Directed acyclic graph (DAG) for analysis of the shared values effect.  If the DAG is faithful, adjusting for scientist values (`sci_values`) and participant values (`part_values`) is sufficient for estimating the effect of shared values on perceived trustworthiness (`METI`).  In particular, adjusting for demographics should not change the regression coefficient on shared values.](../out/03_shared_values_dag.png){#fig-shared-dag}

```{r}
#| echo: false
#| label: tbl-shared
#| tbl-cap: "Sequential regression analysis of hypothesis H4 shared values/scientist values effects"
shared_values_tbl = readRDS('../out/03_shared_values_tbl.Rds')
shared_values_tbl
```

## Variation in Effect Sizes ##

We ran the following analyses to test our hypothesis that (H5) the magnitude of the effects found for the tests of H2-H4 vary depending on whether the participant prioritizes public health or economic growth. Results are presented in Table #.

H5b - consumer risk sensitivity
To test whether the findings regarding consumer risk sensitivity vary as a function of participants' values, we regressed participants' METI ratings of the scientist in the stimuli onto the Conclusions condition, Participants' Values variable, and the Conclusions by Participants' values interaction term. The full model was significant, adj. R^2 = .173, F(3, 840) = 59.95, p < .001 [@tbl-eb model 2]. As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm. However, this effect was qualified by a significant interaction with participants' values. *[is this right? Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values.]*

```{r}
#| echo: false
#| label: tbl-eb
#| tbl-cap: "Regression analysis of H5b, interaction between participant values and consumer risk sensitivity"
readRDS('../out/03_eb_tbl.Rds')
```

H5c - transparency penalty
To test whether the findings above about a hypothesized transparency penalty may vary based on participants' values, we regressed participants' METI ratings of the scientist onto the Disclosure condition variable, Participants' Values variable, and the Disclosure by Participants' Values interaction term. The full model was not significant, adj. R^2 < .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2]. As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values.

```{r}
#| echo: false
#| label: tbl-ec
#| tbl-cap: "Regression analysis of H5c, interaction between participant values and disclosure effect"
readRDS('../out/03_ec_tbl.Rds')
```

H5d - shared values
Without adjustments, Shared Values appears to have a substantial interaction with Participant Values:  Shared Values appears to increase perceived trustworthiness for participants who value public health, while decreasing perceived trustworthiness for participants who value economic growth [@fig-shared-part].  However, as indicated by our analysis for a potential shared values effect, the estimated effects for both shared values and participant values are biased if the model is not adjusted for scientist values.  (@tbl-shared shows that the Shared Values estimate is biased.  Shared Values is a collider between Participant Values and Scientist Values; hence, if a model specification includes Participant Values and Shared values but not Scientist Values, there is an open path between Participant Values and METI, resulting in a biased estimate for Participant Values.)  Because Shared Values is the logical biconditional of Participant Values and Scientist Values, and the Shared Values $\times$ Participant Values is the logical conjunction, including all four variables (Participant, Scientist, and Shared Values, along with the interaction term) in a regression model creates perfect collinearity.

![Data, group means, and 95% confidence intervals for H5d, interaction of shared values and participant values.  Panels correspond to participant values.](../out/03_shared_part.png){#fig-shared-part}

As we found above, rather than a shared values effect there appears to be an effect of scientist values.  Reploting the same data as @fig-shared-part based on scientist values, rather than shared values, suggests a more consistent effect across participant values [@fig-sci-part]. Therefore, to test whether an effect of scientist values might vary as a function of participants values, we regressed participants' METI of the scientist in the stimulus onto the Scientist Values variable, Partiicpants' Values variable, and their interaction [@tbl-sci-part]. This model was significant, adj. R^2 .065, F(3, 563) = 14.15, p < .001. But estimates for both the scientist values term and interaction term were highly uncertain and not significant.  *[finish combining with older text below]*

![Data, group means, and 95% confidence intervals for a potential interaction of scientist values and participant values. Panels correspond to participant values. This figure is identical to @fig-shared-part, except the columns in the economic growth panel have been switched.](../out/03_sci_part.png){#fig-sci-part}

```{r}
#| echo: false
#| label: tbl-sci-part
#| caption: "Regression analysis of interaction between scientist values and participant values"
readRDS('../out/03_sci_part_tbl.Rds')
```

To test whether a potential shared values effect might vary as a function of participants' values, we regressed participants' METI ratings of the scientist in the stimuli onto the Shared Values variable, Participants' Values variable, and the Shared Values by Participants' values interaction term. The full model was significant, adj. R^2 = .065, F(3, 563) = 14.15, p < .001 *[tbl]*.  The model revealed a significant interaction between Shared Values and Participants' Values, such that overall Shared Values had a positive effect when the participant valued public health, but a negative effect when the participant valued economic growth *[@fig-shared-part and tbl]*.

such that the highest perceived trustworthiness ratings were from participants who prioritized public health reading about a scientist who disclosed valuing public health. However, as revealed in our analysis for H4, the evidence that shared values between the participant and scientist are what predicts the perceived trustworthiness of a scientist should be interpreted with caution. Because of the inclusion of Participants' values and the Scientist's values in the analysis for H4, the set of predictor variables for that model and the set of predictor variables for this model are different linear combinations of each other. Hence why the resulting full model parameters are identical between the two models. In spite of this, we report the results of this analysis for the sake of completeness, as this was an a priori planned statistical analysis.

Because of the findings from the analysis run to test H4, we conducted an unplanned post hoc analysis to further test the hypothesis that the effects found in the earlier analysis vary as a function of participants' values. For this test, we regression METI scores onto the Scientist's Values experimental condition, the Shared Values composite variable, and their interaction term. This is, as with the previous model and the model reported to test H4, a different linear combination of the same predictor variables and therefore the full model parameters are identical. However, presentation of the model this way renders a clearer picture of what variables we tested seem to be driving perceptions of the trustworthiness of a scientist. The evidence here suggests that if a scientist chooses to disclose their values alongside their presentation of scientific research, what values they share predicts their perceived trustworthiness, and that the scientist in our study who disclosed valuing public health was perceived a significantly more trustworthy than the scientist who disclosed valuing economic growth.

# Conclusion #

### 2022-06-20 R/reg_tbl.R

@@ -1,0 +1,40 @@
+reg_tbl = function(models, 
+                   labs = c('**EMAD**', '**HL replication**'), 
+                   filters = !variable %in% c('gender', 
+                                              'race_ethnicity', 
+                                              'religious_affil')) {
+    # filters = enquos(filters)
+    models |> 
+        map(tbl_regression, intercept = TRUE) |> 
+        map(add_glance_table, include = c(r.squared, nobs, 
+                                          adj.r.squared,
+                                          statistic, p.value)) |> 
+        tbl_merge(tab_spanner = labs) |> 
+        modify_table_body(~ arrange(.x, row_type == "glance_statistic")) |> 
+        modify_table_body(~ filter(.x, {{filters}}))
+}
+
+# list(emad = model_b_emad, 
+#      hl = model_b) |> 
+#     reg_tbl()
+
+write_reg_tbl = function(tbl, path) {
+    tbl |> 
+        as_flex_table() |> 
+        flextable::save_as_docx(path = str_c(path, '.docx'),
+                                pr_section = officer::prop_section(
+                                    page_size = officer::page_size(width = 11, 
+                                                                   orient = 'landscape')
+                                )
+        )
+    ## Problems exporting gt to PDF are a known issue
+    ## <https://github.com/rstudio/gt/issues/721>
+    tbl |> 
+        as_gt() |> 
+        gtsave(filename = str_c(path, '.pdf'), zoom = 1)
+    
+    tbl |> 
+        tbl_butcher() |> 
+        as_gt() |> 
+        write_rds(file = str_c(path, '.Rds'))
+}

### 2022-06-20 scripts/03_dag.R

@@ -201,6 +201,8 @@ part_values_share
 part_values_plot + part_values_share +
     plot_layout(guides = 'collect') +
     plot_annotation(tag_levels = 'A')
+ggsave(here(out_dir, '03_part_values.png'), 
+       height = 4, width = 8, dpi = 200, scale = 1.5)
 
 table(dataf$political_ideology, dataf$pref)
 
@@ -279,8 +281,8 @@ list(emad = model_b_emad,
     trans_plot_emad = ggplot(emad_df, aes(disclosure, pa_mean)) +
         # geom_violin(draw_quantiles = .5) +
         geom_beeswarm(alpha = .25, size = .3) +
-        stat_summary(fun.data = 'mean_cl_boot', color = 'red', 
-                     size = 1.5, fatten = 1/2) +
+        stat_summary(fun.data = mean_cl_boot, color = 'red', 
+                     size = 1, fatten = 0) +
         stat_summary(geom = 'line', group = 1L, color = 'red') +
         labs(y = 'trust')
     trans_plot_emad
@@ -288,18 +290,18 @@ list(emad = model_b_emad,
     
     trans_plot_us = ggplot(dataf, aes(disclosure, meti_mean)) +
         geom_beeswarm(alpha = .25, size = .3) +
-        stat_summary(fun.data = 'mean_cl_boot', color = 'red', 
-                     size = 1.5, fatten = 1/2) +
+        stat_summary(fun.data = mean_cl_boot, color = 'red', 
+                     size = 1, fatten = 0) +
         stat_summary(geom = 'line', group = 1L, color = 'red') +
         labs(y = 'trust')
     trans_plot_us
     
     trans_plot_emad + 
-        ggtitle('Elliott et al. (2017)') +
+        ggtitle('EMAD') +
         trans_plot_us +
-        ggtitle('Hicks and Lobato')
+        ggtitle('HL replication')
     
-    ggsave(here(out_dir, '04_transparency.png'), 
+    ggsave(here(out_dir, '03_transparency.png'), 
            height = 3, width = 6, scale = 1, 
            bg = 'white')
 }
@@ -323,6 +325,15 @@ list(emad = model_c_emad,
      hl = model_c) |> 
     reg_tbl()
 
+## B + C combined table ----
+model_bc_emad = lm(pa_mean ~ conclusion + disclosure, data = emad_df)
[…]

### 2022-07-05 scripts/03_dag.R
create out dir if it doesn't already exist

@@ -31,6 +31,9 @@ options(contrasts = c('contr.Treatment', 'contr.poly'))
 options(decorate.contr.Treatment = '')
 
 out_dir = here('out')
+if (!dir.exists(out_dir)) {
+    dir.create(out_dir)
+}
 data_dir = here('data')
 
 ## Load data ----

### 2022-07-05 paper/paper.Qmd
switch paper paths to here()

@@
   abstract
+execute:
+    echo: false
 ---
+
+```{r}
+#| message: false
+#| results: hide
+
+library(here)
+here()
+out_dir = here('out')
+```
 
@@
 (A) Absolute counts, (B) shares within political ideology categories.
-Panel (A) shows that our sample substantially over-represents political liberals.](../out/03_part_values.png){#fig-part-values}
+Panel (A) shows that our sample substantially over-represents political liberals.](`r here(out_dir, '03_part_values.png')`){#fig-part-values}
 
@@
 ```{r}
-#| echo: false
 #| label: tbl-bc
 #| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
-bc_tbl = readRDS('../out/03_bc_tbl.Rds')
+bc_tbl = readRDS(here(out_dir, '03_bc_tbl.Rds'))
 bc_tbl
@@
 If the DAG is faithful, adjusting for scientist values (`sci_values`) and participant values (`part_values`) is sufficient for estimating the effect of shared values on perceived trustworthiness (`METI`).
-In particular, adjusting for demographics should not change the regression coefficient on shared values.](../out/03_shared_values_dag.png){#fig-shared-dag}
+In particular, adjusting for demographics should not change the regression coefficient on shared values.](`r here(out_dir, '03_shared_values_dag.png')`){#fig-shared-dag}
 
 ```{r}
-#| echo: false
 #| label: tbl-shared
 #| tbl-cap: "Sequential regression analysis of hypothesis H4 shared values/scientist values effects"
-shared_values_tbl = readRDS('../out/03_shared_values_tbl.Rds')
+shared_values_tbl = readRDS(here(out_dir, '03_shared_values_tbl.Rds'))
 shared_values_tbl
@@
 ```{r}
-#| echo: false
 #| label: tbl-eb
 #| tbl-cap: "Regression analysis of H5b, interaction between participant values and consumer risk sensitivity"
-readRDS('../out/03_eb_tbl.Rds')
+readRDS(here(out_dir, '03_eb_tbl.Rds'))
 ```
@@
 ```{r}
-#| echo: false
 #| label: tbl-ec
 #| tbl-cap: "Regression analysis of H5c, interaction between participant values and disclosure effect"
-readRDS('../out/03_ec_tbl.Rds')
+readRDS(here(out_dir, '03_ec_tbl.Rds'))
 ```
@@
 
-![Data, group means, and 95% confidence intervals for H5d, interaction of shared values and participant values.  Panels correspond to participant values.](../out/03_shared_part.png){#fig-shared-part}
+![Data, group means, and 95% confidence intervals for H5d, interaction of shared values and participant values.
+Panels correspond to participant values.](`r here(out_dir, '03_shared_part.png')`){#fig-shared-part}
 
@@
 Panels correspond to participant values.
-This figure is identical to @fig-shared-part, except the columns in the economic growth panel have been switched.](../out/03_sci_part.png){#fig-sci-part}
+This figure is identical to @fig-shared-part, except the columns in the economic growth panel have been switched.](`r here(out_dir, '03_sci_part.png')`){#fig-sci-part}
 
 ```{r}
-#| echo: false
 #| label: tbl-sci-part
-#| caption: "Regression analysis of interaction between scientist values and participant values"
-readRDS('../out/03_sci_part_tbl.Rds')
+#| tbl-cap: "Regression analysis of interaction between scientist values and participant values"
+readRDS(here(out_dir, '03_sci_part_tbl.Rds'))
 ```

### 2022-07-05 paper/paper.Qmd
updated paper
updated the paper to include regression stats in text

--- 2022-06-20 paper/paper.Qmd
@@
 Specifically, results revealed that the conclusions reported by the scientist predicted participants' perceived trustworthiness in line with our hypothesis.
-Participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm.
-By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values are perceived as less trustworthy than a scientist who does not disclose values.
+Participants rated the scientist who reported that BPA does not cause harm as less trustworthy ($$M_{sd}$$ = $$4.48_{1.26}$$) than the scientist who reported that BPA causes harm ($$M_{sd}$$ = $$5.47_{1.12} $$), $\beta $ = -0.99, t(985) = -13.09, p < .001.
+By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values ($$M_{sd}$$ = $$4.94_{1.33}$$) are perceived as less trustworthy than a scientist who does not disclose values  ($$M_{sd}$$ = $$5.05_{1.21}$$), $\beta $ = -0.12, t(985) = -1.46, p = .143.
 
@@
 #| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
-bc_tbl = readRDS('../out/03_bc_tbl.Rds')
+bc_tbl = readRDS('out/03_bc_tbl.Rds')
 bc_tbl
@@
 R^2 = .173, F(3, 840) = 59.95, p < .001 [@tbl-eb model 2].
-As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm.
-However, this effect was qualified by a significant interaction with participants' values.
+As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta $ = -0.75, t(840) = -4.51, p < .001.
+However, this effect was qualified by a significant interaction with participants' values, $\beta $ = -0.41, t(840) = -2.16, p = .031.
 *[is this right?
-Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values.]*
+Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($$M_{sd}$$ = $$5.56_{1.09}$$) than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values ($$M_{sd}$$ = $$4.63_{1.22}$$) for participants who prioritized economic growth and ($$M_{sd}$$ = $$4.40_{1.24}$$) for participants who prioritized public health].]* Further, participants who prioritized public health over economic growth reading and read about a scientist who concluded that BPA does not cause harm rated the scientist as less trustworthy ($$M_{sd}$$ = $$4.40_{1.24}$$) than participants who read about a scientist who concluded that BPA does cause harm, regardless of those participants' values ($$M_{sd}$$ = $$5.39_{1.15}$$) for participants who prioritized economic growth; $$M_{sd}$$ = $$5.56_{1.09}$$ for participants who prioritized public health).
 
@@
 R^2 < .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2].
-As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values.
+As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta $ = -0.23, t(840) = -1.07, p = 0.286.
 
@@
 Reploting the same data as @fig-shared-part based on scientist values, rather than shared values, suggests a more consistent effect across participant values [@fig-sci-part].
-Therefore, to test whether an effect of scientist values might vary as a function of participants values, we regressed participants' METI of the scientist in the stimulus onto the Scientist Values variable, Partiicpants' Values variable, and their interaction [@tbl-sci-part].
+Therefore, to test whether an effect of scientist values might vary as a function of participants values, we regressed participants' METI of the scientist in the stimulus onto the Scientist Values variable, Participants' Values variable, and their interaction [@tbl-sci-part].
 This model was significant, adj.
 R^2 .065, F(3, 563) = 14.15, p < .001.
-But estimates for both the scientist values term and interaction term were highly uncertain and not significant.
+But estimates for both the scientist values term _[EL: shouldn't this be participants values term?]_ and interaction term were highly uncertain and not significant.
 *[finish combining with older text below]*

### 2022-07-06 paper/paper.Qmd
fix branches todo
overlooked merge issue?

@@
 ```{r}
-#| echo: false
 #| label: tbl-bc
 #| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
-bc_tbl = readRDS('out/03_bc_tbl.Rds')
+bc_tbl = readRDS(here(out_dir, '03_bc_tbl.Rds'))
 bc_tbl

### 2022-07-06 paper/paper.Qmd
latex issues

--- 2022-07-05 paper/paper.Qmd
@@
 Specifically, results revealed that the conclusions reported by the scientist predicted participants' perceived trustworthiness in line with our hypothesis.
-Participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm.
-By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values are perceived as less trustworthy than a scientist who does not disclose values.
+Participants rated the scientist who reported that BPA does not cause harm as less trustworthy ($M_{sd} = 4.48_{1.26}$) than the scientist who reported that BPA causes harm ($M_{sd} = 5.47_{1.12}$), $\beta$ = -0.99, t(985) = -13.09, p < .001.
+By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values ($M_{sd} = 4.94_{1.33}$) are perceived as less trustworthy than a scientist who does not disclose values  ($M_{sd} = 5.05_{1.21}$), $\beta$ = -0.12, t(985) = -1.46, p = .143.
 
@@
 R^2 = .173, F(3, 840) = 59.95, p < .001 [@tbl-eb model 2].
-As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA causes harm as more trustworthy than the scientist who reported that BPA does not cause harm.
-However, this effect was qualified by a significant interaction with participants' values.
+As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta$ = -0.75, t(840) = -4.51, p < .001.
+However, this effect was qualified by a significant interaction with participants' values, $\beta$ = -0.41, t(840) = -2.16, p = .031.
 *[is this right?
-Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values.]*
+Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values ($M_{sd} = 4.63_{1.22}$) for participants who prioritized economic growth and ($M_{sd} = 4.40_{1.24}$) for participants who prioritized public health].]* Further, participants who prioritized public health over economic growth reading and read about a scientist who concluded that BPA does not cause harm rated the scientist as less trustworthy ($M_{sd} = 4.40_{1.24}$) than participants who read about a scientist who concluded that BPA does cause harm, regardless of those participants' values ($M_{sd} = 5.39_{1.15}$) for participants who prioritized economic growth; $M_{sd} = 5.56_{1.09}$ for participants who prioritized public health).
 
@@
 R^2 < .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2].
-As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values.
+As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta$ = -0.23, t(840) = -1.07, p = 0.286.
 

### 2022-07-06 paper/paper.Qmd
mostly working on interaction sections

@@
   abstract
+toc: true
 execute:
@@
 
-*[consistent labeling of hypotheses]*
+H1. Modest correlation between values and ideology
+ ~ \(a\) Political liberals are more likely to prioritize public health over economic growth, compared to political conservatives; but (b) a majority of political conservatives prioritize public health.
 
-a. Modest correlation between values and ideology:  
-	(i) Political liberals are more likely to prioritize public health over economic growth, compared to political conservatives; but (ii) a majority of political conservatives prioritize public health.
-b. Consumer risk sensitivity:  Scientists who find that a chemical harms human health are perceived as more trustworthy than scientists who find that a chemical does not cause harm. 
-c. Transparency penalty: Scientists who disclose values are perceived as less trustworthy than scientists who do not.  
-d.
-Shared values:  Given that the scientist discloses values, if the participant and the scientist share the same values, the scientist is perceived as more trustworthy than if the participant and scientist have discordant values.
-e. Variation in effects:  The magnitude of the effects above vary depending on whether the participant prioritizes public health or economic growth.
+H2. Consumer risk sensitivity
+ ~ Scientists who find that a chemical harms human health are perceived as more trustworthy than scientists who find that a chemical does not cause harm.
 
-Hypothesis c, the transparency penalty, corresponds to the objection to transparency:  disclosing values undermines trust in science.
+H3. Transparency penalty
+ ~ Scientists who disclose values are perceived as less trustworthy than scientists who do not.
+
+H4. Shared values
+ ~ Given that the scientist discloses values, if the participant and the scientist share the same values, the scientist is perceived as more trustworthy than if the participant and scientist have discordant values.
+
+H5. Variation in effects
+ ~ The magnitude of the effects 2-4 vary depending on whether the participant prioritizes public health or economic growth.
+
+Hypothesis H3, the transparency penalty, corresponds to the objection to transparency:  disclosing values undermines trust in science.
 However, the shared values effect works in the opposite direction, counteracting the transparency penalty.
 
-*[hypothesis a]*
-
-Hypotheses b-e were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed a priori to identify appropriate adjustments (covariates) for hypotheses d and e.
+Hypothesis 1 was analyzed using Spearman rank order correlation and visually.
+Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed a priori to identify appropriate adjustments (covariates) for H4 and 5.
+Because @ElliottValuesEnvironmentalResearch2017 made their data publicly available, and our analytical approaches differ somewhat from theirs, we conducted parallel analyses of their data, and report these parallel results for H2 and 3.
+Exploratory data analysis was used throughout to support data validation and aid interpretation.
 
@@
 
+## Software and reproducibility ##
+
+Data cleaning and analysis was conducted in R *[version, cite]*, with extensive use of the `tidyverse` suite of packages *[cite]*.
+Regression tables were generated using the `gt` and `gtsummary` packages *[cites]*.
+
+*[All original data and reproducible code is available at <https://github.com/dhicks/transparency>.  Instructions in that repository explain how to automatically reproduce our analysis.]*
+
 # Results #
@@
 660 participants (67%) were randomly assigned to a values disclosure condition (scientist prioritizes either economic growth or public health); 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).
-Consequently, subsamples for hypotheses d and e were substantially smaller than the full analysis sample.
+Consequently, subsamples for hypotheses 4 and 5 were substantially smaller than the full analysis sample.
 
@@
 
+## H1: Correlation between values and ideology
+
 We tested the hypothesis of a modest correlation between values and ideology in two ways.
@@
 
+## H2 and H3: Consumer risk sensitivity and transparency penalty
+
 Next, we tested the hypotheses that (H2) scientists who find a chemical harms human health are perceived as more trustworthy than scientist who find that a chemical does not cause harm and (H3) a scientist who discloses values are perceived as less trustworthy than a scientist who does not.
@@
 The full model was significant, adj.
-R^2 = .148, F(2, 985) = 86.59, p < .001 (see @tbl-bc).
+$R^2$= .148, F(2, 985) = 86.59, p < .001 (see @tbl-bc).
 Specifically, results revealed that the conclusions reported by the scientist predicted participants' perceived trustworthiness in line with our hypothesis.
@@
 
+## H4: Shared values
+
 Next, we tested the hypothesis that (H4) if the participant and scientist share the same values, the scientist is perceived as more trustworthy than if the participant and scientist do not share the same values.
@@
 While the univariate model was significant, adj.
-R^2 = .029, F(1, 565) = 18.0, p < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.
+$R^2$= .029, F(1, 565) = 18.0, p < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.
 Rather, only the scientist's disclosed values significantly predicted how trustworthy participants rated the scientist, such that a scientist who disclosed valuing public health was rated as more trustworthy than a scientist who disclosed valuing economic growth.
@@
 
-## Variation in Effect Sizes ##
+## H5: Variation in effects
 
@@
 
-H5b - consumer risk sensitivity
+### Consumer risk sensitivity
+
 To test whether the findings regarding consumer risk sensitivity vary as a function of participants' values, we regressed participants' METI ratings of the scientist in the stimuli onto the Conclusions condition, Participants' Values variable, and the Conclusions by Participants' values interaction term.
 The full model was significant, adj.
-R^2 = .173, F(3, 840) = 59.95, p < .001 [@tbl-eb model 2].
+$R^2$ = .173, F(3, 840) = 59.95, p < .001 ([@tbl-eb model 2]).
 As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta$ = -0.75, t(840) = -4.51, p < .001.
 However, this effect was qualified by a significant interaction with participants' values, $\beta$ = -0.41, t(840) = -2.16, p = .031.
-*[is this right?
-Participants who prioritized public health over economic growth and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants who read about a scientist who concluded that BPA does not cause harm, regardless of those participants' values ($M_{sd} = 4.63_{1.22}$) for participants who prioritized economic growth and ($M_{sd} = 4.40_{1.24}$) for participants who prioritized public health].]* Further, participants who prioritized public health over economic growth reading and read about a scientist who concluded that BPA does not cause harm rated the scientist as less trustworthy ($M_{sd} = 4.40_{1.24}$) than participants who read about a scientist who concluded that BPA does cause harm, regardless of those participants' values ($M_{sd} = 5.39_{1.15}$) for participants who prioritized economic growth; $M_{sd} = 5.56_{1.09}$ for participants who prioritized public health).
+Participants who prioritized public health and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants with the same values who read about a scientist who concluded that BPA does not cause harm ($M_{sd} = 4.40_{1.24}$) ([@fig-conclusion-part]).
 
@@
 #| label: tbl-eb
-#| tbl-cap: "Regression analysis of H5b, interaction between participant values and consumer risk sensitivity"
+#| tbl-cap: "Regression analysis of H5-consumer, interaction between participant values and consumer risk sensitivity"
 readRDS(here(out_dir, '03_eb_tbl.Rds'))
@@
 
-H5c - transparency penalty
+![Data, group means, and 95% confidence intervals for H5-consumer, interaction of consumer risk sensitivity and participant values.
+Panels correspond to participant values.](`r here(out_dir, '03_conclusion_part.png')`){#fig-conclusion-part}
+
+### Transparency penalty
+
 To test whether the findings above about a hypothesized transparency penalty may vary based on participants' values, we regressed participants' METI ratings of the scientist onto the Disclosure condition variable, Participants' Values variable, and the Disclosure by Participants' Values interaction term.
 The full model was not significant, adj.
-R^2 < .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2].
+$R^2$< .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2].
 As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta$ = -0.23, t(840) = -1.07, p = 0.286.
@@
 #| label: tbl-ec
-#| tbl-cap: "Regression analysis of H5c, interaction between participant values and disclosure effect"
+#| tbl-cap: "Regression analysis of H5-transparency, interaction between participant values and transparency penalty"
 readRDS(here(out_dir, '03_ec_tbl.Rds'))
@@
 
-H5d - shared values
+### Shared values and scientist values
+
 Without adjustments, Shared Values appears to have a substantial interaction with Participant Values:  Shared Values appears to increase perceived trustworthiness for participants who value public health, while decreasing perceived trustworthiness for participants who value economic growth [@fig-shared-part].
 However, as indicated by our analysis for a potential shared values effect, the estimated effects for both shared values and participant values are biased if the model is not adjusted for scientist values.
-(@tbl-shared shows that the Shared Values estimate is biased.
-Shared Values is a collider between Participant Values and Scientist Values; hence, if a model specification includes Participant Values and Shared values but not Scientist Values, there is an open path between Participant Values and METI, resulting in a biased estimate for Participant Values.)  Because Shared Values is the logical biconditional of Participant Values and Scientist Values, and the Shared Values $\times$ Participant Values is the logical conjunction, including all four variables (Participant, Scientist, and Shared Values, along with the interaction term) in a regression model creates perfect collinearity.
+(@fig-shared-dag and @tbl-shared show how this bias occurs.
+Shared Values is a collider between Participant Values and Scientist Values; hence, if a model specification includes Participant Values and Shared values but not Scientist Values, there is an open path between Participant Values and METI, resulting in a biased estimate for Participant Values.)  Because Shared Values is the logical biconditional of Participant Values and Scientist Values, and the Shared Values $\times$ Participant Values interaction term is their logical conjunction, including all four variables (Participant, Scientist, and Shared Values, along with the interaction term) in a regression model creates perfect collinearity.
 
@@
 As we found above, rather than a shared values effect there appears to be an effect of scientist values.
-Reploting the same data as @fig-shared-part based on scientist values, rather than shared values, suggests a more consistent effect across participant values [@fig-sci-part].
-Therefore, to test whether an effect of scientist values might vary as a function of participants values, we regressed participants' METI of the scientist in the stimulus onto the Scientist Values variable, Participants' Values variable, and their interaction [@tbl-sci-part].
+Replotting the same data as @fig-shared-part based on scientist values, rather than shared values, suggests a more consistent effect across participant values (@fig-sci-part).
+Therefore, we conducted an unplanned post hoc analysis to test whether an effect of scientist values might vary as a function of participants values.
+We regressed participants' METI of the scientist in the stimulus onto the Scientist Values variable, Participants' Values variable, and their interaction (@tbl-sci-part).
 This model was significant, adj.
-R^2 .065, F(3, 563) = 14.15, p < .001.
-But estimates for both the scientist values term _[EL: shouldn't this be participants values term?]_ and interaction term were highly uncertain and not significant.
-*[finish combining with older text below]*
+$R^2$ = 0.065, F(3, 563) = 14.15, p < .001.
+The estimate for the scientist values term was significant, $\beta$ = 0.53, t(563) = 2.26, p = 0.02, while the estimate for the interaction term was not, $\beta$ = 0.20, t(563) = 0.75, p = 0.45.
+Estimates for both variables had large uncertainties, with confidence intervals of about 1 (95% confidence intervals, scientist values: 0.07, 1.0; interaction term: -0.32, 0.72).
 
@@
 
-To test whether a potential shared values effect might vary as a function of participants' values, we regressed participants' METI ratings of the scientist in the stimuli onto the Shared Values variable, Participants' Values variable, and the Shared Values by Participants' values interaction term.
-The full model was significant, adj.
-R^2 = .065, F(3, 563) = 14.15, p < .001 *[tbl]*.
-The model revealed a significant interaction between Shared Values and Participants' Values, such that overall Shared Values had a positive effect when the participant valued public health, but a negative effect when the participant valued economic growth *[@fig-shared-part and tbl]*.
-
-such that the highest perceived trustworthiness ratings were from participants who prioritized public health reading about a scientist who disclosed valuing public health.
-However, as revealed in our analysis for H4, the evidence that shared values between the participant and scientist are what predicts the perceived trustworthiness of a scientist should be interpreted with caution.
-Because of the inclusion of Participants' values and the Scientist's values in the analysis for H4, the set of predictor variables for that model and the set of predictor variables for this model are different linear combinations of each other.
-Hence why the resulting full model parameters are identical between the two models.
-In spite of this, we report the results of this analysis for the sake of completeness, as this was an a priori planned statistical analysis.
-
-Because of the findings from the analysis run to test H4, we conducted an unplanned post hoc analysis to further test the hypothesis that the effects found in the earlier analysis vary as a function of participants' values.
-For this test, we regression METI scores onto the Scientist's Values experimental condition, the Shared Values composite variable, and their interaction term.
-This is, as with the previous model and the model reported to test H4, a different linear combination of the same predictor variables and therefore the full model parameters are identical.
-However, presentation of the model this way renders a clearer picture of what variables we tested seem to be driving perceptions of the trustworthiness of a scientist.
-The evidence here suggests that if a scientist chooses to disclose their values alongside their presentation of scientific research, what values they share predicts their perceived trustworthiness, and that the scientist in our study who disclosed valuing public health was perceived a significantly more trustworthy than the scientist who disclosed valuing economic growth.
+*[maybe move to conclusion]*
+The evidence here suggests that if a scientist discloses their values alongside their presentation of scientific research, what values they share predicts their perceived trustworthiness, and that a scientist who discloses valuing public health will tend to be perceived as more trustworthy than the scientist who discloses valuing economic growth.
 

### 2022-07-06 scripts/03_dag.R

@@ -541,11 +541,25 @@ write_reg_tbl(eb_tbl, here(out_dir, '03_eb_tbl'))
 
 
 emad_df |> 
-    filter(!is.na(part_values)) |> 
+    filter(!is.na(part_values), disclosure) |> 
     ggplot(aes(conclusion, pa_mean)) +
     geom_boxplot() +
     facet_wrap(vars(part_values))
 
+dataf |> 
+    filter(!is.na(part_values), disclosure) |> 
+    ggplot(aes(conclusion, meti_mean)) +
+    geom_beeswarm(alpha = .5, cex = 1.5, size = .5) +
+    stat_summary(fun.data = mean_cl_boot, color = 'red', 
+                 size = 1, fatten = 0) +
+    stat_summary(fun.data = mean_cl_boot, geom = 'line', group = 1L, color = 'red') +
+    facet_wrap(vars(part_values)) +
+    labs(x = 'scientist conclusion: BPA ...', 
+         y = 'perceived trustworthiness')
+
+ggsave(here(out_dir, '03_conclusion_part.png'), 
+       height = 3, width = 6, scale = 1, 
+       bg = 'white')
 
 #' ## C: Disclosure ##
 dag |> 

### 2022-07-06 scripts/01_clean.R
automatically download emad data file

@@ -227,7 +227,20 @@ emad_xwalk = tribble(
 )
 
 ## Elliott, McCright, Allen, and Dietz
-emad = read_csv(here('data', 'elliot_et_al', 'Values and Science Experiment 1.csv'))
+## <https://figshare.com/articles/dataset/Elliott_McCright_Allen_and_Dietz_Data_SPSS_Stata_/4695793>
+emad_dir = here('data', 'elliot_et_al')
+if (!dir.exists('data')) {
+    dir.create('data')
+}
+if (!dir.exists(emad_dir)) {
+    dir.create(emad_dir)
+}
+emad_file = here(emad_dir, 'Values and Science Experiment 1.csv')
+if (!file.exists(emad_dir)) {
+    download.file('https://figshare.com/ndownloader/files/8371322', 
+                  emad_file)
+}
+emad = read_csv(emad_file)
 
 emad_clean = emad |> 
     ## Condition should be 1 variable, not 6

### 2022-07-08 paper/paper.Qmd
reproducibility; include emad data in repo
fill in references; working on conclusion

@@
     echo: false
+bibliography: |
+  transparency-Hicks.yaml
 ---
@@
 
-Over the past 15 years, many philosophers of science have rejected the ideal of value-free science and related, traditional understandings of objectivity and political neutrality of science.
-*[Douglas and Elliott]*  According to the value-free ideal, social and political values — such as feminism, environmentalism, or the protection of human health — have no legitimate role to play in the evaluation of scientific hypotheses.
+Over the past 15 years, many philosophers of science have rejected the ideal of value-free science and related, traditional understandings of objectivity and political neutrality of science [@DouglasSciencePolicyValuefree2009; @ElliottTapestryValuesIntroduction2017].
+According to the value-free ideal, social and political values — such as feminism, environmentalism, or the protection of human health — have no legitimate role to play in the evaluation of scientific hypotheses.
 The value-free ideal is compatible with allowing social and political values to play important roles earlier and later in inquiry.
@@
 Keeping with the example, the question of whether chemical X causes cancer has much more significant social and political implications than the question of whether chemical X fluoresces under ultraviolet light.
-So, in the interest of protecting human health, it would be appropriate to *scientifically* accept the hypothesis that chemical X is carcinogenic based on relatively provisional evidence *[cite]*.
+So, in the interest of protecting human health, it would be appropriate to *scientifically* accept the hypothesis that chemical X is carcinogenic based on relatively provisional evidence [@CranorSocialBenefitsExpedited1995; @ElliottNonepistemicValuesMultiple2014; @FernandezPintoLegitimizingValuesRegulatory2019].
 
 While this kind of argument indicates that at least some social and political values may or should play at least some role in the evaluation of at least some scientific hypotheses, it doesn't provide us with much positive guidance:  which values, playing which roles, in the evaluation of which hypotheses?
-*[New Direction]*  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research *[Elliott]*.
+[@HicksNewDirectionScience2014]  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research [@ElliottSciencePolicyTransparency2014; @McKaughanBacktrackingEthicsFraming2013].
 But some philosophers have objected to this transparency proposal, arguing that it might actually undermine trust in science.
-According to this objection, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy *[John, Kovaka]*.
+According to this objection, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021].
 
@@
 This first bullet is not present in the "no disclosure" values condition.
-The second bullet, identical across conditions, states "I examined the scientific evidence on potential health risks of BPA."  The third bullet concludes that BPA either does or does not cause harm to people, depending on whether the subject is in the harm or *[not-harm]* condition.
+The second bullet, identical across conditions, states "I examined the scientific evidence on potential health risks of BPA."  The third bullet concludes that BPA either does or does not cause harm to people, depending on whether the subject is in the "causes harm" or "does not cause harm" condition.
 
@@
 Spence's trustworthiness using a semantic differential scale.
-@ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @ElliottValuesEnvironmentalResearch2017], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
+@ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
 On the METI, participants rate the target scientist (the fictional Dr.
@@
 In both @ElliottValuesEnvironmentalResearch2017 and our replication analysis, all items were averaged together to form a 1-7 composite measure of trustworthiness.
-To aid interpretation, in the analysis the direction of the scale has been set so that increased values correspond to increasing perceived trustworthiness.
+To aid interpretation, in the analysis the direction of the scale has been set so that increased values correspond to increased perceived trustworthiness.
 
-Ethicists make a key distinction between trust and trustworthiness *[Baier]*.
+Ethicists make a key distinction between trust and trustworthiness [@BaierTrustAntitrust1986].
 Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.
@@
 For our replication, we chose to keep BPA, rather than switching to a different public scientific controversy with a higher profile in 2021, such as climate change, police violence, voter fraud, or any of numerous aspects of the Covid-19 pandemic.
-All of these controversies are highly politically charged, with prominent public experts and counterexperts *[cite]*.
+All of these controversies are highly politically charged, with prominent public experts and counterexperts [@GoldenbergVaccineHesitancyPublic2021 100-1, ch.
+6].
 While we anticipated some effects of political partisanship in the BPA case, we felt it would be less likely to swamp the values disclosure that was our primary interest.
@@
 A recent analysis finds that Prolific produces substantially higher quality data than Amazon Mechanical Turk for online survey studies, though three of the five authors are affiliated with Prolific [@PeerDataQualityPlatforms2021].
-Preliminary power analysis recommended a sample of approximately 1,000 participants to reliably detect non-interaction effects (hypotheses a-d).
+Preliminary power analysis recommended a sample of approximately 1,000 participants to reliably detect non-interaction effects (H1-4).
 
@@
 
-*[All original data and reproducible code is available at <https://github.com/dhicks/transparency>.  Instructions in that repository explain how to automatically reproduce our analysis.]*
+*[Anonymized original data and reproducible code are available at <https://github.com/dhicks/transparency>.  Instructions in that repository explain how to automatically reproduce our analysis.]*
 
@@
 2]*  For political ideology, the General Social Survey has consistently found over several decades that about 30% of US adults identify as liberal, about 30% identify as conservative, and about 40% as moderate *[https://gssdataexplorer.norc.org/variables/178/vshow]*.
-Among our participants, liberals outnumber conservatives *[by over 2:1; hypothesis a fig]*.
-Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling *[https://academic.oup.com/poq/article/82/1/1/4837043?login=false]*.
+Among our participants, liberals outnumber conservatives *[by over 2:1]* (@fig-part-values).
+Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling [@KennedyEvaluation2016Election2018].
 In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement]*.
-This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  General Social Survey data shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public controversies align with liberal-conservative partisanship *[cite]*.
-Insofar as some political conservatives are both less likely to participate in studies on Prolific (or at least in our particular study) and likely to perceive Spence as less trustworthy, this will lead to omitted variable bias for analyses that require adjustment by political ideology.
-*[however, most of our analyses do not require this adjustment]*
+This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  data from the General Social Survey shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public controversies align with liberal-conservative partisanship *[cite]*.
+
+Because of these representation issues, insofar as some political conservatives are both less likely to participate in studies on Prolific (or at least in our particular study) and likely to perceive Spence as less trustworthy, this will produce omitted variable bias for analyses that require adjustment by political ideology.
+Analysis of our a priori DAG indicated that, for the hypotheses examined here, this adjustment was not necessary in any case.
+When some adjustment was necessary (H4-5), we included political ideology along with other demographic variables in an alternative model specification as a robustness check.
+In line with the DAG, in no case did these robustness checks produce indications of omitted variable bias.
 
@@
 We ran the following analyses to test our hypothesis that (H5) the magnitude of the effects found for the tests of H2-H4 vary depending on whether the participant prioritizes public health or economic growth.
-Results are presented in Table #.
+Results are presented in [@tbl-eb; @tbl-ec; @tbl-sci-part].
 
@@
 # Conclusion #
+
+@tbl-summary summarizes the results of our replication attempts.
+We were able to successfully replicate the correlation between political ideology and valuing public health and consumer risk sensitivity (a scientist who concludes BPA is harmful is perceived as more trustworthy), and in addition found evidence of an interaction between consumer risk sensitivity and whether the participant values public health (participants who value public health have more extreme reactions to the scientist's conclusion).
+
+We did not find evidence for a transparency penalty or shared values effect.  And we found evidence of a scientist values effect, which also did not appear in @ElliottValuesEnvironmentalResearch2017.
+
++------------------+------------------------------------------+-------------+
+|                  | Hypothesis                               | Replicated? |
++==================+==========================================+=============+
+|              H1a | Liberals prioritize public health        | yes         |
+|                  | rel. to conservatives                    |             |
++------------------+------------------------------------------+-------------+
+|              H1b | Majority of conservatives prioritize     | no          |
+|                  | public health                            |             |
++------------------+------------------------------------------+-------------+
+|               H2 | Sci. who find chemical is harmful        | yes         |
+|                  | perceived as more trustworthy            |             |
++------------------+------------------------------------------+-------------+
+|               H3 | Sci. who disclose values perceived as    | no          |
+|                  | less trustworthy                         |             |
++------------------+------------------------------------------+-------------+
+|               H4 | Given sci. discloses values,             | no          |
+|                  | sci. who shared values with participant  |             |
+|                  | perceived as more trustworthy            |             |
++------------------+------------------------------------------+-------------+
+|        Unplanned | Given sci. discloses values,             |yes$^\dagger$|
+|                  | sci. who values public health            |             |
+|                  | perceived as more trustworthy            |             |
++------------------+------------------------------------------+-------------+
+|      H5-consumer | Interaction btwn. sci. conclusion (H2)   | yes         |
+|                  | and participant values                   |             |
++------------------+------------------------------------------+-------------+
+|  H5-transparency | Interaction btwn. values disclosure (H3) | no          |
+|                  | and participant values                   |             |
++------------------+------------------------------------------+-------------+
+|        H5-shared | Interaction btwn. shared values (H4)     | $^\star$    |
+|                  | and participant values                   |             |
++------------------+------------------------------------------+-------------+
+|   H5-sci. values | Interaction btwn. scientist values       | no$^\dagger$|
+|                  | and participant values                   |             |
++------------------+------------------------------------------+-------------+
+
+: Summary of replication results.
+$^\star$: The shared values $\times$ participant values interaction specification had perfect multicollinearity and could not be fit.
+$^\dagger$: For these unplanned analyses, "yes" indicates that we found evidence supporting the stated hypothesis, and "no" indicates that we did not find such evidence.
+{#tbl-summary}
+
+@ElliottValuesEnvironmentalResearch2017 claim evidence for a shared values effect by comparing values disclosure vs.
+no disclosure (that is, estimating a transparency effect) in quasi-independent subsamples across 8 cells (participant values $\times$ scientist conclusion $\times$ scientist values; @ElliottValuesEnvironmentalResearch2017 fig.
+1).
+This approach is difficult to interpret; their reasoning seems to be that this estimate is statistically significant in 4/8 cells, and in 3/4 of these cells (1 where respondent values economic growth) the participant and scientist share values.
+But these analyses are underpowered (for example, sample sizes are well below 100 for some of the cells in which participants value economic growth) and the overall approach cannot distinguish a shared values effect from other potential effects (scientist values, participant values).
+Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$=0.29, t(335)=1.75, p=0.08; while the estimate for scientist values was statistically significant, $\beta$=0.41, t(335)=2.48, p=0.014.
+It seems likely that claims of a shared values effect were the result of misinterpreting a confusing statistical approach.
+*[trying to put this as nice as possible]*
+
+The disagreement over a transparency penalty is more difficult to explain.
+Using our approach and the data from @ElliottValuesEnvironmentalResearch2017, the estimate for a values disclosure was statistically significant (@tbl-bc model 1).
+One possible explanation is that, over the course of the Covid-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.
+(And, more recently, Leana Wen and some other prominent physicians and public health officials have argued for the importance of a "return to normal," which might be understood as prioritizing economic growth.)  A values disclosure that was regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.
+
+*[further work]*
+
+# References #

### 2022-07-11 paper/paper.Qmd
working on wide table problem for PDF output

@@
   transparency-Hicks.yaml
+header-includes:
+  - \usepackage{pdflscape}
+  - \newcommand{\blandscape}{\landscape}
+  - \newcommand{\elandscape}{\endlandscape}
+  # - \usepackage{endfloat}
 ---
@@
 
+library(gtsummary)
 library(here)
@@
 ```
+
+```{r}
+#| label: tbl-test
+#| tbl-cap: "Testing"
+knitr::include_graphics('../out/03_eb_tbl.pdf')
+```
+
+@tbl-test
+
+<!-- \blandscape -->
+```{r}
+#| tbl-cap: "Sequential regression analysis of hypothesis H4 shared values/scientist values effects"
+readRDS(here(out_dir, '03_shared_values_tbl.Rds')) |> 
+    as_hux_table() |> 
+    huxtable::set_width(1)
+```
+<!-- \elandscape -->
 
@@
 #| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
-bc_tbl = readRDS(here(out_dir, '03_bc_tbl.Rds'))
-bc_tbl
+readRDS(here(out_dir, '03_bc_tbl.Rds')) |> 
+    as_hux_table()
 ```
@@
 
+<!-- \blandscape -->
+
 ```{r}
@@
 #| tbl-cap: "Sequential regression analysis of hypothesis H4 shared values/scientist values effects"
-shared_values_tbl = readRDS(here(out_dir, '03_shared_values_tbl.Rds'))
-shared_values_tbl
+readRDS(here(out_dir, '03_shared_values_tbl.Rds')) |> 
+    as_hux_table()
 ```
+<!-- \elandscape -->
 
@@
 #| tbl-cap: "Regression analysis of H5-consumer, interaction between participant values and consumer risk sensitivity"
-readRDS(here(out_dir, '03_eb_tbl.Rds'))
+readRDS(here(out_dir, '03_eb_tbl.Rds')) |> 
+    as_hux_table()
 ```
@@
 #| tbl-cap: "Regression analysis of H5-transparency, interaction between participant values and transparency penalty"
-readRDS(here(out_dir, '03_ec_tbl.Rds'))
+readRDS(here(out_dir, '03_ec_tbl.Rds')) |> 
+    as_hux_table()
 ```
@@
 #| tbl-cap: "Regression analysis of interaction between scientist values and participant values"
-readRDS(here(out_dir, '03_sci_part_tbl.Rds'))
+readRDS(here(out_dir, '03_sci_part_tbl.Rds')) |> 
+    as_hux_table()
 ```

### 2022-07-11 R/reg_tbl.R

@@ -35,6 +35,6 @@ write_reg_tbl = function(tbl, path) {
     
     tbl |> 
         tbl_butcher() |> 
-        as_gt() |> 
+        # as_gt() |> 
         write_rds(file = str_c(path, '.Rds'))
 }

### 2022-07-12 paper/paper.Qmd
wide table problem solved

--- 2022-07-08 paper/paper.Qmd
@@
 
+library(gtsummary)
+library(kableExtra)
 library(here)
@@
 out_dir = here('out')
+
+print_tbl = function(x) {
+    x |> 
+        as_kable_extra(booktabs = TRUE) |> 
+        kable_styling(latex_options = c('scale_down'))
+}
 ```
@@
 #| tbl-cap: "Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3 (disclosure effect)"
-bc_tbl = readRDS(here(out_dir, '03_bc_tbl.Rds'))
-bc_tbl
+readRDS(here(out_dir, '03_bc_tbl.Rds')) |> 
+    print_tbl()
 ```
@@
 #| tbl-cap: "Sequential regression analysis of hypothesis H4 shared values/scientist values effects"
-shared_values_tbl = readRDS(here(out_dir, '03_shared_values_tbl.Rds'))
-shared_values_tbl
+readRDS(here(out_dir, '03_shared_values_tbl.Rds')) |> 
+    print_tbl()
 ```
@@
 #| tbl-cap: "Regression analysis of H5-consumer, interaction between participant values and consumer risk sensitivity"
-readRDS(here(out_dir, '03_eb_tbl.Rds'))
+readRDS(here(out_dir, '03_eb_tbl.Rds')) |> 
+    print_tbl()
 ```
@@
 #| tbl-cap: "Regression analysis of H5-transparency, interaction between participant values and transparency penalty"
-readRDS(here(out_dir, '03_ec_tbl.Rds'))
+readRDS(here(out_dir, '03_ec_tbl.Rds')) |> 
+    print_tbl()
 ```
@@
 #| tbl-cap: "Regression analysis of interaction between scientist values and participant values"
-readRDS(here(out_dir, '03_sci_part_tbl.Rds'))
+readRDS(here(out_dir, '03_sci_part_tbl.Rds')) |> 
+    print_tbl()
 ```

### 2022-07-12 paper/paper.Qmd
both html and pdf formats available

@@
   transparency-Hicks.yaml
-header-includes:
-  - \usepackage{pdflscape}
-  - \newcommand{\blandscape}{\landscape}
-  - \newcommand{\elandscape}{\endlandscape}
-  # - \usepackage{endfloat}
+format: 
+    html: default
+    pdf: default
 ---
@@
 
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |                   | Hypothesis                               | Replicated? |
-+==================+==========================================+=============+
++===================+==========================================+=============+
 |               H1a | Liberals prioritize public health        | yes         |
 |                   | rel. to conservatives                    |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |               H1b | Majority of conservatives prioritize     | no          |
 |                   | public health                            |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |                H2 | Sci. who find chemical is harmful        | yes         |
 |                   | perceived as more trustworthy            |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |                H3 | Sci. who disclose values perceived as    | no          |
 |                   | less trustworthy                         |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |                H4 | Given sci. discloses values,             | no          |
@@
 |                   | perceived as more trustworthy            |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |         Unplanned | Given sci. discloses values,             |yes$^\dagger$|
@@
 |                   | perceived as more trustworthy            |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |       H5-consumer | Interaction btwn. sci. conclusion (H2)   | yes         |
 |                   | and participant values                   |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |   H5-transparency | Interaction btwn. values disclosure (H3) | no          |
 |                   | and participant values                   |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |         H5-shared | Interaction btwn. shared values (H4)     | $^\star$    |
 |                   | and participant values                   |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 |    H5-sci. values | Interaction btwn. scientist values       | no$^\dagger$|
 |                   | and participant values                   |             |
-+------------------+------------------------------------------+-------------+
++-------------------+------------------------------------------+-------------+
 

### 2022-07-13 paper/paper.Qmd
simplify setup chunk options
citations, readme

@@
 ---
-title: "Something something values disclosure replication"
+title: "Values disclosures and trust in science: A replication study"
 author: "Daniel J. Hicks and Emilio Lobato"
 abstract: |
-  abstract
+  While philosophers of science generally agree that social, political, and ethical values can play legitimate roles in science, there is active debate over whether scientists should disclosure such values in their public communications.
+This debate depends, in part, on empirical claims about whether values disclosures might undermine public trust in science.
+In a previous study, Elliott et al.
+used an online experiment to test this empirical claim.
+The current paper reports a replication attempt of their experiment.
+Comparing results of the two replications, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
 toc: true
@@
 ```{r}
-#| message: false
-#| results: hide
+#| include: false
 
@@
 Participants were recruited using the online survey platform Prolific, and the survey was administered in a web browser using Qualtrics.
-Prolific has an option to draw samples that are balanced to be representative by age, binary gender, and a 5-category race variable (taking values Asian, Black, Mixed, Other, and White) for US adults *[https://researcher-help.prolific.co/hc/en-gb/articles/360019238413-Representative-samples-FAQ]*.
+Prolific has an option to draw samples that are balanced to be representative by age, binary gender, and a 5-category race variable (taking values Asian, Black, Mixed, Other, and White) for US adults [@RepresentativeSamplesFAQ2022].
 A recent analysis finds that Prolific produces substantially higher quality data than Amazon Mechanical Turk for online survey studies, though three of the five authors are affiliated with Prolific [@PeerDataQualityPlatforms2021].
@@
 
-Data cleaning and analysis was conducted in R *[version, cite]*, with extensive use of the `tidyverse` suite of packages *[cite]*.
-Regression tables were generated using the `gt` and `gtsummary` packages *[cites]*.
+Data cleaning and analysis was conducted in R version 4.1.2 [@RCoreTeamLanguageEnvironmentStatistical2021], with extensive use of the `tidyverse` suite of packages version 1.3.1 [@WickhamWelcomeTidyverse2019].
+Regression tables were generated using the packages `gt` version 0.5.0 [@IannoneGtEasilyCreate2022] and `gtsummary` version 1.6.0 [@SjobergReproducibleSummaryTables2021].
 
@@
 Critically, our data are unlikely to be representative by education level and political ideology.
-In 2021, about 9% of US adults 25 or over had a less than high school education, and 38% had a Bachelor's degree or higher.
+In 2021, about 9% of US adults 25 or over had a less than high school education, and 38% had a Bachelor's degree or higher [@CPSHistoricalTime2022 fig.
+2].
 Only 1% of our participants reported a less than high school education, and 57% reported a Bachelor's degree or higher.
-*[<https://www.census.gov/library/visualizations/time-series/demo/cps-historical-time-series.html> fig.
-2]*  For political ideology, the General Social Survey has consistently found over several decades that about 30% of US adults identify as liberal, about 30% identify as conservative, and about 40% as moderate *[https://gssdataexplorer.norc.org/variables/178/vshow]*.
-Among our participants, liberals outnumber conservatives *[by over 2:1]* (@fig-part-values).
+For political ideology, the General Social Survey has consistently found over several decades that about 30% of US adults identify as liberal, about 30% identify as conservative, and about 40% as moderate [@GSSDataExplorer2022].
+Among our participants, liberals (574) heavily outnumber conservatives (248; @fig-part-values).
 Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling [@KennedyEvaluation2016Election2018].
 In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement]*.
-This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  data from the General Social Survey shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public controversies align with liberal-conservative partisanship *[cite]*.
+This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  data from the General Social Survey shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public scientific controversies align with liberal-conservative partisanship [@FunkAmericansPoliticsScience2015].
 
@@
 
-+-------------------+------------------------------------------+-------------+
-|                   | Hypothesis                               | Replicated? |
-+===================+==========================================+=============+
-|               H1a | Liberals prioritize public health        | yes         |
-|                   | rel. to conservatives                    |             |
-+-------------------+------------------------------------------+-------------+
-|               H1b | Majority of conservatives prioritize     | no          |
-|                   | public health                            |             |
-+-------------------+------------------------------------------+-------------+
-|                H2 | Sci. who find chemical is harmful        | yes         |
-|                   | perceived as more trustworthy            |             |
-+-------------------+------------------------------------------+-------------+
-|                H3 | Sci. who disclose values perceived as    | no          |
-|                   | less trustworthy                         |             |
-+-------------------+------------------------------------------+-------------+
-|                H4 | Given sci. discloses values,             | no          |
-|                   | sci. who shared values with participant  |             |
-|                   | perceived as more trustworthy            |             |
-+-------------------+------------------------------------------+-------------+
-|         Unplanned | Given sci. discloses values,             |yes$^\dagger$|
-|                   | sci. who values public health            |             |
-|                   | perceived as more trustworthy            |             |
-+-------------------+------------------------------------------+-------------+
-|       H5-consumer | Interaction btwn. sci. conclusion (H2)   | yes         |
-|                   | and participant values                   |             |
-+-------------------+------------------------------------------+-------------+
-|   H5-transparency | Interaction btwn. values disclosure (H3) | no          |
-|                   | and participant values                   |             |
-+-------------------+------------------------------------------+-------------+
-|         H5-shared | Interaction btwn. shared values (H4)     | $^\star$    |
-|                   | and participant values                   |             |
-+-------------------+------------------------------------------+-------------+
-|    H5-sci. values | Interaction btwn. scientist values       | no$^\dagger$|
-|                   | and participant values                   |             |
-+-------------------+------------------------------------------+-------------+
-
-: Summary of replication results.
-$^\star$: The shared values $\times$ participant values interaction specification had perfect multicollinearity and could not be fit.
-$^\dagger$: For these unplanned analyses, "yes" indicates that we found evidence supporting the stated hypothesis, and "no" indicates that we did not find such evidence.
-{#tbl-summary}
+{{< include summary_table.md >}}
 

### 2022-07-13 readme.Rmd
citations, readme

---
title: "Data and code for \"Values disclosures and trust in science: A replication study\""
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

This repository contains anonymized data and code for the paper "Values disclosures and trust in science: A replication study" by Daniel J. Hicks and Emilio Lobato.

# Data files

Data files are contained in the folder `data`.

The files `data/data.csv` and `data/data.Rds` contain anonymized versions of the original data collected for this replication study.  An overview is included at the bottom of this readme.

The files `data/emad.csv` and `data/emad.Rds` contain versions of the data from Elliot et al. experiment 1, downloaded from <https://figshare.com/articles/dataset/Elliott_McCright_Allen_and_Dietz_Data_SPSS_Stata_/4695793> on 2021-06-09. This dataset was reshaped and variables recoded to facilitate parallel analysis with our original data using the code in script `01_clean.R`. This dataset is made available under a Creative Commons Attribution 4.0 International license.

# Reproducibility

To preserve participant anonymity, raw data (as downloaded from Qualtrics and cleaned by script `01_clean.R`) is not publicly available. Only analysis scripts `02_analysis.R` and `03_dag.R` are reproduced, along with assembling the manuscript PDF.

Either method below will require prior installation of R, the `renv` package, and [Quarto](https://quarto.org/).

## Using make

After cloning and downloading, in Bash or a similar shell run
```
make install
make
```

## Manually

In R, run
```
renv::restore()
```
to create a local (within-project) package library, with the specific versions used in this project.  Then run the scripts `02_analysis.R` and `03_dag.R`, in that order, to run the analyses.  These scripts will write plots and tables to the `out` folder, creating it if necessary.  Optionally, use `rmarkdown::render()` or the "Knit" button in RStudio to generate HTML documents from these scripts, for comparison with the ones included in the repository.  (Be sure to rename the included HTML files first.)

Finally, in the `paper` folder, from Bash or a similar shell run
```
quarto render paper.Qmd --to pdf
```
to generate the PDF version of the manuscript.

# Data overview

```{r}
library(skimr)

dataf = readRDS(file.path('data', 'data.Rds'))
skim(dataf)
```

### 2022-07-13 scripts/03_dag.R

@@ -21,6 +21,9 @@ library(ggdag)
 library(gt)
 library(gtsummary)
 
+## Need Hmisc for bootstrap CIs in some plots, but don't want to load it
+find.package('Hmisc')
+
 library(here)
 source(here('R', 'reg_tbl.R'))
 source(here('R', 'plot_adjustments.R'))
@@ -208,6 +211,14 @@ ggsave(here(out_dir, '03_part_values.png'),
        height = 4, width = 8, dpi = 200, scale = 1.5)
 
 table(dataf$political_ideology, dataf$pref)
+
+dataf |> 
+    mutate(political_ideology = case_when(
+        political_ideology < 4 ~ 'liberal', 
+        political_ideology == 4 ~ 'moderate', 
+        political_ideology > 4 ~ 'conservative'
+    )) |> 
+    count(political_ideology)
 
 cor(emad_df$ideology, emad_df$tradeoff, 
     use = 'complete.obs',

### 2022-07-15 paper/paper.Qmd
99% first draft

@@
 
-*[Anonymized original data and reproducible code are available at <https://github.com/dhicks/transparency>.  Instructions in that repository explain how to automatically reproduce our analysis.]*
+Anonymized original data and reproducible code are available at <https://github.com/dhicks/transparency>.  Instructions in that repository explain how to automatically reproduce our analysis.
 
@@
 Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling [@KennedyEvaluation2016Election2018].
-In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement]*.
+In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement?]*.
 This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  data from the General Social Survey shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public scientific controversies align with liberal-conservative partisanship [@FunkAmericansPoliticsScience2015].
@@
 
-*[maybe move to conclusion]*
-The evidence here suggests that if a scientist discloses their values alongside their presentation of scientific research, what values they share predicts their perceived trustworthiness, and that a scientist who discloses valuing public health will tend to be perceived as more trustworthy than the scientist who discloses valuing economic growth.
-
 # Conclusion #
@@
 
-We did not find evidence for a transparency penalty or shared values effect.  And we found evidence of a scientist values effect, which also did not appear in @ElliottValuesEnvironmentalResearch2017.
+We did not find evidence for a transparency penalty or shared values effect.
+And we found evidence of a scientist values effect (a scientist who discloses valuing public health is perceived as more trustworthy), which was not identified by @ElliottValuesEnvironmentalResearch2017.
 
@@
 One possible explanation is that, over the course of the Covid-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.
-(And, more recently, Leana Wen and some other prominent physicians and public health officials have argued for the importance of a "return to normal," which might be understood as prioritizing economic growth.)  A values disclosure that was regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.
+A values disclosure that might have been regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.
 
-*[further work]*
+The scientist values effect — a scientist who discloses valuing public health is perceived as more trustworthy — might be interpreted as supporting some of the claims of the "aims approach" in philosophy of science.
+This approach argues that scientific fields often have both social or practical aims, along with epistemic aims such as the pursuit of truth [@ElliottNonepistemicValuesMultiple2014; @IntemannDistinguishingLegitimateIllegitimate2015; @PotochnikIdealizationAimsScience2017; @FernandezPintoLegitimizingValuesRegulatory2019; @HicksWhenVirtuesAre2022].
+For example, the field of public health might have the practical aim of promoting the health of the public.
+Where certain values are seen as conflicting with these aims, it is morally wrong for a scientist to promote those values.
+Rather than violating the value-free ideal, a scientist who discloses valuing public health might be seen as following the norms of the field of public health.
+
+## Limitations
+
+As noted above, a major limitation of the current study is that the sample substantially underrepresents political conservatives.
+Our a priori DAG analysis and empirical robustness checks do not indicate any bias in our reported estimates due to this representation problem.
+However, our estimates could still be biased if any of the effects examined here have interactions with political ideology.
+
+Another limitation is that our sample only considers US adults.
+It does not include residents of any other country, and does not track where our participants might live within the US.
+Public scientific controversies often have different dynamics both across national borders and within different regions of the US.
+
+A third limitation is that the stimulus only considers a single controversy, over the safety of the chemical bisphenol A (BPA).
+Public attention to this controversy peaked around 2009.
+While it remains unsettled in terms of both science and policy, it is much less socially and politically salient than controversies over climate change or vaccination.
+
+## Directions for future work
+
+We suggest three directions for future work in this area, two in the experimental stimulus and one in the endpoint or outcome measured.
+
+The stimulus developed by @ElliottValuesEnvironmentalResearch2017 involves a single (fictional) scientist.
+But public scientific controversies often feature conflicting claims made by contesting experts and counterexperts.
+For example, as the Omicron wave of the Covid-19 pandemic waned in the US in spring 2022, various physicians, public health experts, journalists, and government officials made conflicting claims about the effective severity of Omicron, the relative effectiveness of masks in highly vaccinated populations, and the need to "return to normal" [@Adler-BellPandemicInterpreter2022; @KhullarWillCoronavirusPandemic2022; @YongHowDidThis2022].
+For members of the general public, the question was not whether to trust the claims of a given expert, but instead which experts to trust.
+Publics might perceive both expert A and expert B to be highly trustworthy, but would need to make a decision about who to trust if these experts are making conflicting claims.
+It would be straightforward to modify the single-scientist stimulus to cover this kind of "dueling experts" scenario.
+To modify METI, participants might be asked which of the two experts they would consider more competent, ethical, honest, etc.
+
+@BrownTrustExpertiseScientific2022 emphasized a distinction between individuals, groups, and institutions, as both trustors and trustees, and argued that much of the philosophical and empirical research on trust in science has focused on either individual-individual trust (as in @ElliottValuesEnvironmentalResearch2017) or individual-group trust (as in the General Social Survey, which asks individual respondents about their confidence in "the scientific community").
+In future experiments, "Dr.
+Riley Spence" could be represented with an institutional affiliation, such as "Dr.
+Riley Spence of the Environmental Protection Agency" or "Dr.
+Riley Spence of the American Petroleum Institute" Or claims could be attributed directly to institutions, such as "Environmental Protection Agency scientists" or "a report published by the American Petroleum Institute."
+
+Finally, research on the climate controversy suggests that acceptance of scientific claims might underpredict policy support.
+Since 2008, public opinion studies by the Yale Program on Climate Change Communication have found that only about 50-60% of the US public understands that there's a scientific consensus on climate change, that this understanding is politically polarized (conservatives are substantially less likely to recognize the consensus), but also that some climate policies enjoy majority support even among conservatives [@LeiserowitzPoliticsGlobalWarming2022].
+Future research could ask participants whether they would support, for example, increased regulation of BPA, either instead of or along with the perceived trustworthiness of the scientific expert.
 

### 2022-07-15 scripts/03_dag.R
update makefile

@@ -8,7 +8,6 @@
 # TODO: 
 #     -[] clean code
 #     -[] predicted value plots for E and F
-#     -[] webshot::install_phantomjs() in make install
 
 library(tidyverse)
 theme_set(theme_minimal())

### 2022-07-17 R/reg_plots.R
cleaned 03 code

@@ -24,7 +24,7 @@ plot_estimate = function(models, ...) {
                ymin = conf.low, ymax = conf.high, 
                color = study)) +
         geom_pointrange(position = position_dodge(width = .25)) +
-        geom_hline(yintercept = c(meaningful, -meaningful), linetype = 'dashed') +
+        geom_hline(yintercept = c(1, -1), linetype = 'dashed') +
         coord_flip()
 }
 

### 2022-07-17 R/reg_tbl.R
docx tables now autofit

@@ -21,20 +21,22 @@ reg_tbl = function(models,
 write_reg_tbl = function(tbl, path) {
     tbl |> 
         as_flex_table() |> 
-        flextable::save_as_docx(path = str_c(path, '.docx'),
-                                pr_section = officer::prop_section(
-                                    page_size = officer::page_size(width = 11, 
-                                                                   orient = 'landscape')
+        set_table_properties(layout = 'autofit') |> 
+        save_as_docx(path = str_c(path, '.docx'),
+                                pr_section = prop_section(
+                                    page_size = page_size(orient = 'landscape', 
+                                                          height = 11, 
+                                                          width = 8.5)
                                 )
         )
     ## Problems exporting gt to PDF are a known issue
     ## <https://github.com/rstudio/gt/issues/721>
-    tbl |> 
-        as_gt() |> 
-        gtsave(filename = str_c(path, '.pdf'), zoom = 1)
+    # tbl |> 
+    #     as_gt() |> 
+    #     gtsave(filename = str_c(path, '.pdf'), zoom = 1)
     
-    tbl |> 
-        tbl_butcher() |> 
-        # as_gt() |> 
+    tbl |>
+        tbl_butcher() |>
+        # as_gt() |>
         write_rds(file = str_c(path, '.Rds'))
 }

### 2022-07-19 paper/paper.Qmd
docx tables now autofit
update paper
updating the paper

@@
 The current paper reports a replication attempt of their experiment.
-Comparing results of the two replications, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
+Comparing results of the two *[EL: two?
+does a re-analysis of their data count as a replication?]* replications, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
 toc: true
@@
 @ElliottValuesEnvironmentalResearch2017 conducted an online survey study to evaluate this prediction.
+The authors found tentative evidence that disclosing values may reduce the perceived trustworthiness of a scientist, and that this effect may be moderated by whether or not participants share the same values as the scientist and whether or not the scientist reports findings contrary to their stated values.
 However, these authors collected data from a somewhat small sample using Amazon Mechanical Turk, and adopted an analytical approach that significantly diluted their sample across different conditions.
-
-In this paper, we report the results of a replication of @ElliottValuesEnvironmentalResearch2017 study 1, using a larger, more reliable sample and a more statistically efficient analytical approach.
+In this paper, we report the results of a replication of @ElliottValuesEnvironmentalResearch2017 study 1, using a larger sample and a more statistically efficient analytical approach.
+The major results found by @ElliottValuesEnvironmentalResearch2017 serve as the basis for our hypotheses, described below.
 
@@
 
-We use the same experimental stimulus and design as @ElliottValuesEnvironmentalResearch2017, embedded within a larger survey with other results not reported here.
-In the replication component, participants are randomly assigned to one of $3 \times 2$ (values \times harm) conditions.
+We used the same experimental stimulus and design as @ElliottValuesEnvironmentalResearch2017, which we embedded within a larger project examining the public's perceptions of the values in science.
+In this paper, we are only reporting the replication component of the project.
+The replication component was the same 3 (Values: No Disclosure, Economic Growth, Public Health) \times 2 (Harm: Causes Harm, Does Not Cause Harm) experimental design as used by @ElliottValuesEnvironmentalResearch2017.
 In each condition, participants are first shown a single presentation slide and the following explanatory text:
@@
 Spence is provided.
-The content of the slide varies across the conditions.
+The content of the slide varies across the conditions *[EL: is it possible to add a picture of the stimuli?]*.
 Each slide has the header "My conclusion," followed by a list of 2 or 3 bullets.
@@
 Spence's trustworthiness using a semantic differential scale.
-@ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
+@ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale *[EL: would it be fair to describe their scale as an ad hoc measure?]*; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
 On the METI, participants rate the target scientist (the fictional Dr.
-Spence) on a 1-7 scale for 14 items, where each item is anchored at the ends by a pair of words, such as competent-incompetent or responsible-irresponsible.
+Spence) on a 1-7 scale for 14 semantic differential items, where each item is anchored at the ends by a pair of words, such as competent-incompetent or responsible-irresponsible.
+The METI scale captures perceived trustworthiness of a target along three dimensions: competence, benevolence, and integrity.
+However, the three dimensions were strongly correlated in our sample.
+To avoid issues of multicollinearity in our analyses, we averaged participant scores into a single composite measure.
 In both @ElliottValuesEnvironmentalResearch2017 and our replication analysis, all items were averaged together to form a 1-7 composite measure of trustworthiness.
@@
 As its name indicates, METI assesses perceived trustworthiness, not trust (such as accepting Dr.
-Spencer's conclusion about BPA).
+Spence's conclusion about BPA).
+*[EL: could this be a limitation or future directions?
+measuring both trust and trustworthiness?]*
 
 @ElliottValuesEnvironmentalResearch2017 explain that they selected BPA as a complex, ongoing, public scientific controversy.
-For our replication, we chose to keep BPA, rather than switching to a different public scientific controversy with a higher profile in 2021, such as climate change, police violence, voter fraud, or any of numerous aspects of the Covid-19 pandemic.
+For our replication, we chose to keep BPA, rather than switching to a different public scientific controversy with a higher profile in 2021, such as climate change, police violence, voter fraud, or any of numerous aspects of the COVID-19 pandemic.
 All of these controversies are highly politically charged, with prominent public experts and counterexperts [@GoldenbergVaccineHesitancyPublic2021 100-1, ch.
@@
 Hypothesis 1 was analyzed using Spearman rank order correlation and visually.
-Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed a priori to identify appropriate adjustments (covariates) for H4 and 5.
+Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed *a priori* to identify appropriate adjustments (covariates) for H4 and 5.
 Because @ElliottValuesEnvironmentalResearch2017 made their data publicly available, and our analytical approaches differ somewhat from theirs, we conducted parallel analyses of their data, and report these parallel results for H2 and 3.
@@
 Preliminary power analysis recommended a sample of approximately 1,000 participants to reliably detect non-interaction effects (H1-4).
+
+After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample ($M_{age}$ = 44-years-old, $SD_{age}$ = 16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124, Asian or Pacific Islander = 63, Hispanic = 33, American Indian or Alaskan Native = 5, Mixed or Other = 51).
+Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition *[EL: I think a table would work better for this, but I don't know if there are limits on the number of tables allowed for Frontiers]*.
+Of the full 988 participants, 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).
+Consequently, subsamples for hypotheses 4 and 5 were substantially smaller than the full analysis sample.
+
+*[participant demographics]* *[EL: Demographics usually goes in the Participants section rather than results]*
 
@@
 
-After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample.
-660 participants (67%) were randomly assigned to a values disclosure condition (scientist prioritizes either economic growth or public health); 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).
-Consequently, subsamples for hypotheses 4 and 5 were substantially smaller than the full analysis sample.
-
-*[participant demographics]*
-
 Critically, our data are unlikely to be representative by education level and political ideology.
@@
 Because of these representation issues, insofar as some political conservatives are both less likely to participate in studies on Prolific (or at least in our particular study) and likely to perceive Spence as less trustworthy, this will produce omitted variable bias for analyses that require adjustment by political ideology.
-Analysis of our a priori DAG indicated that, for the hypotheses examined here, this adjustment was not necessary in any case.
+Analysis of our *a priori* DAG indicated that, for the hypotheses examined here, this adjustment was not necessary in any case.
 When some adjustment was necessary (H4-5), we included political ideology along with other demographic variables in an alternative model specification as a robustness check.
@@
 First, to test (H1a) whether political liberals are more likely to prioritize public health over economic growth compared to conservatives, we conducted a Spearman's rank order correlation.
-Results revealed a significant correlation in line with the hypothesis, Spearman's rho = -.47, p < .001.
+Results revealed a significant correlation in line with the hypothesis, Spearman's $\rho $ = -.47, *p* < .001.
 Political liberals were more likely than conservatives to value public health over economic growth.
@@
 The full model was significant, adj.
-$R^2$= .148, F(2, 985) = 86.59, p < .001 (see @tbl-bc).
+$R^2$ = .148, *F*(2, 985) = 86.59, *p* < .001 (see @tbl-bc).
 Specifically, results revealed that the conclusions reported by the scientist predicted participants' perceived trustworthiness in line with our hypothesis.
-Participants rated the scientist who reported that BPA does not cause harm as less trustworthy ($M_{sd} = 4.48_{1.26}$) than the scientist who reported that BPA causes harm ($M_{sd} = 5.47_{1.12}$), $\beta$ = -0.99, t(985) = -13.09, p < .001.
-By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values ($M_{sd} = 4.94_{1.33}$) are perceived as less trustworthy than a scientist who does not disclose values  ($M_{sd} = 5.05_{1.21}$), $\beta$ = -0.12, t(985) = -1.46, p = .143.
+Participants rated the scientist who reported that BPA does not cause harm as less trustworthy ($M_{sd} = 4.48_{1.26}$) than the scientist who reported that BPA causes harm ($M_{sd} = 5.47_{1.12}$), $\beta$ = -0.99, *t*(985) = -13.09, *p* < .001.
+By contrast, the results do not provide evidence in favor of our hypothesis that a scientist disclosing their values ($M_{sd} = 4.94_{1.33}$) are perceived as less trustworthy than a scientist who does not disclose values  ($M_{sd} = 5.05_{1.21}$), $\beta$ = -0.12, *t*(985) = -1.46, *p* = .143.
 
@@
 While the univariate model was significant, adj.
-$R^2$= .029, F(1, 565) = 18.0, p < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.
+$R^2$= .029, *F*(1, 565) = 18.0, *p* < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.
 Rather, only the scientist's disclosed values significantly predicted how trustworthy participants rated the scientist, such that a scientist who disclosed valuing public health was rated as more trustworthy than a scientist who disclosed valuing economic growth.
@@
 The full model was significant, adj.
-$R^2$ = .173, F(3, 840) = 59.95, p < .001 ([@tbl-eb model 2]).
-As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta$ = -0.75, t(840) = -4.51, p < .001.
-However, this effect was qualified by a significant interaction with participants' values, $\beta$ = -0.41, t(840) = -2.16, p = .031.
+$R^2$ = .173, *F*(3, 840) = 59.95, *p* < .001 ([@tbl-eb model 2]).
+As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta$ = -0.75, *t*(840) = -4.51, *p* < .001.
+However, this effect was qualified by a significant interaction with participants' values, $\beta$ = -0.41, *t*(840) = -2.16, *p* = .031.
 Participants who prioritized public health and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants with the same values who read about a scientist who concluded that BPA does not cause harm ($M_{sd} = 4.40_{1.24}$) ([@fig-conclusion-part]).
@@
 The full model was not significant, adj.
-$R^2$< .001, F(3, 840) = 1.19, p = .31 [@tbl-ec model 2].
-As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta$ = -0.23, t(840) = -1.07, p = 0.286.
+$R^2$< .001, *F*(3, 840) = 1.19, *p* = .31 [@tbl-ec model 2].
+As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta$ = -0.23, *t*(840) = -1.07, *p* = 0.286.
 
@@
 This model was significant, adj.
-$R^2$ = 0.065, F(3, 563) = 14.15, p < .001.
-The estimate for the scientist values term was significant, $\beta$ = 0.53, t(563) = 2.26, p = 0.02, while the estimate for the interaction term was not, $\beta$ = 0.20, t(563) = 0.75, p = 0.45.
+$R^2$ = 0.065, *F*(3, 563) = 14.15, *p* < .001.
+The estimate for the scientist values term was significant, $\beta$ = 0.53, *t*(563) = 2.26, *p* = 0.02, while the estimate for the interaction term was not, $\beta$ = 0.20, *t*(563) = 0.75, *p* = 0.45.
 Estimates for both variables had large uncertainties, with confidence intervals of about 1 (95% confidence intervals, scientist values: 0.07, 1.0; interaction term: -0.32, 0.72).
@@
 @ElliottValuesEnvironmentalResearch2017 claim evidence for a shared values effect by comparing values disclosure vs.
-no disclosure (that is, estimating a transparency effect) in quasi-independent subsamples across 8 cells (participant values $\times$ scientist conclusion $\times$ scientist values; @ElliottValuesEnvironmentalResearch2017 fig.
+no disclosure (that is, estimating a transparency effect) in quasi-independent subsamples across a 2 $\times$ 2 $\times$ 2 design (participant values $\times$ scientist conclusion $\times$ scientist values; @ElliottValuesEnvironmentalResearch2017 fig.
 1).
@@
 But these analyses are underpowered (for example, sample sizes are well below 100 for some of the cells in which participants value economic growth) and the overall approach cannot distinguish a shared values effect from other potential effects (scientist values, participant values).
-Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$=0.29, t(335)=1.75, p=0.08; while the estimate for scientist values was statistically significant, $\beta$=0.41, t(335)=2.48, p=0.014.
+Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$ = 0.29, *t*(335) = 1.75, *p* = 0.08; while the estimate for scientist values was statistically significant, $\beta$ = 0.41, *t*(335) = 2.48, *p* = 0.014.
 It seems likely that claims of a shared values effect were the result of misinterpreting a confusing statistical approach.
 *[trying to put this as nice as possible]*
+*[EL: suggested alternative "...could be the result of a less appropriate analytic approach given the nature of their data"]*
 
@@
 Using our approach and the data from @ElliottValuesEnvironmentalResearch2017, the estimate for a values disclosure was statistically significant (@tbl-bc model 1).
-One possible explanation is that, over the course of the Covid-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.
+One possible explanation is that, over the course of the COVID-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.
 A values disclosure that might have been regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.
@@
 As noted above, a major limitation of the current study is that the sample substantially underrepresents political conservatives.
-Our a priori DAG analysis and empirical robustness checks do not indicate any bias in our reported estimates due to this representation problem.
+Our *a priori* DAG analysis and empirical robustness checks do not indicate any bias in our reported estimates due to this representation problem.
 However, our estimates could still be biased if any of the effects examined here have interactions with political ideology.
@@
 Public scientific controversies often have different dynamics both across national borders and within different regions of the US.
+*[EL: this strikes me as a strong enough claim to warrant a citation]*
 
@@
 Public attention to this controversy peaked around 2009.
-While it remains unsettled in terms of both science and policy, it is much less socially and politically salient than controversies over climate change or vaccination.
+While it remains unsettled in terms of both science and policy, it is much less socially and politically salient than controversies over topics such as climate change or vaccination.
+This limitation affects our ability to speculate about the generalizability of our findings.
+That is, the effects we reported may only apply narrowly to a subset of sociopolitically controversial scientific topics that have not achieved the degree of salience or persistence in sociopolitical discourse for as long as topics like climate change or vaccinations.
 
@@
 We suggest three directions for future work in this area, two in the experimental stimulus and one in the endpoint or outcome measured.
+*[EL: Should we include a future work direction foreshadowing the other aspect of this project, looking at what values or perceptions people have about what science is or should be?]*
 
 The stimulus developed by @ElliottValuesEnvironmentalResearch2017 involves a single (fictional) scientist.
-But public scientific controversies often feature conflicting claims made by contesting experts and counterexperts.
-For example, as the Omicron wave of the Covid-19 pandemic waned in the US in spring 2022, various physicians, public health experts, journalists, and government officials made conflicting claims about the effective severity of Omicron, the relative effectiveness of masks in highly vaccinated populations, and the need to "return to normal" [@Adler-BellPandemicInterpreter2022; @KhullarWillCoronavirusPandemic2022; @YongHowDidThis2022].
+But public scientific controversies often feature conflicting claims made by contesting experts and counterexperts *[EL: as well as non-experts]*.
+For example, as the Omicron wave of the COVID-19 pandemic waned in the US in spring 2022, various physicians, public health experts, journalists, and government officials made conflicting claims about the effective severity of Omicron, the relative effectiveness of masks in highly vaccinated populations, and the need to "return to normal" [@Adler-BellPandemicInterpreter2022; @KhullarWillCoronavirusPandemic2022; @YongHowDidThis2022].
 For members of the general public, the question was not whether to trust the claims of a given expert, but instead which experts to trust.

### 2022-07-19 readme.Rmd

@@
 
-After cloning and downloading, in Bash or a similar shell run
+After cloning and downloading, in bash or a similar shell run
 ```
@@
 ```
+
+The installation step primarily installs R packages.  This can take a long time if many packages need to be downloaded.
 
@@
 These scripts will write plots and tables to the `out` folder, creating it if necessary.
+
 Optionally, use `rmarkdown::render()` or the "Knit" button in RStudio to generate HTML documents from these scripts, for comparison with the ones included in the repository.
@@
 
-Finally, in the `paper` folder, from Bash or a similar shell run
+Finally, in the `paper` folder, from bash or a similar shell run
 ```

### 2022-07-21 paper/paper.Qmd
Emilio's notes

@@
 The current paper reports a replication attempt of their experiment.
-Comparing results of the two *[EL: two?
-does a re-analysis of their data count as a replication?]* replications, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
+Comparing results of the original study and our replication, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
 toc: true
@@
     pdf: default
+always_allow_html: true
 ---
@@
 Spence is provided.
-The content of the slide varies across the conditions *[EL: is it possible to add a picture of the stimuli?]*.
+The content of the slide varies across the conditions (@fig-stimulus).
 Each slide has the header "My conclusion," followed by a list of 2 or 3 bullets.
@@
 
+![Example of the experimental stimulus, from the "public health values disclosure" + "causes harm" conclusion.](fig1_stimulus.png){#fig-stimulus}
+
 After viewing the slide, the participant is then asked to rate Dr.
 Spence's trustworthiness using a semantic differential scale.
-@ElliottValuesEnvironmentalResearch2017 did not cite a source for their scale *[EL: would it be fair to describe their scale as an ad hoc measure?]*; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
+@ElliottValuesEnvironmentalResearch2017 used an ad hoc scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
 On the METI, participants rate the target scientist (the fictional Dr.
@@
 Spence's conclusion about BPA).
-*[EL: could this be a limitation or future directions?
-measuring both trust and trustworthiness?]*
 
@@
 After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample ($M_{age}$ = 44-years-old, $SD_{age}$ = 16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124, Asian or Pacific Islander = 63, Hispanic = 33, American Indian or Alaskan Native = 5, Mixed or Other = 51).
-Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition *[EL: I think a table would work better for this, but I don't know if there are limits on the number of tables allowed for Frontiers]*.
+Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition (@tbl-condition).
 Of the full 988 participants, 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).
@@
 
-*[participant demographics]* *[EL: Demographics usually goes in the Participants section rather than results]*
+```{r}
+#| label: tbl-condition
+#| tbl-cap: "Assignment of participants to conditions"
+readRDS(here(out_dir, '03_condition_tbl.Rds')) |> 
+	print_tbl()
+```
 
@@
 Both overrepresentation of college graduates and underrepresentation of conservatives (especially conservatives with strong anti-institutional views) are known issues in public opinion polling [@KennedyEvaluation2016Election2018].
-In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness *[supplement?]*.
+In exploratory data analysis, we noted that there was essentially no correlation between political ideology and perceived trustworthiness (in the online supplement, see subsection H5-shared).
 This was surprising, since general trust/distrust in science has become a partisan phenomenon over the last few decades:  data from the General Social Survey shows increasing trust in science from liberals and decreasing trust from conservatives [@GauchatPoliticizationSciencePublic2012; @LeePartyPolarizationTrust2021], and many (though not all) prominent public scientific controversies align with liberal-conservative partisanship [@FunkAmericansPoliticsScience2015].
@@
 Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$ = 0.29, *t*(335) = 1.75, *p* = 0.08; while the estimate for scientist values was statistically significant, $\beta$ = 0.41, *t*(335) = 2.48, *p* = 0.014.
-It seems likely that claims of a shared values effect were the result of misinterpreting a confusing statistical approach.
-*[trying to put this as nice as possible]*
-*[EL: suggested alternative "...could be the result of a less appropriate analytic approach given the nature of their data"]*
+It seems likely that claims of a shared values effect could have been the result of a less appropriate analytic approach given the nature of the data.
 
@@
 It does not include residents of any other country, and does not track where our participants might live within the US.
-Public scientific controversies often have different dynamics both across national borders and within different regions of the US.
-*[EL: this strikes me as a strong enough claim to warrant a citation]*
+Public scientific controversies often have different dynamics both across national borders and within different regions of geographically large countries [@MilesPublicPerceptionScientific2003; @HoweGeographicVariationOpinions2015; @MildenbergerDistributionClimateChange2016; @MildenbergerSpatialDistributionRepublican2017; @SturgisTrustScienceSocial2021].
 
@@
 
+The design from @ElliottValuesEnvironmentalResearch2017 assesses public perceptions of the value-free ideal indirectly, by probing whether violations of this ideal would lead to decreased perceived trustworthiness.
+But do members of the general public accept the value-free ideal?
+Concurrently with this replication study, we also asked participants directly for their views on the value-free ideal and related issues and arguments from the philosophy of science literature.
+A manuscript discussing this effort to develop a "values in science scale" is currently under preparation.
+
 We suggest three directions for future work in this area, two in the experimental stimulus and one in the endpoint or outcome measured.
-*[EL: Should we include a future work direction foreshadowing the other aspect of this project, looking at what values or perceptions people have about what science is or should be?]*
 
 The stimulus developed by @ElliottValuesEnvironmentalResearch2017 involves a single (fictional) scientist.
-But public scientific controversies often feature conflicting claims made by contesting experts and counterexperts *[EL: as well as non-experts]*.
-For example, as the Omicron wave of the COVID-19 pandemic waned in the US in spring 2022, various physicians, public health experts, journalists, and government officials made conflicting claims about the effective severity of Omicron, the relative effectiveness of masks in highly vaccinated populations, and the need to "return to normal" [@Adler-BellPandemicInterpreter2022; @KhullarWillCoronavirusPandemic2022; @YongHowDidThis2022].
+But public scientific controversies often feature conflicting claims made by contesting experts and counterexperts.
+For example, as the Omicron wave of the COVID-19 pandemic waned in the US in spring 2022, various physicians, public health experts, science journalists, and government officials made conflicting claims about the effective severity of Omicron, the relative effectiveness of masks in highly vaccinated populations, and the need to "return to normal" [@Adler-BellPandemicInterpreter2022; @KhullarWillCoronavirusPandemic2022; @YongHowDidThis2022].
 For members of the general public, the question was not whether to trust the claims of a given expert, but instead which experts to trust.
@@
 To modify METI, participants might be asked which of the two experts they would consider more competent, ethical, honest, etc.
+(Public scientific controversies can include other prominent actors with no relevant expertise, such as political journalists, elected officials, or social media conspiracy theorists.
+Such individuals might nonetheless be treated by publics as trusted sources for factual claims.
+Understanding why many people might trust a figure like Alex Jones over someone like Anthony Fauci would likely require prior work disentangling trustworthiness, expertise, and institutional standing.)
 
@@
 
-Finally, research on the climate controversy suggests that acceptance of scientific claims might underpredict policy support.
+As noted above, ethicists make a distinction between trust and trustworthiness.
+Instruments such as METI assess trustworthiness: whether a speaker is perceived to have qualities that would make it appropriate to trust them.
+Trust itself is closer to behavior than attitudes, and would probably be better measured (in online survey experiments and similar designs) by asking whether participants accept claims made by speakers or support policy positions endorsed by speakers.
+In addition, research on the climate controversy suggests that acceptance of scientific claims might underpredict policy support.
 Since 2008, public opinion studies by the Yale Program on Climate Change Communication have found that only about 50-60% of the US public understands that there's a scientific consensus on climate change, that this understanding is politically polarized (conservatives are substantially less likely to recognize the consensus), but also that some climate policies enjoy majority support even among conservatives [@LeiserowitzPoliticsGlobalWarming2022].
-Future research could ask participants whether they would support, for example, increased regulation of BPA, either instead of or along with the perceived trustworthiness of the scientific expert.
+Putting these points together, future research could ask participants whether they would support, for example, increased regulation of BPA, either instead of or along with assessing the perceived trustworthiness of the scientific expert.
 

### 2022-07-21 scripts/03_dag.R

@@ -49,6 +49,7 @@ dataf = read_rds(here(data_dir, 'data.Rds'))
 
 #' # Descriptive summary of our data
 ## Descriptive summary of our data ----
+#' ## Demographics
 #' - age
 #' - gender
 #' - race/ethnicity
@@ -157,8 +158,23 @@ demo_gt |>
                  )
     )
 
+#' ## Condition assignment
+condition_tbl = dataf |> 
+    mutate(disclose_values = case_when(
+        !disclosure ~ 'no disclosure', 
+        disclosure ~ sci_values), 
+        disclose_values = fct_relevel(disclose_values, 
+                                      'no disclosure',
+                                      'public health', 
+                                      'economic growth')) |> 
+    select(disclose_values, conclusion) |> 
+    tbl_summary(by = conclusion,
+                label = list(disclose_values ~ 'disclosure/values'))
+condition_tbl
+write_reg_tbl(condition_tbl, here(out_dir, '03_condition_tbl'))
 
-#' # Trust, overall
+
+#' ## Trust, overall
 ## Trust, overall ----
 ggplot() +
     geom_violin(aes(x = 'EMAD', pa_mean), 

### 2022-07-25 paper/paper.Qmd
as emailed to KE, 2022-07-25

@@
 Keeping with the example, the question of whether chemical X causes cancer has much more significant social and political implications than the question of whether chemical X fluoresces under ultraviolet light.
-So, in the interest of protecting human health, it would be appropriate to *scientifically* accept the hypothesis that chemical X is carcinogenic based on relatively provisional evidence [@CranorSocialBenefitsExpedited1995; @ElliottNonepistemicValuesMultiple2014; @FernandezPintoLegitimizingValuesRegulatory2019].
+So, in the interest of protecting human health, it would be appropriate to *scientifically* accept the hypothesis that chemical X is carcinogenic for regulatory purposes based on relatively weak or provisional evidence [@CranorSocialBenefitsExpedited1995; @ElliottNonepistemicValuesMultiple2014; @FernandezPintoLegitimizingValuesRegulatory2019].
 
@@
 The authors found tentative evidence that disclosing values may reduce the perceived trustworthiness of a scientist, and that this effect may be moderated by whether or not participants share the same values as the scientist and whether or not the scientist reports findings contrary to their stated values.
-However, these authors collected data from a somewhat small sample using Amazon Mechanical Turk, and adopted an analytical approach that significantly diluted their sample across different conditions.
+However, these authors collected data from a somewhat small sample using Amazon Mechanical Turk, and adopted an analytical approach that diluted their sample across different conditions.
 In this paper, we report the results of a replication of @ElliottValuesEnvironmentalResearch2017 study 1, using a larger sample and a more statistically efficient analytical approach.
@@
 First, to test (H1a) whether political liberals are more likely to prioritize public health over economic growth compared to conservatives, we conducted a Spearman's rank order correlation.
-Results revealed a significant correlation in line with the hypothesis, Spearman's $\rho $ = -.47, *p* < .001.
+Results revealed a significant correlation in line with the hypothesis, Spearman's $\rho$ = -.47, *p* < .001.
 Political liberals were more likely than conservatives to value public health over economic growth.
@@
 Following this, we created a new Shared Values variable as a composite of the participants' reported values and the scientist's values.
-For this analysis, an a priori directed acyclic graphic (DAG) indicated that a univariate regression of participants' METI rating onto Shared Values would produce a biased estimate, that adjustments were required for both the participant's and scientist's values, and that including demographic variables would not effect this estimate [@fig-shared-dag].
+For this analysis, an a priori directed acyclic graphic (DAG) indicated that a univariate regression of participants' METI rating onto Shared Values would produce a biased estimate, that adjustments were required for both the participant's and scientist's values, and that including demographic variables would not effect this estimate (@fig-shared-dag).
 While the univariate model was significant, adj.
-$R^2$= .029, *F*(1, 565) = 18.0, *p* < .001 [@tbl-shared model 1], after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist [@tbl-shared models 2-4], indicating that the univariate estimate was indeed biased.
+$R^2$= .029, *F*(1, 565) = 18.0, *p* < .001 (@tbl-shared model 1), after adjusting for scientist values Shared Values did not emerge as a significant predictor of participants' perceptions of the trustworthiness of the scientist (@tbl-shared models 2-4), indicating that the univariate estimate was indeed biased.
 Rather, only the scientist's disclosed values significantly predicted how trustworthy participants rated the scientist, such that a scientist who disclosed valuing public health was rated as more trustworthy than a scientist who disclosed valuing economic growth.
-This scientist values effect was robust across alternative model specifications, with an estimated effect of 0.6 (95% CI 0.4-0.8) [@tbl-shared models 2-5], again consistent with the DAG.
+This scientist values effect was robust across alternative model specifications, with an estimated effect of 0.6 (95% CI 0.4-0.8; @tbl-shared models 2-5), again consistent with the DAG.
 
@@
 The full model was significant, adj.
-$R^2$ = .173, *F*(3, 840) = 59.95, *p* < .001 ([@tbl-eb model 2]).
+$R^2$ = .173, *F*(3, 840) = 59.95, *p* < .001 (@tbl-eb model 2).
 As with the earlier analysis, results showed a main effect of the Conclusions condition, such that participants rated the scientist who reported that BPA does not cause harm as less trustworthy than the scientist who reported that BPA causes harm, $\beta$ = -0.75, *t*(840) = -4.51, *p* < .001.
 However, this effect was qualified by a significant interaction with participants' values, $\beta$ = -0.41, *t*(840) = -2.16, *p* = .031.
-Participants who prioritized public health and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants with the same values who read about a scientist who concluded that BPA does not cause harm ($M_{sd} = 4.40_{1.24}$) ([@fig-conclusion-part]).
+Participants who prioritized public health and read about a scientist who concluded that BPA causes harm rated the scientist as more trustworthy ($M_{sd} = 5.56_{1.09}$) than participants with the same values who read about a scientist who concluded that BPA does not cause harm ($M_{sd} = 4.40_{1.24}$; @fig-conclusion-part).
 
@@
 The full model was not significant, adj.
-$R^2$< .001, *F*(3, 840) = 1.19, *p* = .31 [@tbl-ec model 2].
+$R^2$< .001, *F*(3, 840) = 1.19, *p* = .31 (@tbl-ec model 2).
 As with the earlier analysis, our results do not provide evidence for a transparency penalty to the perceived trustworthiness of a scientist, either in general or interacting with participants' own reported values, $\beta$ = -0.23, *t*(840) = -1.07, *p* = 0.286.
@@
 
-Without adjustments, Shared Values appears to have a substantial interaction with Participant Values:  Shared Values appears to increase perceived trustworthiness for participants who value public health, while decreasing perceived trustworthiness for participants who value economic growth [@fig-shared-part].
+Without adjustments, Shared Values appears to have a substantial interaction with Participant Values:  Shared Values appears to increase perceived trustworthiness for participants who value public health, while decreasing perceived trustworthiness for participants who value economic growth (@fig-shared-part).
 However, as indicated by our analysis for a potential shared values effect, the estimated effects for both shared values and participant values are biased if the model is not adjusted for scientist values.
@@
 
-![Data, group means, and 95% confidence intervals for H5d, interaction of shared values and participant values.
+![Data, group means, and 95% confidence intervals for H5-shared, interaction of shared values and participant values.
 Panels correspond to participant values.](`r here(out_dir, '03_shared_part.png')`){#fig-shared-part}

### 2022-08-11 paper/paper.Qmd
ready to submit to Frontiers

@@
 title: "Values disclosures and trust in science: A replication study"
-author: "Daniel J. Hicks and Emilio Lobato"
+# author: 
+#     - name: Daniel J. Hicks
+#       orcid: 0000-0001-7945-4416
+#       email: dhicks4@ucmerced.edu
+#       affiliations:
+#           - ref: ucm
+#       attributes:
+#           corresponding: true
+#     - name: Emilio Lobato
+#       orcid: 0000-0002-3066-2932
+#       email: elobato@ucmerced.edu
+#       affiliations: 
+#         - ref: ucm
+
+affiliations: 
+    - id: ucm
+      name: University of California, Merced
+      address: 5200 N Lake Road
+      city: Merced
+      state: CA
+      postal-code: 95343
+
 abstract: |
@@
 Comparing results of the original study and our replication, we do not find evidence for a transparency penalty or "shared values" effect, but do find evidence that the content of scientific conclusions (whether or not a chemical is found to cause harm) might effect perceived trustworthiness and that scientists who value public health and disclose this value might be perceived as more trustworthy.
-toc: true
+
+toc: false
 execute:
@@
 always_allow_html: true
+header-includes:
+  - \usepackage[left]{lineno}
+  - \linenumbers
 ---
+
+| Daniel J. Hicks^1\*^, Emilio Lobato^1^
+| ^1^: University of California, Merced, 5200 N Lake Road, Merced, CA, 95343
+| ^\*^: Address correspondence to <dhicks4@ucmerced.edu>
+| 
+| Keywords: trust, trust in science, values in science, replication, philosophy of science
 
@@
 (A) Absolute counts, (B) shares within political ideology categories.
-Panel (A) shows that our sample substantially over-represents political liberals.](`r here(out_dir, '03_part_values.png')`){#fig-part-values}
+Panel (A) shows that our sample substantially over-represents political liberals.](fig2_part_values.png){#fig-part-values}
 
@@
 If the DAG is faithful, adjusting for scientist values (`sci_values`) and participant values (`part_values`) is sufficient for estimating the effect of shared values on perceived trustworthiness (`METI`).
-In particular, adjusting for demographics should not change the regression coefficient on shared values.](`r here(out_dir, '03_shared_values_dag.png')`){#fig-shared-dag}
+In particular, adjusting for demographics should not change the regression coefficient on shared values.](fig3_shared_values_dag.png){#fig-shared-dag}
 
@@
 ![Data, group means, and 95% confidence intervals for H5-consumer, interaction of consumer risk sensitivity and participant values.
-Panels correspond to participant values.](`r here(out_dir, '03_conclusion_part.png')`){#fig-conclusion-part}
+Panels correspond to participant values.](fig4_conclusion_part.png){#fig-conclusion-part}
 
@@
 
-![Data, group means, and 95% confidence intervals for H5-shared, interaction of shared values and participant values.
-Panels correspond to participant values.](`r here(out_dir, '03_shared_part.png')`){#fig-shared-part}
+![Data, group means, and 95% confidence intervals for H5-shared, interaction of shared values and participant values.  Panels correspond to participant values.](fig5_shared_part.png){#fig-shared-part}
 
@@
 Panels correspond to participant values.
-This figure is identical to @fig-shared-part, except the columns in the economic growth panel have been switched.](`r here(out_dir, '03_sci_part.png')`){#fig-sci-part}
+This figure is identical to @fig-shared-part, except the columns in the economic growth panel have been switched.](fig6_sci_part.png){#fig-sci-part}
 
@@
 
+# Acknowledgments #
+
+For feedback on this project, thanks to Kevin Elliott and participants at the Values in Science conference and Political Philosophy and Values in Medicine, Science, and Technology 2022 conference.
+
+# Funding #
+
+Funding for this project was provided by the University of California, Merced.
+
 # References #

### 2022-08-11 paper/paper.tex
ready to submit to Frontiers

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage[left]{lineno}
\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Values disclosures and trust in science: A replication study},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Values disclosures and trust in science: A replication study}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
While philosophers of science generally agree that social, political,
and ethical values can play legitimate roles in science, there is active
debate over whether scientists should disclosure such values in their
public communications. This debate depends, in part, on empirical claims
about whether values disclosures might undermine public trust in
science. In a previous study, Elliott et al.~used an online experiment
to test this empirical claim. The current paper reports a replication
attempt of their experiment. Comparing results of the original study and
our replication, we do not find evidence for a transparency penalty or
``shared values'' effect, but do find evidence that the content of
scientific conclusions (whether or not a chemical is found to cause
harm) might effect perceived trustworthiness and that scientists who
value public health and disclose this value might be perceived as more
trustworthy.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi

Daniel J. Hicks\textsuperscript{1*}, Emilio Lobato\textsuperscript{1}\\
\textsuperscript{1}: University of California, Merced, 5200 N Lake Road,
Merced, CA, 95343\\
\textsuperscript{*}: Address correspondence to
\href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}\\
\strut \\
Keywords: trust, trust in science, values in science, replication,
philosophy of science

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Over the past 15 years, many philosophers of science have rejected the
ideal of value-free science and related, traditional understandings of
objectivity and political neutrality of science (Douglas 2009; Elliott
2017). According to the value-free ideal, social and political values
--- such as feminism, environmentalism, or the protection of human
health --- have no legitimate role to play in the evaluation of
scientific hypotheses. The value-free ideal is compatible with allowing
social and political values to play important roles earlier and later in
inquiry. Specifically, these values may legitimately shape the content
and framing of research questions --- researchers might decide to
investigate whether chemical X causes cancer out of a concern to protect
human health --- and will be essential when scientific findings are used
to inform public policy --- say, banning the use of chemical X. But,
according to the value-free ideal, these values must not influence the
collection and analysis of data and the process of reaching an overall
conclusion about whether or not chemical X causes cancer.

Challenges to the value-free ideal argue that at least some social and
political values may, or even should, play a role in the evaluation of
scientific hypotheses. Keeping with the example, the question of whether
chemical X causes cancer has much more significant social and political
implications than the question of whether chemical X fluoresces under
ultraviolet light. So, in the interest of protecting human health, it
would be appropriate to \emph{scientifically} accept the hypothesis that
chemical X is carcinogenic for regulatory purposes based on relatively
weak or provisional evidence (Cranor 1995; Elliott and McKaughan 2014;
Fernández Pinto and Hicks 2019).

While this kind of argument indicates that at least some social and
political values may or should play at least some role in the evaluation
of at least some scientific hypotheses, it doesn't provide us with much
positive guidance: which values, playing which roles, in the evaluation
of which hypotheses? (Hicks 2014) One (partial) answer to this set of
questions appeals to the ideal of transparency: scientists should
disclose to the public the social and political values that have
influenced their research (Elliott and Resnik 2014; McKaughan and
Elliott 2013). But some philosophers have objected to this transparency
proposal, arguing that it might actually undermine trust in science.
According to this objection, members of the public generally accept the
ideal of value-free science; values disclosures would violate this
ideal, making scientists (incorrectly) appear biased and untrustworthy
(John 2017; Kovaka 2021).

This objection is an empirical prediction: if scientists disclose their
values, they will be perceived as less trustworthy. Elliott et al.
(2017) conducted an online survey study to evaluate this prediction. The
authors found tentative evidence that disclosing values may reduce the
perceived trustworthiness of a scientist, and that this effect may be
moderated by whether or not participants share the same values as the
scientist and whether or not the scientist reports findings contrary to
their stated values. However, these authors collected data from a
somewhat small sample using Amazon Mechanical Turk, and adopted an
analytical approach that diluted their sample across different
conditions. In this paper, we report the results of a replication of
Elliott et al. (2017) study 1, using a larger sample and a more
statistically efficient analytical approach. The major results found by
Elliott et al. (2017) serve as the basis for our hypotheses, described
below.

\hypertarget{methods-and-materials}{%
\section{Methods and Materials}\label{methods-and-materials}}

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

We used the same experimental stimulus and design as Elliott et al.
(2017), which we embedded within a larger project examining the public's
perceptions of the values in science. In this paper, we are only
reporting the replication component of the project. The replication
component was the same 3 (Values: No Disclosure, Economic Growth, Public
Health) \times 2 (Harm: Causes Harm, Does Not Cause Harm) experimental
design as used by Elliott et al. (2017). In each condition, participants
are first shown a single presentation slide and the following
explanatory text:

\begin{quote}
For several decades, a scientist named Dr.~Riley Spence has been doing
research on chemicals used in consumer products. One
chemical---Bisphenol A, popularly known as BPA---is found in a large
number of consumer products and is suspected of posing risks to human
health. Yet, scientists do not agree about these possible health risks.
Dr.~Spence recently gave a public talk in Washington, D.C., about BPA
research. Here is the final slide from Dr.~Spence's presentation.
\end{quote}

No other information about the (fictional) Dr.~Spence is provided. The
content of the slide varies across the conditions
(Figure~\ref{fig-stimulus}). Each slide has the header ``My
conclusion,'' followed by a list of 2 or 3 bullets. The first bullet, if
present, makes a values disclosure, stating that either economic growth
or public health should be a top national priority. This first bullet is
not present in the ``no disclosure'' values condition. The second
bullet, identical across conditions, states ``I examined the scientific
evidence on potential health risks of BPA.'' The third bullet concludes
that BPA either does or does not cause harm to people, depending on
whether the subject is in the ``causes harm'' or ``does not cause harm''
condition.

\begin{figure}

{\centering \includegraphics{fig1_stimulus.png}

}

\caption{\label{fig-stimulus}Example of the experimental stimulus, from
the ``public health values disclosure'' + ``causes harm'' conclusion.}

\end{figure}

After viewing the slide, the participant is then asked to rate
Dr.~Spence's trustworthiness using a semantic differential scale.
Elliott et al. (2017) used an ad hoc scale; we used the Muenster
Epistemic Trustworthiness Inventory (METI; Hendriks, Kienhues, and
Bromme 2015), which substantially overlaps but is not identical to the
Elliott et al. (2017) scale. On the METI, participants rate the target
scientist (the fictional Dr.~Spence) on a 1-7 scale for 14 semantic
differential items, where each item is anchored at the ends by a pair of
words, such as competent-incompetent or responsible-irresponsible. The
METI scale captures perceived trustworthiness of a target along three
dimensions: competence, benevolence, and integrity. However, the three
dimensions were strongly correlated in our sample. To avoid issues of
multicollinearity in our analyses, we averaged participant scores into a
single composite measure. In both Elliott et al. (2017) and our
replication analysis, all items were averaged together to form a 1-7
composite measure of trustworthiness. To aid interpretation, in the
analysis the direction of the scale has been set so that increased
values correspond to increased perceived trustworthiness.

Ethicists make a key distinction between trust and trustworthiness
(Baier 1986). Trust is a verb: placing one's trust in someone else, with
respect to some activity or domain. Trustworthiness is an assessment, of
whether or not that trust is appropriate. In the experiment,
participants might trust Dr.~Spence's conclusion even if they judge
Spence to be untrustworthy, or vice versa; though we would typically
expect these two to go together. As its name indicates, METI assesses
perceived trustworthiness, not trust (such as accepting Dr.~Spence's
conclusion about BPA).

Elliott et al. (2017) explain that they selected BPA as a complex,
ongoing, public scientific controversy. For our replication, we chose to
keep BPA, rather than switching to a different public scientific
controversy with a higher profile in 2021, such as climate change,
police violence, voter fraud, or any of numerous aspects of the COVID-19
pandemic. All of these controversies are highly politically charged,
with prominent public experts and counterexperts (Goldenberg 2021,
100--101, ch.~6). While we anticipated some effects of political
partisanship in the BPA case, we felt it would be less likely to swamp
the values disclosure that was our primary interest.

After filling in the METI, participants provided demographic information
and other sections of the survey that are not examined here. Due to
researcher error, a question about the participants' values (whether
they prioritize economic growth or public health) was omitted in the
first wave of data collection; this question was asked in a followup
wave.

\hypertarget{replication-hypotheses-and-analytical-approach}{%
\subsection{Replication hypotheses and analytical
approach}\label{replication-hypotheses-and-analytical-approach}}

We identified 5 major findings from Elliott et al. (2017) for our
replication attempt:

\begin{description}
\tightlist
\item[H1. Modest correlation between values and ideology]
(a) Political liberals are more likely to prioritize public health over
economic growth, compared to political conservatives; but (b) a majority
of political conservatives prioritize public health.
\item[H2. Consumer risk sensitivity]
Scientists who find that a chemical harms human health are perceived as
more trustworthy than scientists who find that a chemical does not cause
harm.
\item[H3. Transparency penalty]
Scientists who disclose values are perceived as less trustworthy than
scientists who do not.
\item[H4. Shared values]
Given that the scientist discloses values, if the participant and the
scientist share the same values, the scientist is perceived as more
trustworthy than if the participant and scientist have discordant
values.
\item[H5. Variation in effects]
The magnitude of the effects 2-4 vary depending on whether the
participant prioritizes public health or economic growth.
\end{description}

Hypothesis H3, the transparency penalty, corresponds to the objection to
transparency: disclosing values undermines trust in science. However,
the shared values effect works in the opposite direction, counteracting
the transparency penalty.

Hypothesis 1 was analyzed using Spearman rank order correlation and
visually. Hypotheses 2-5 were analyzed using linear regression models as
a common framework, with a direct acyclic graph (DAG) constructed
\emph{a priori} to identify appropriate adjustments (covariates) for H4
and 5. Because Elliott et al. (2017) made their data publicly available,
and our analytical approaches differ somewhat from theirs, we conducted
parallel analyses of their data, and report these parallel results for
H2 and 3. Exploratory data analysis was used throughout to support data
validation and aid interpretation.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

Participants were recruited using the online survey platform Prolific,
and the survey was administered in a web browser using Qualtrics.
Prolific has an option to draw samples that are balanced to be
representative by age, binary gender, and a 5-category race variable
(taking values Asian, Black, Mixed, Other, and White) for US adults
({``Representative Samples FAQ''} 2022). A recent analysis finds that
Prolific produces substantially higher quality data than Amazon
Mechanical Turk for online survey studies, though three of the five
authors are affiliated with Prolific (Peer et al. 2021). Preliminary
power analysis recommended a sample of approximately 1,000 participants
to reliably detect non-interaction effects (H1-4).

After excluding participants who declined consent after opening the
survey or did not complete the survey, we had 988 participants in the
full analysis sample (\(M_{age}\) = 44-years-old, \(SD_{age}\) =
16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124,
Asian or Pacific Islander = 63, Hispanic = 33, American Indian or
Alaskan Native = 5, Mixed or Other = 51). Participants were randomly
assigned to condition, with 163 assigned to the No Disclosure + Causes
Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm
condition, 165 assigned to the Economic Growth + Causes Harm condition,
165 assigned to the Economic Growth + Does Not Cause Harm condition, 168
assigned to the Public Health + Causes Harm condition, and 162 assigned
to the Public Health + Does Not Cause Harm condition
(Table~\ref{tbl-condition}). Of the full 988 participants, 844
participants (85\%) responded to the followup question about their own
values (participant prioritizes economic growth or public health).
Consequently, subsamples for hypotheses 4 and 5 were substantially
smaller than the full analysis sample.

\hypertarget{tbl-condition}{}
\begin{table}
\caption{\label{tbl-condition}Assignment of participants to conditions }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lcc}
\toprule
\textbf{Characteristic} & \textbf{causes harm}, N = 496 & \textbf{does not cause harm}, N = 492\\
\midrule
disclosure/values &  & \\
\hspace{1em}no disclosure & 163 (33\%) & 165 (34\%)\\
\hspace{1em}public health & 168 (34\%) & 162 (33\%)\\
\hspace{1em}economic growth & 165 (33\%) & 165 (34\%)\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
\end{tabular}}
\end{table}

The study was approved by the UC Merced IRB on August 17, 2021, and data
collection ran October 18-20, 2021. As mentioned above, due to
researcher error a question about participants' values was not included
in the original survey. A followup survey asking this question of the
same group of participants was conducted December 8, 2021 through March
5, 2022.

\hypertarget{software-and-reproducibility}{%
\subsection{Software and
reproducibility}\label{software-and-reproducibility}}

Data cleaning and analysis was conducted in R version 4.1.2 (R Core Team
2021), with extensive use of the \texttt{tidyverse} suite of packages
version 1.3.1 (Wickham et al. 2019). Regression tables were generated
using the packages \texttt{gt} version 0.5.0 (Iannone, Cheng, and
Schloerke 2022) and \texttt{gtsummary} version 1.6.0 (Sjoberg et al.
2021).

Anonymized original data and reproducible code are available at
\url{https://github.com/dhicks/transparency}. Instructions in that
repository explain how to automatically reproduce our analysis.

\hypertarget{results}{%
\section{Results}\label{results}}

Critically, our data are unlikely to be representative by education
level and political ideology. In 2021, about 9\% of US adults 25 or over
had a less than high school education, and 38\% had a Bachelor's degree
or higher ({``CPS Historical Time Series Visualizations''} 2022, fig.
2). Only 1\% of our participants reported a less than high school
education, and 57\% reported a Bachelor's degree or higher. For
political ideology, the General Social Survey has consistently found
over several decades that about 30\% of US adults identify as liberal,
about 30\% identify as conservative, and about 40\% as moderate ({``GSS
Data Explorer''} 2022). Among our participants, liberals (574) heavily
outnumber conservatives (248; Figure~\ref{fig-part-values}). Both
overrepresentation of college graduates and underrepresentation of
conservatives (especially conservatives with strong anti-institutional
views) are known issues in public opinion polling (Kennedy et al. 2018).
In exploratory data analysis, we noted that there was essentially no
correlation between political ideology and perceived trustworthiness (in
the online supplement, see subsection H5-shared). This was surprising,
since general trust/distrust in science has become a partisan phenomenon
over the last few decades: data from the General Social Survey shows
increasing trust in science from liberals and decreasing trust from
conservatives (Gauchat 2012; Lee 2021), and many (though not all)
prominent public scientific controversies align with
liberal-conservative partisanship (Funk and Rainie 2015).

Because of these representation issues, insofar as some political
conservatives are both less likely to participate in studies on Prolific
(or at least in our particular study) and likely to perceive Spence as
less trustworthy, this will produce omitted variable bias for analyses
that require adjustment by political ideology. Analysis of our \emph{a
priori} DAG indicated that, for the hypotheses examined here, this
adjustment was not necessary in any case. When some adjustment was
necessary (H4-5), we included political ideology along with other
demographic variables in an alternative model specification as a
robustness check. In line with the DAG, in no case did these robustness
checks produce indications of omitted variable bias.

\hypertarget{h1-correlation-between-values-and-ideology}{%
\subsection{H1: Correlation between values and
ideology}\label{h1-correlation-between-values-and-ideology}}

We tested the hypothesis of a modest correlation between values and
ideology in two ways. First, to test (H1a) whether political liberals
are more likely to prioritize public health over economic growth
compared to conservatives, we conducted a Spearman's rank order
correlation. Results revealed a significant correlation in line with the
hypothesis, Spearman's \(\rho\) = -.47, \emph{p} \textless{} .001.
Political liberals were more likely than conservatives to value public
health over economic growth. To test (H1b) that a majority of political
conservatives priortize public health over economic growth we
cross-tabulated the data. Results revealed that, contrary to the
hypothesis, slightly more than half of the self-reported political
conservatives in our sample reported valuing economic growth (51.7\%)
over public health (48.3\%). See Figure~\ref{fig-part-values}.

\begin{figure}

{\centering \includegraphics{fig2_part_values.png}

}

\caption{\label{fig-part-values}Participant values by political
ideology. (A) Absolute counts, (B) shares within political ideology
categories. Panel (A) shows that our sample substantially
over-represents political liberals.}

\end{figure}

\hypertarget{h2-and-h3-consumer-risk-sensitivity-and-transparency-penalty}{%
\subsection{H2 and H3: Consumer risk sensitivity and transparency
penalty}\label{h2-and-h3-consumer-risk-sensitivity-and-transparency-penalty}}

Next, we tested the hypotheses that (H2) scientists who find a chemical
harms human health are perceived as more trustworthy than scientist who
find that a chemical does not cause harm and (H3) a scientist who
discloses values are perceived as less trustworthy than a scientist who
does not. For this analysis, we regressed participants' METI ratings
onto both the Conclusions and Disclosure experimental conditions. The
full model was significant, adj. \(R^2\) = .148, \emph{F}(2, 985) =
86.59, \emph{p} \textless{} .001 (see Table~\ref{tbl-bc}). Specifically,
results revealed that the conclusions reported by the scientist
predicted participants' perceived trustworthiness in line with our
hypothesis. Participants rated the scientist who reported that BPA does
not cause harm as less trustworthy (\(M_{sd} = 4.48_{1.26}\)) than the
scientist who reported that BPA causes harm (\(M_{sd} = 5.47_{1.12}\)),
\(\beta\) = -0.99, \emph{t}(985) = -13.09, \emph{p} \textless{} .001. By
contrast, the results do not provide evidence in favor of our hypothesis
that a scientist disclosing their values (\(M_{sd} = 4.94_{1.33}\)) are
perceived as less trustworthy than a scientist who does not disclose
values (\(M_{sd} = 5.05_{1.21}\)), \(\beta\) = -0.12, \emph{t}(985) =
-1.46, \emph{p} = .143.

\hypertarget{tbl-bc}{}
\begin{table}
\caption{\label{tbl-bc}Regression analysis for hypotheses H2 (consumer risk sensitivity) and H3
(disclosure effect) }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{EMAD}} & \multicolumn{3}{c}{\textbf{HL replication}} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
\textbf{Characteristic} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
(Intercept) & 6.0 & 5.7, 6.2 & <0.001 & 5.6 & 5.4, 5.7 & <0.001\\
conclusion &  &  &  &  &  & \\
\hspace{1em}conclusion[does not cause harm] & -1.1 & -1.3, -0.82 & <0.001 & -1.0 & -1.1, -0.84 & <0.001\\
disclosure &  &  &  &  &  & \\
\hspace{1em}disclosure[TRUE] & -0.52 & -0.78, -0.26 & <0.001 & -0.12 & -0.28, 0.04 & 0.14\\
\midrule
\addlinespace
R² & 0.152 &  &  & 0.150 &  & \\
No. Obs. & 498 &  &  & 988 &  & \\
Adjusted R² & 0.148 &  &  & 0.148 &  & \\
Statistic & 44.3 &  &  & 86.6 &  & \\
p-value & <0.001 &  &  & <0.001 &  & \\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} CI = Confidence Interval}\\
\end{tabular}}
\end{table}

\hypertarget{h4-shared-values}{%
\subsection{H4: Shared values}\label{h4-shared-values}}

Next, we tested the hypothesis that (H4) if the participant and
scientist share the same values, the scientist is perceived as more
trustworthy than if the participant and scientist do not share the same
values. For this and related analyses, we only included data from
participants who were assigned to either of the Disclosure conditions
and self-reported their own values, reducing the sample to 567.
Following this, we created a new Shared Values variable as a composite
of the participants' reported values and the scientist's values. For
this analysis, an a priori directed acyclic graphic (DAG) indicated that
a univariate regression of participants' METI rating onto Shared Values
would produce a biased estimate, that adjustments were required for both
the participant's and scientist's values, and that including demographic
variables would not effect this estimate (Figure~\ref{fig-shared-dag}).
While the univariate model was significant, adj. \(R^2\)= .029,
\emph{F}(1, 565) = 18.0, \emph{p} \textless{} .001
(Table~\ref{tbl-shared} model 1), after adjusting for scientist values
Shared Values did not emerge as a significant predictor of participants'
perceptions of the trustworthiness of the scientist
(Table~\ref{tbl-shared} models 2-4), indicating that the univariate
estimate was indeed biased. Rather, only the scientist's disclosed
values significantly predicted how trustworthy participants rated the
scientist, such that a scientist who disclosed valuing public health was
rated as more trustworthy than a scientist who disclosed valuing
economic growth. This scientist values effect was robust across
alternative model specifications, with an estimated effect of 0.6 (95\%
CI 0.4-0.8; Table~\ref{tbl-shared} models 2-5), again consistent with
the DAG.

\begin{figure}

{\centering \includegraphics{fig3_shared_values_dag.png}

}

\caption{\label{fig-shared-dag}Directed acyclic graph (DAG) for analysis
of the shared values effect. If the DAG is faithful, adjusting for
scientist values (\texttt{sci\_values}) and participant values
(\texttt{part\_values}) is sufficient for estimating the effect of
shared values on perceived trustworthiness (\texttt{METI}). In
particular, adjusting for demographics should not change the regression
coefficient on shared values.}

\end{figure}

\hypertarget{tbl-shared}{}
\begin{table}
\caption{\label{tbl-shared}Sequential regression analysis of hypothesis H4 shared values/scientist
values effects }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lccccccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{(1) univariate} & \multicolumn{3}{c}{(2) scientist values} & \multicolumn{3}{c}{(3) participant values} & \multicolumn{3}{c}{(4) demographics} & \multicolumn{3}{c}{(5) scientist values alone} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13} \cmidrule(l{3pt}r{3pt}){14-16}
\textbf{Characteristic} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
(Intercept) & 4.7 & 4.5, 4.9 & <0.001 & 4.6 & 4.4, 4.7 & <0.001 & 4.7 & 4.5, 5.0 & <0.001 & 4.5 & 3.7, 5.3 & <0.001 & 4.6 & 4.5, 4.8 & <0.001\\
shared values &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}shared\_values[TRUE] & 0.47 & 0.25, 0.68 & <0.001 & 0.11 & -0.15, 0.37 & 0.4 & 0.10 & -0.16, 0.36 & 0.5 & 0.07 & -0.21, 0.35 & 0.6 &  &  & \\
scientist values &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}sci\_values[public health] &  &  &  & 0.62 & 0.36, 0.88 & <0.001 & 0.63 & 0.37, 0.89 & <0.001 & 0.62 & 0.34, 0.90 & <0.001 & 0.65 & 0.45, 0.84 & <0.001\\
\addlinespace
participant values &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}part\_values[public health] &  &  &  &  &  &  & -0.18 & -0.44, 0.08 & 0.2 & -0.17 & -0.49, 0.15 & 0.3 &  &  & \\
age &  &  &  &  &  &  &  &  &  & 0.00 & -0.01, 0.01 & 0.7 &  &  & \\
rel. serv. attendance &  &  &  &  &  &  &  &  &  & 0.01 & -0.08, 0.10 & 0.9 &  &  & \\
political id. &  &  &  &  &  &  &  &  &  & -0.01 & -0.09, 0.06 & 0.8 &  &  & \\
\addlinespace
education &  &  &  &  &  &  &  &  &  & 0.05 & -0.18, 0.28 & 0.6 &  &  & \\
\midrule
R² & 0.031 &  &  & 0.067 &  &  & 0.070 &  &  & 0.143 &  &  & 0.060 &  & \\
No. Obs. & 567 &  &  & 567 &  &  & 567 &  &  & 538 &  &  & 660 &  & \\
Adjusted R² & 0.029 &  &  & 0.064 &  &  & 0.065 &  &  & 0.065 &  &  & 0.058 &  & \\
Statistic & 18.0 &  &  & 20.3 &  &  & 14.1 &  &  & 1.83 &  &  & 41.7 &  & \\
\addlinespace
p-value & <0.001 &  &  & <0.001 &  &  & <0.001 &  &  & 0.001 &  &  & <0.001 &  & \\
\bottomrule
\multicolumn{16}{l}{\rule{0pt}{1em}\textsuperscript{1} CI = Confidence Interval}\\
\end{tabular}}
\end{table}

\hypertarget{h5-variation-in-effects}{%
\subsection{H5: Variation in effects}\label{h5-variation-in-effects}}

We ran the following analyses to test our hypothesis that (H5) the
magnitude of the effects found for the tests of H2-H4 vary depending on
whether the participant prioritizes public health or economic growth.
Results are presented in
Table~\ref{tbl-eb}, Table~\ref{tbl-ec}, Table~\ref{tbl-sci-part}.

\hypertarget{consumer-risk-sensitivity}{%
\subsubsection{Consumer risk
sensitivity}\label{consumer-risk-sensitivity}}

To test whether the findings regarding consumer risk sensitivity vary as
a function of participants' values, we regressed participants' METI
ratings of the scientist in the stimuli onto the Conclusions condition,
Participants' Values variable, and the Conclusions by Participants'
values interaction term. The full model was significant, adj. \(R^2\) =
.173, \emph{F}(3, 840) = 59.95, \emph{p} \textless{} .001
(Table~\ref{tbl-eb} model 2). As with the earlier analysis, results
showed a main effect of the Conclusions condition, such that
participants rated the scientist who reported that BPA does not cause
harm as less trustworthy than the scientist who reported that BPA causes
harm, \(\beta\) = -0.75, \emph{t}(840) = -4.51, \emph{p} \textless{}
.001. However, this effect was qualified by a significant interaction
with participants' values, \(\beta\) = -0.41, \emph{t}(840) = -2.16,
\emph{p} = .031. Participants who prioritized public health and read
about a scientist who concluded that BPA causes harm rated the scientist
as more trustworthy (\(M_{sd} = 5.56_{1.09}\)) than participants with
the same values who read about a scientist who concluded that BPA does
not cause harm (\(M_{sd} = 4.40_{1.24}\);
Figure~\ref{fig-conclusion-part}).

\hypertarget{tbl-eb}{}
\begin{table}
\caption{\label{tbl-eb}Regression analysis of H5-consumer, interaction between participant
values and consumer risk sensitivity }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{base} & \multicolumn{3}{c}{interaction} & \multicolumn{3}{c}{demographics} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
\textbf{Characteristic} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
(Intercept) & 5.5 & 5.4, 5.6 & <0.001 & 5.4 & 5.2, 5.6 & <0.001 & 5.3 & 4.7, 5.9 & <0.001\\
conclusion &  &  &  &  &  &  &  &  & \\
\hspace{1em}conclusion[does not cause harm] & -1.0 & -1.1, -0.84 & <0.001 & -0.75 & -1.1, -0.43 & <0.001 & -0.73 & -1.1, -0.38 & <0.001\\
part\_values &  &  &  &  &  &  &  &  & \\
\hspace{1em}part\_values[public health] &  &  &  & 0.18 & -0.08, 0.43 & 0.2 & 0.20 & -0.09, 0.49 & 0.2\\
\addlinespace
conclusion * part\_values &  &  &  &  &  &  &  &  & \\
\hspace{1em}conclusion[does not cause harm] * part\_values[public health] &  &  &  & -0.41 & -0.79, -0.04 & 0.031 & -0.46 & -0.86, -0.07 & 0.022\\
age &  &  &  &  &  &  & 0.00 & 0.00, 0.01 & 0.4\\
religious\_serv &  &  &  &  &  &  & -0.01 & -0.08, 0.05 & 0.7\\
political\_ideology &  &  &  &  &  &  & -0.02 & -0.08, 0.03 & 0.4\\
\addlinespace
education &  &  &  &  &  &  & -0.04 & -0.21, 0.13 & 0.6\\
\midrule
R² & 0.148 &  &  & 0.176 &  &  & 0.236 &  & \\
No. Obs. & 988 &  &  & 844 &  &  & 801 &  & \\
Adjusted R² & 0.147 &  &  & 0.173 &  &  & 0.186 &  & \\
Statistic & 171 &  &  & 59.9 &  &  & 4.64 &  & \\
\addlinespace
p-value & <0.001 &  &  & <0.001 &  &  & <0.001 &  & \\
\bottomrule
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{1} CI = Confidence Interval}\\
\end{tabular}}
\end{table}

\begin{figure}

{\centering \includegraphics{fig4_conclusion_part.png}

}

\caption{\label{fig-conclusion-part}Data, group means, and 95\%
confidence intervals for H5-consumer, interaction of consumer risk
sensitivity and participant values. Panels correspond to participant
values.}

\end{figure}

\hypertarget{transparency-penalty}{%
\subsubsection{Transparency penalty}\label{transparency-penalty}}

To test whether the findings above about a hypothesized transparency
penalty may vary based on participants' values, we regressed
participants' METI ratings of the scientist onto the Disclosure
condition variable, Participants' Values variable, and the Disclosure by
Participants' Values interaction term. The full model was not
significant, adj. \(R^2\)\textless{} .001, \emph{F}(3, 840) = 1.19,
\emph{p} = .31 (Table~\ref{tbl-ec} model 2). As with the earlier
analysis, our results do not provide evidence for a transparency penalty
to the perceived trustworthiness of a scientist, either in general or
interacting with participants' own reported values, \(\beta\) = -0.23,
\emph{t}(840) = -1.07, \emph{p} = 0.286.

\hypertarget{tbl-ec}{}
\begin{table}
\caption{\label{tbl-ec}Regression analysis of H5-transparency, interaction between participant
values and transparency penalty }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{base} & \multicolumn{3}{c}{interaction} & \multicolumn{3}{c}{demographics} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
\textbf{Characteristic} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
(Intercept) & 5.1 & 4.9, 5.2 & <0.001 & 5.0 & 4.7, 5.3 & <0.001 & 4.9 & 4.3, 5.6 & <0.001\\
disclosure &  &  &  &  &  &  &  &  & \\
\hspace{1em}disclosure[TRUE] & -0.11 & -0.28, 0.06 & 0.2 & 0.04 & -0.32, 0.41 & 0.8 & 0.05 & -0.34, 0.43 & 0.8\\
part\_values &  &  &  &  &  &  &  &  & \\
\hspace{1em}part\_values[public health] &  &  &  & 0.07 & -0.26, 0.41 & 0.7 & 0.14 & -0.23, 0.51 & 0.5\\
\addlinespace
disclosure * part\_values &  &  &  &  &  &  &  &  & \\
\hspace{1em}disclosure[TRUE] * part\_values[public health] &  &  &  & -0.23 & -0.66, 0.19 & 0.3 & -0.29 & -0.74, 0.15 & 0.2\\
age &  &  &  &  &  &  & 0.00 & 0.00, 0.01 & 0.5\\
religious\_serv &  &  &  &  &  &  & 0.02 & -0.06, 0.09 & 0.7\\
political\_ideology &  &  &  &  &  &  & -0.02 & -0.08, 0.04 & 0.5\\
\addlinespace
education &  &  &  &  &  &  & -0.07 & -0.26, 0.11 & 0.4\\
\midrule
R² & 0.002 &  &  & 0.004 &  &  & 0.071 &  & \\
No. Obs. & 988 &  &  & 844 &  &  & 801 &  & \\
Adjusted R² & 0.001 &  &  & 0.001 &  &  & 0.009 &  & \\
Statistic & 1.60 &  &  & 1.19 &  &  & 1.15 &  & \\
\addlinespace
p-value & 0.2 &  &  & 0.3 &  &  & 0.2 &  & \\
\bottomrule
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{1} CI = Confidence Interval}\\
\end{tabular}}
\end{table}

\hypertarget{shared-values-and-scientist-values}{%
\subsubsection{Shared values and scientist
values}\label{shared-values-and-scientist-values}}

Without adjustments, Shared Values appears to have a substantial
interaction with Participant Values: Shared Values appears to increase
perceived trustworthiness for participants who value public health,
while decreasing perceived trustworthiness for participants who value
economic growth (Figure~\ref{fig-shared-part}). However, as indicated by
our analysis for a potential shared values effect, the estimated effects
for both shared values and participant values are biased if the model is
not adjusted for scientist values. (Figure~\ref{fig-shared-dag} and
Table~\ref{tbl-shared} show how this bias occurs. Shared Values is a
collider between Participant Values and Scientist Values; hence, if a
model specification includes Participant Values and Shared values but
not Scientist Values, there is an open path between Participant Values
and METI, resulting in a biased estimate for Participant Values.)
Because Shared Values is the logical biconditional of Participant Values
and Scientist Values, and the Shared Values \(\times\) Participant
Values interaction term is their logical conjunction, including all four
variables (Participant, Scientist, and Shared Values, along with the
interaction term) in a regression model creates perfect collinearity.

\begin{figure}

{\centering \includegraphics{fig5_shared_part.png}

}

\caption{\label{fig-shared-part}Data, group means, and 95\% confidence
intervals for H5-shared, interaction of shared values and participant
values. Panels correspond to participant values.}

\end{figure}

As we found above, rather than a shared values effect there appears to
be an effect of scientist values. Replotting the same data as
Figure~\ref{fig-shared-part} based on scientist values, rather than
shared values, suggests a more consistent effect across participant
values (Figure~\ref{fig-sci-part}). Therefore, we conducted an unplanned
post hoc analysis to test whether an effect of scientist values might
vary as a function of participants values. We regressed participants'
METI of the scientist in the stimulus onto the Scientist Values
variable, Participants' Values variable, and their interaction
(Table~\ref{tbl-sci-part}). This model was significant, adj. \(R^2\) =
0.065, \emph{F}(3, 563) = 14.15, \emph{p} \textless{} .001. The estimate
for the scientist values term was significant, \(\beta\) = 0.53,
\emph{t}(563) = 2.26, \emph{p} = 0.02, while the estimate for the
interaction term was not, \(\beta\) = 0.20, \emph{t}(563) = 0.75,
\emph{p} = 0.45. Estimates for both variables had large uncertainties,
with confidence intervals of about 1 (95\% confidence intervals,
scientist values: 0.07, 1.0; interaction term: -0.32, 0.72).

\begin{figure}

{\centering \includegraphics{fig6_sci_part.png}

}

\caption{\label{fig-sci-part}Data, group means, and 95\% confidence
intervals for a potential interaction of scientist values and
participant values. Panels correspond to participant values. This figure
is identical to Figure~\ref{fig-shared-part}, except the columns in the
economic growth panel have been switched.}

\end{figure}

\hypertarget{tbl-sci-part}{}
\begin{table}
\caption{\label{tbl-sci-part}Regression analysis of interaction between scientist values and
participant values }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lcccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{base} & \multicolumn{3}{c}{participant values} & \multicolumn{3}{c}{interaction} & \multicolumn{3}{c}{demographics} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
\textbf{Characteristic} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value} & \textbf{Beta} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
(Intercept) & 4.6 & 4.5, 4.8 & <0.001 & 4.7 & 4.5, 5.0 & <0.001 & 4.8 & 4.5, 5.1 & <0.001 & 4.5 & 3.7, 5.4 & <0.001\\
sci\_values &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}sci\_values[public health] & 0.65 & 0.45, 0.84 & <0.001 & 0.69 & 0.47, 0.90 & <0.001 & 0.53 & 0.07, 1.0 & 0.024 & 0.55 & 0.05, 1.1 & 0.031\\
part\_values &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}part\_values[public health] &  &  &  & -0.19 & -0.44, 0.07 & 0.2 & -0.28 & -0.63, 0.08 & 0.12 & -0.24 & -0.65, 0.17 & 0.3\\
\addlinespace
sci\_values * part\_values &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}sci\_values[public health] * part\_values[public health] &  &  &  &  &  &  & 0.20 & -0.32, 0.72 & 0.5 & 0.14 & -0.42, 0.71 & 0.6\\
age &  &  &  &  &  &  &  &  &  & 0.00 & -0.01, 0.01 & 0.7\\
religious\_serv &  &  &  &  &  &  &  &  &  & 0.01 & -0.08, 0.10 & 0.9\\
political\_ideology &  &  &  &  &  &  &  &  &  & -0.01 & -0.09, 0.06 & 0.8\\
\addlinespace
education &  &  &  &  &  &  &  &  &  & 0.05 & -0.18, 0.28 & 0.6\\
\midrule
R² & 0.060 &  &  & 0.069 &  &  & 0.070 &  &  & 0.143 &  & \\
No. Obs. & 660 &  &  & 567 &  &  & 567 &  &  & 538 &  & \\
Adjusted R² & 0.058 &  &  & 0.066 &  &  & 0.065 &  &  & 0.065 &  & \\
Statistic & 41.7 &  &  & 21.0 &  &  & 14.1 &  &  & 1.83 &  & \\
\addlinespace
p-value & <0.001 &  &  & <0.001 &  &  & <0.001 &  &  & 0.001 &  & \\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{1} CI = Confidence Interval}\\
\end{tabular}}
\end{table}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Table~\ref{tbl-summary} summarizes the results of our replication
attempts. We were able to successfully replicate the correlation between
political ideology and valuing public health and consumer risk
sensitivity (a scientist who concludes BPA is harmful is perceived as
more trustworthy), and in addition found evidence of an interaction
between consumer risk sensitivity and whether the participant values
public health (participants who value public health have more extreme
reactions to the scientist's conclusion).

We did not find evidence for a transparency penalty or shared values
effect. And we found evidence of a scientist values effect (a scientist
who discloses valuing public health is perceived as more trustworthy),
which was not identified by Elliott et al. (2017).

\hypertarget{tbl-summary}{}
\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2471}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5412}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2118}}@{}}
\caption{\label{tbl-summary}Summary of replication results. 95\%
confidence intervals. \(^\star\)The shared values \(\times\) participant
values interaction specification had perfect multicollinearity and could
not be fit. \(^\dagger\)For these unplanned analyses, ``yes'' indicates
that we found evidence supporting the stated hypothesis, and ``no''
indicates that we did not find such evidence. \(^\ddagger\)Our analysis
method found no evidence of a shared values effect in the Elliott et al.
(2017) data.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Hypothesis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Replicated? (CI)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Hypothesis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Replicated? (CI)
\end{minipage} \\
\midrule()
\endhead
H1a & Liberals more likely to prioritize public health than
conservatives & yes \\
H1b & Majority of conservatives prioritize public health & no \\
H2 & Scientist who finds chemical is harmful perceived as more
trustworthy & yes (-1.1, -0.84) \\
H3 & Scientist who discloses values perceived as less trustworthy & no
(-0.28, 0.04) \\
H4 & Given scientist discloses values, scientist who shares values with
participant perceived as more trustworthy & no\(^\ddagger\) (-0.16,
0.36) \\
Unplanned & Given scientist discloses values, scientist who values
public health perceived as more trustworthy & yes\(^\dagger\) (0.37,
0.89) \\
H5-consumer & Conclusion (H2) \(\times\) participant values & yes
(-0.79, -0.04) \\
H5-transparenc & Disclosure (H3) \(\times\) participant values & no
(-0.66, 0.19) \\
H5-shared & Shared values (H4) \(\times\) participant values &
\(^\star\) \\
H5-sci. values & Scientist values \(\times\) participant values &
no\(^\dagger\) (-0.32, 0.72) \\
\bottomrule()
\end{longtable}

Elliott et al. (2017) claim evidence for a shared values effect by
comparing values disclosure vs.~no disclosure (that is, estimating a
transparency effect) in quasi-independent subsamples across a 2
\(\times\) 2 \(\times\) 2 design (participant values \(\times\)
scientist conclusion \(\times\) scientist values; Elliott et al. (2017)
fig.~1). This approach is difficult to interpret; their reasoning seems
to be that this estimate is statistically significant in 4/8 cells, and
in 3/4 of these cells (1 where respondent values economic growth) the
participant and scientist share values. But these analyses are
underpowered (for example, sample sizes are well below 100 for some of
the cells in which participants value economic growth) and the overall
approach cannot distinguish a shared values effect from other potential
effects (scientist values, participant values). Applying our regression
analysis approach to the data published by Elliott et al. (2017), the
estimate for shared values was not statistically significant, \(\beta\)
= 0.29, \emph{t}(335) = 1.75, \emph{p} = 0.08; while the estimate for
scientist values was statistically significant, \(\beta\) = 0.41,
\emph{t}(335) = 2.48, \emph{p} = 0.014. It seems likely that claims of a
shared values effect could have been the result of a less appropriate
analytic approach given the nature of the data.

The disagreement over a transparency penalty is more difficult to
explain. Using our approach and the data from Elliott et al. (2017), the
estimate for a values disclosure was statistically significant
(Table~\ref{tbl-bc} model 1). One possible explanation is that, over the
course of the COVID-19 pandemic, members of the general public have
become used to ``scientists'' (including both bench researchers and
public health officials) making claims about the importance of
protecting public health. A values disclosure that might have been
regarded as violating the value-free ideal, pre-pandemic, might now be
seen as routine.

The scientist values effect --- a scientist who discloses valuing public
health is perceived as more trustworthy --- might be interpreted as
supporting some of the claims of the ``aims approach'' in philosophy of
science. This approach argues that scientific fields often have both
social or practical aims, along with epistemic aims such as the pursuit
of truth (Elliott and McKaughan 2014; Intemann 2015; Potochnik 2017;
Fernández Pinto and Hicks 2019; Hicks 2022). For example, the field of
public health might have the practical aim of promoting the health of
the public. Where certain values are seen as conflicting with these
aims, it is morally wrong for a scientist to promote those values.
Rather than violating the value-free ideal, a scientist who discloses
valuing public health might be seen as following the norms of the field
of public health.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

As noted above, a major limitation of the current study is that the
sample substantially underrepresents political conservatives. Our
\emph{a priori} DAG analysis and empirical robustness checks do not
indicate any bias in our reported estimates due to this representation
problem. However, our estimates could still be biased if any of the
effects examined here have interactions with political ideology.

Another limitation is that our sample only considers US adults. It does
not include residents of any other country, and does not track where our
participants might live within the US. Public scientific controversies
often have different dynamics both across national borders and within
different regions of geographically large countries (Miles and Frewer
2003; Howe et al. 2015; Mildenberger et al. 2016; Mildenberger et al.
2017; Sturgis, Brunton-Smith, and Jackson 2021).

A third limitation is that the stimulus only considers a single
controversy, over the safety of the chemical bisphenol A (BPA). Public
attention to this controversy peaked around 2009. While it remains
unsettled in terms of both science and policy, it is much less socially
and politically salient than controversies over topics such as climate
change or vaccination. This limitation affects our ability to speculate
about the generalizability of our findings. That is, the effects we
reported may only apply narrowly to a subset of sociopolitically
controversial scientific topics that have not achieved the degree of
salience or persistence in sociopolitical discourse for as long as
topics like climate change or vaccinations.

\hypertarget{directions-for-future-work}{%
\subsection{Directions for future
work}\label{directions-for-future-work}}

The design from Elliott et al. (2017) assesses public perceptions of the
value-free ideal indirectly, by probing whether violations of this ideal
would lead to decreased perceived trustworthiness. But do members of the
general public accept the value-free ideal? Concurrently with this
replication study, we also asked participants directly for their views
on the value-free ideal and related issues and arguments from the
philosophy of science literature. A manuscript discussing this effort to
develop a ``values in science scale'' is currently under preparation.

We suggest three directions for future work in this area, two in the
experimental stimulus and one in the endpoint or outcome measured.

The stimulus developed by Elliott et al. (2017) involves a single
(fictional) scientist. But public scientific controversies often feature
conflicting claims made by contesting experts and counterexperts. For
example, as the Omicron wave of the COVID-19 pandemic waned in the US in
spring 2022, various physicians, public health experts, science
journalists, and government officials made conflicting claims about the
effective severity of Omicron, the relative effectiveness of masks in
highly vaccinated populations, and the need to ``return to normal''
(Adler-Bell 2022; Khullar 2022; Yong 2022). For members of the general
public, the question was not whether to trust the claims of a given
expert, but instead which experts to trust. Publics might perceive both
expert A and expert B to be highly trustworthy, but would need to make a
decision about who to trust if these experts are making conflicting
claims. It would be straightforward to modify the single-scientist
stimulus to cover this kind of ``dueling experts'' scenario. To modify
METI, participants might be asked which of the two experts they would
consider more competent, ethical, honest, etc. (Public scientific
controversies can include other prominent actors with no relevant
expertise, such as political journalists, elected officials, or social
media conspiracy theorists. Such individuals might nonetheless be
treated by publics as trusted sources for factual claims. Understanding
why many people might trust a figure like Alex Jones over someone like
Anthony Fauci would likely require prior work disentangling
trustworthiness, expertise, and institutional standing.)

Brown (2022) emphasized a distinction between individuals, groups, and
institutions, as both trustors and trustees, and argued that much of the
philosophical and empirical research on trust in science has focused on
either individual-individual trust (as in Elliott et al. (2017)) or
individual-group trust (as in the General Social Survey, which asks
individual respondents about their confidence in ``the scientific
community''). In future experiments, ``Dr.~Riley Spence'' could be
represented with an institutional affiliation, such as ``Dr.~Riley
Spence of the Environmental Protection Agency'' or ``Dr.~Riley Spence of
the American Petroleum Institute'' Or claims could be attributed
directly to institutions, such as ``Environmental Protection Agency
scientists'' or ``a report published by the American Petroleum
Institute.''

As noted above, ethicists make a distinction between trust and
trustworthiness. Instruments such as METI assess trustworthiness:
whether a speaker is perceived to have qualities that would make it
appropriate to trust them. Trust itself is closer to behavior than
attitudes, and would probably be better measured (in online survey
experiments and similar designs) by asking whether participants accept
claims made by speakers or support policy positions endorsed by
speakers. In addition, research on the climate controversy suggests that
acceptance of scientific claims might underpredict policy support. Since
2008, public opinion studies by the Yale Program on Climate Change
Communication have found that only about 50-60\% of the US public
understands that there's a scientific consensus on climate change, that
this understanding is politically polarized (conservatives are
substantially less likely to recognize the consensus), but also that
some climate policies enjoy majority support even among conservatives
(Leiserowitz et al. 2022). Putting these points together, future
research could ask participants whether they would support, for example,
increased regulation of BPA, either instead of or along with assessing
the perceived trustworthiness of the scientific expert.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

For feedback on this project, thanks to Kevin Elliott and participants
at the Values in Science conference and Political Philosophy and Values
in Medicine, Science, and Technology 2022 conference.

\hypertarget{funding}{%
\section{Funding}\label{funding}}

Funding for this project was provided by the University of California,
Merced.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Adler-BellPandemicInterpreter2022}{}}%
Adler-Bell, Sam. 2022. {``The Pandemic Interpreter.''} Intelligencer.
February 24, 2022.
\url{https://nymag.com/intelligencer/2022/02/david-leonhardt-the-pandemic-interpreter.html}.

\leavevmode\vadjust pre{\hypertarget{ref-BaierTrustAntitrust1986}{}}%
Baier, Annette. 1986. {``Trust and Antitrust.''} \emph{Ethics} 96 (2):
231--60. \url{http://www.jstor.org/stable/2381376}.

\leavevmode\vadjust pre{\hypertarget{ref-BrownTrustExpertiseScientific2022}{}}%
Brown, Matthew. 2022. {``Trust, Expertise and Scientific Authority in
Democracy.''} Michigan State University, February 20.
\url{https://www.youtube.com/watch?v=y3XeP6e646g}.

\leavevmode\vadjust pre{\hypertarget{ref-CPSHistoricalTime2022}{}}%
{``CPS Historical Time Series Visualizations.''} 2022. United States
Census Bureau. March 12, 2022.
\url{https://web.archive.org/web/20220312234236/https://www.census.gov/library/visualizations/time-series/demo/cps-historical-time-series.html}.

\leavevmode\vadjust pre{\hypertarget{ref-CranorSocialBenefitsExpedited1995}{}}%
Cranor, Carl F. 1995. {``The Social Benefits of Expedited Risk
Assessments.''} \emph{Risk Analysis} 15 (3): 353--58.
\url{https://doi.org/10.1111/j.1539-6924.1995.tb00328.x}.

\leavevmode\vadjust pre{\hypertarget{ref-DouglasSciencePolicyValuefree2009}{}}%
Douglas, Heather E. 2009. \emph{Science, Policy, and the Value-Free
Ideal}. Pittsburgh, Pa: University of Pittsburgh Press.

\leavevmode\vadjust pre{\hypertarget{ref-ElliottTapestryValuesIntroduction2017}{}}%
Elliott, Kevin C. 2017. \emph{A Tapestry of Values: An Introduction to
Values in Science}. Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-ElliottValuesEnvironmentalResearch2017}{}}%
Elliott, Kevin C., Aaron M. McCright, Summer Allen, and Thomas Dietz.
2017. {``Values in Environmental Research: Citizens' Views of Scientists
Who Acknowledge Values.''} \emph{PLOS ONE} 12 (10): e0186049.
\url{https://doi.org/10.1371/journal.pone.0186049}.

\leavevmode\vadjust pre{\hypertarget{ref-ElliottNonepistemicValuesMultiple2014}{}}%
Elliott, Kevin C., and Daniel~J. McKaughan. 2014. {``Nonepistemic Values
and the Multiple Goals of Science.''} \emph{Philosophy of Science} 81
(1): 1--21. \url{https://doi.org/10.1086/674345}.

\leavevmode\vadjust pre{\hypertarget{ref-ElliottSciencePolicyTransparency2014}{}}%
Elliott, Kevin C., and David B. Resnik. 2014. {``Science, Policy, and
the Transparency of Values.''} \emph{Environmental Health Perspectives},
March. \url{https://doi.org/10.1289/ehp.1408107}.

\leavevmode\vadjust pre{\hypertarget{ref-FernandezPintoLegitimizingValuesRegulatory2019}{}}%
Fernández Pinto, Manuela, and Daniel J. Hicks. 2019. {``Legitimizing
Values in Regulatory Science.''} \emph{Environmental Health
Perspectives} 127 (3): 035001. \url{https://doi.org/10.1289/EHP3317}.

\leavevmode\vadjust pre{\hypertarget{ref-FunkAmericansPoliticsScience2015}{}}%
Funk, Cary, and Lee Rainie. 2015. {``Americans, Politics and Science
Issues.''} Pew Research Center.
\url{http://www.pewinternet.org/2015/07/01/americans-politics-and-science-issues/}.

\leavevmode\vadjust pre{\hypertarget{ref-GauchatPoliticizationSciencePublic2012}{}}%
Gauchat, Gordon. 2012. {``Politicization of Science in the Public Sphere
A Study of Public Trust in the United States, 1974 to 2010.''}
\emph{American Sociological Review} 77 (2): 167--87.
\url{https://doi.org/10.1177/0003122412438225}.

\leavevmode\vadjust pre{\hypertarget{ref-GoldenbergVaccineHesitancyPublic2021}{}}%
Goldenberg, Maya J. 2021. \emph{Vaccine Hesitancy: Public Trust,
Expertise, and the War on Science}. University of Pittsburgh Press.

\leavevmode\vadjust pre{\hypertarget{ref-GSSDataExplorer2022}{}}%
{``GSS Data Explorer.''} 2022. NORC at the University of Chicago. June
29, 2022.
\url{https://web.archive.org/web/20220629231111/https://gssdataexplorer.norc.org/variables/178/vshow}.

\leavevmode\vadjust pre{\hypertarget{ref-HendriksMeasuringLaypeopleTrust2015}{}}%
Hendriks, Friederike, Dorothe Kienhues, and Rainer Bromme. 2015.
{``Measuring Laypeople's Trust in Experts in a Digital Age: The Muenster
Epistemic Trustworthiness Inventory (METI).''} \emph{PLOS ONE} 10 (10):
e0139309. \url{https://doi.org/10.1371/journal.pone.0139309}.

\leavevmode\vadjust pre{\hypertarget{ref-HicksNewDirectionScience2014}{}}%
Hicks, Daniel J. 2014. {``A New Direction for Science and Values.''}
\emph{Synthese} 191 (14): 3271--95.
\url{https://doi.org/10.1007/s11229-014-0447-9}.

\leavevmode\vadjust pre{\hypertarget{ref-HicksWhenVirtuesAre2022}{}}%
---------. 2022. {``When Virtues are Vices: {`Anti-Science'} Epistemic
Values in Environmental Politics.''} \emph{Philosophy, Theory, and
Practice in Biology} 14 (0). \url{https://doi.org/10.3998/.2629}.

\leavevmode\vadjust pre{\hypertarget{ref-HoweGeographicVariationOpinions2015}{}}%
Howe, Peter D., Matto Mildenberger, Jennifer R. Marlon, and Anthony
Leiserowitz. 2015. {``Geographic Variation in Opinions on Climate Change
at State and Local Scales in the USA.''} \emph{Nature Climate Change} 5
(6): 596--603. \url{https://doi.org/10.1038/nclimate2583}.

\leavevmode\vadjust pre{\hypertarget{ref-IannoneGtEasilyCreate2022}{}}%
Iannone, Richard, Joe Cheng, and Barret Schloerke. 2022. {``Gt: Easily
Create Presentation-Ready Display Tables.''}
\url{https://CRAN.R-project.org/package=gt}.

\leavevmode\vadjust pre{\hypertarget{ref-IntemannDistinguishingLegitimateIllegitimate2015}{}}%
Intemann, Kristen. 2015. {``Distinguishing Between Legitimate and
Illegitimate Values in Climate Modeling.''} \emph{European Journal for
Philosophy of Science} 5 (2): 217--32.
\url{https://doi.org/10.1007/s13194-014-0105-6}.

\leavevmode\vadjust pre{\hypertarget{ref-JohnEpistemicTrustEthics2017}{}}%
John, Stephen. 2017. {``Epistemic Trust and the Ethics of Science
Communication: Against Transparency, Openness, Sincerity and Honesty.''}
\emph{Social Epistemology} 32 (2): 75--87.
\url{https://doi.org/10.1080/02691728.2017.1410864}.

\leavevmode\vadjust pre{\hypertarget{ref-KennedyEvaluation2016Election2018}{}}%
Kennedy, Courtney, Mark Blumenthal, Scott Clement, Joshua D Clinton,
Claire Durand, Charles Franklin, Kyley McGeeney, et al. 2018. {``An
Evaluation of the 2016 Election Polls in the United States.''}
\emph{Public Opinion Quarterly} 82 (1): 1--33.
\url{https://doi.org/10.1093/poq/nfx047}.

\leavevmode\vadjust pre{\hypertarget{ref-KhullarWillCoronavirusPandemic2022}{}}%
Khullar, Dhruv. 2022. {``Will the Coronavirus Pandemic Ever End?''}
\emph{The New Yorker}, May 23, 2022.
\url{https://www.newyorker.com/news/daily-comment/will-the-coronavirus-pandemic-ever-end}.

\leavevmode\vadjust pre{\hypertarget{ref-KovakaClimateChangeDenial2021}{}}%
Kovaka, Karen. 2021. {``Climate Change Denial and Beliefs about
Science.''} \emph{Synthese} 198 (3): 2355--74.
\url{https://doi.org/10.1007/s11229-019-02210-z}.

\leavevmode\vadjust pre{\hypertarget{ref-LeePartyPolarizationTrust2021}{}}%
Lee, John J. 2021. {``Party Polarization and Trust in Science: What
about Democrats?''} \emph{Socius} 7 (January): 23780231211010101.
\url{https://doi.org/10.1177/23780231211010101}.

\leavevmode\vadjust pre{\hypertarget{ref-LeiserowitzPoliticsGlobalWarming2022}{}}%
Leiserowitz, Anthony, Edward Maibach, Seth Rosenthal, and John Kotcher.
2022. {``Politics \& Global Warming, April 2022.''} Yale Program on
Climate Change Communication.
\url{https://climatecommunication.yale.edu/publications/politics-global-warming-april-2022/}.

\leavevmode\vadjust pre{\hypertarget{ref-McKaughanBacktrackingEthicsFraming2013}{}}%
McKaughan, Daniel J., and Kevin C. Elliott. 2013. {``Backtracking and
the Ethics of Framing: Lessons from Voles and Vasopressin.''}
\emph{Accountability in Research} 20 (3): 206--26.
\url{https://doi.org/10.1080/08989621.2013.788384}.

\leavevmode\vadjust pre{\hypertarget{ref-MildenbergerDistributionClimateChange2016}{}}%
Mildenberger, Matto, Peter Howe, Erick Lachapelle, Leah Stokes, Jennifer
Marlon, and Timothy Gravelle. 2016. {``The Distribution of Climate
Change Public Opinion in Canada.''} \emph{PLOS ONE} 11 (8): e0159774.
\url{https://doi.org/10.1371/journal.pone.0159774}.

\leavevmode\vadjust pre{\hypertarget{ref-MildenbergerSpatialDistributionRepublican2017}{}}%
Mildenberger, Matto, Jennifer R. Marlon, Peter D. Howe, and Anthony
Leiserowitz. 2017. {``The Spatial Distribution of Republican and
Democratic Climate Opinions at State and Local Scales.''} \emph{Climatic
Change}, November, 1--10.
\url{https://doi.org/10.1007/s10584-017-2103-0}.

\leavevmode\vadjust pre{\hypertarget{ref-MilesPublicPerceptionScientific2003}{}}%
Miles, Susan, and Lynn J. Frewer. 2003. {``Public Perception of
Scientific Uncertainty in Relation to Food Hazards.''} \emph{Journal of
Risk Research} 6 (3): 267--83.
\url{https://doi.org/10.1080/1366987032000088883}.

\leavevmode\vadjust pre{\hypertarget{ref-PeerDataQualityPlatforms2021}{}}%
Peer, Eyal, David Rothschild, Andrew Gordon, Zak Evernden, and Ekaterina
Damer. 2021. {``Data Quality of Platforms and Panels for Online
Behavioral Research.''} \emph{Behavior Research Methods}, September.
\url{https://doi.org/10.3758/s13428-021-01694-3}.

\leavevmode\vadjust pre{\hypertarget{ref-PotochnikIdealizationAimsScience2017}{}}%
Potochnik, Angela. 2017. \emph{Idealization and the Aims of Science}.
Chicago and London: University of Chicago Press.

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeamLanguageEnvironmentStatistical2021}{}}%
R Core Team. 2021. {``R: A Language and Environment for Statistical
Computing.''} R Foundation for Statistical Computing.
\url{https://www.r-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-RepresentativeSamplesFAQ2022}{}}%
{``Representative Samples FAQ.''} 2022. Prolific. March 13, 2022.
\url{https://web.archive.org/web/20220313050153/https://researcher-help.prolific.co/hc/en-gb/articles/360019238413-Representative-samples-FAQ}.

\leavevmode\vadjust pre{\hypertarget{ref-SjobergReproducibleSummaryTables2021}{}}%
Sjoberg, Daniel D., Karissa Whiting, Michael Curry, Jessica A. Lavery,
and Joseph Larmarange. 2021. {``Reproducible Summary Tables with the
Gtsummary Package.''} \emph{The R Journal} 13 (1): 570--80.
\url{https://journal.r-project.org/archive/2021/RJ-2021-053/index.html}.

\leavevmode\vadjust pre{\hypertarget{ref-SturgisTrustScienceSocial2021}{}}%
Sturgis, Patrick, Ian Brunton-Smith, and Jonathan Jackson. 2021.
{``Trust in Science, Social Consensus and Vaccine Confidence.''}
\emph{Nature Human Behaviour}, May, 1--7.
\url{https://doi.org/10.1038/s41562-021-01115-7}.

\leavevmode\vadjust pre{\hypertarget{ref-WickhamWelcomeTidyverse2019}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software} 4
(43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-YongHowDidThis2022}{}}%
Yong, Ed. 2022. {``How Did This Many Deaths Become Normal?''} The
Atlantic. March 8, 2022.
\url{https://www.theatlantic.com/health/archive/2022/03/covid-us-death-rate/626972/}.

\end{CSLReferences}

\end{document}

### 2022-10-11 paper/paper.Qmd
edit paper + meti subscale correlation
made revisions to Paper.qmd and added code to assess the correlations between the METI subscales in the 02_analysis script

@@
 @ElliottValuesEnvironmentalResearch2017 used an ad hoc scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.
+Even though the ad hoc scale created by Elliot and colleagues had acceptable internal reliability, the METI was developed and psychometrically validated to measure the perceived trustworthiness of experts.
 On the METI, participants rate the target scientist (the fictional Dr.
@@
 The METI scale captures perceived trustworthiness of a target along three dimensions: competence, benevolence, and integrity.
-However, the three dimensions were strongly correlated in our sample.
+However, the three dimensions were strongly correlated in our sample (Pearson r values ranging from .75 to .92).
 To avoid issues of multicollinearity in our analyses, we averaged participant scores into a single composite measure.
@@
 
-After filling in the METI, participants provided demographic information and other sections of the survey that are not examined here.
+After filling in the METI, participants provided demographic information, including self-identifying their political ideology, and other sections of the survey that are not examined here.
 Due to researcher error, a question about the participants' values (whether they prioritize economic growth or public health) was omitted in the first wave of data collection; this question was asked in a followup wave.
@@
 H5. Variation in effects
-~ The magnitude of the effects 2-4 vary depending on whether the participant prioritizes public health or economic growth.
+~ The magnitude of the effects for hypotheses 2-4 vary depending on whether the participant prioritizes public health or economic growth.
 
@@
 
-Hypothesis 1 was analyzed using Spearman rank order correlation and visually.
+Hypothesis 1 was analyzed using Spearman rank order correlation to test H1a and a visual inspection of descriptive statistics to test H1b.
 Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed *a priori* to identify appropriate adjustments (covariates) for H4 and 5.
@@
 Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition (@tbl-condition).
+Due to researcher error a question about participants' values was not included in the original survey.
 Of the full 988 participants, 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).
@@
 The study was approved by the UC Merced IRB on August 17, 2021, and data collection ran October 18-20, 2021.
-As mentioned above, due to researcher error a question about participants' values was not included in the original survey.
-A followup survey asking this question of the same group of participants was conducted December 8, 2021 through March 5, 2022.
+The followup survey asking the initial sample of participants about their own values regarding economic growth and public health was conducted December 8, 2021 through March 5, 2022.
 

### 2022-10-11 scripts/02_analysis.R

@@ -18,6 +18,12 @@ data_dir = here("data")
 #read the data in
 df <- read.csv(here(data_dir, 'data.csv'))
 
+#correlations of the METI subscales
+meti_cor <- cor(df |> select(meti_competence,
+                             meti_integrity,
+                             meti_benevolence), use = "complete.obs") |> 
+    formatC(digits=2,format="f")
+meti_cor
 
 ## Values & Ideology Hypotheses ----
 #have to subset out the "other" and "prefer not to say" political ideology responses for the correlation

### 2022-10-12 paper/paper.Qmd
finished revisions per reviewers

@@
 [@HicksNewDirectionScience2014]  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research [@ElliottSciencePolicyTransparency2014; @McKaughanBacktrackingEthicsFraming2013].
-But some philosophers have objected to this transparency proposal, arguing that it might actually undermine trust in science.
-According to this objection, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021].
+But some philosophers have objected to this transparency proposal, arguing that it might undermine trust in science.
+
+Specifically, ethicists make a distinction between trust and trustworthiness [@BaierTrustAntitrust1986].
+Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.
+Trustworthiness is an assessment, of whether or not that trust is appropriate.
+According to the objection to the ideal of transparency, transparency would not change how trustworthy science actually is — the science is conducted in the same way — but might reduce public perceptions of trustworthiness — as members of the public see how the scientific sausage is made.
+For instance, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021].
 
@@
 
-Ethicists make a key distinction between trust and trustworthiness [@BaierTrustAntitrust1986].
-Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.
-Trustworthiness is an assessment, of whether or not that trust is appropriate.
-In the experiment, participants might trust Dr.
-Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa; though we would typically expect these two to go together.
-As its name indicates, METI assesses perceived trustworthiness, not trust (such as accepting Dr.
+As its name indicates, METI assesses perceived trustworthiness rather than trust (such as accepting Dr.
 Spence's conclusion about BPA).
+It is conceivable that participants might trust Dr.
+Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa.
+But we would typically trust and trustworthiness to go together.
 
@@
 Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$ = 0.29, *t*(335) = 1.75, *p* = 0.08; while the estimate for scientist values was statistically significant, $\beta$ = 0.41, *t*(335) = 2.48, *p* = 0.014.
-It seems likely that claims of a shared values effect could have been the result of a less appropriate analytic approach given the nature of the data.
+It seems likely that claims of a shared values effect could have been an artifact of a less appropriate analytic approach given the nature of the data.
 
@@
 Using our approach and the data from @ElliottValuesEnvironmentalResearch2017, the estimate for a values disclosure was statistically significant (@tbl-bc model 1).
-One possible explanation is that, over the course of the COVID-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.
+One possible explanation is that, over the course of the COVID-19 pandemic, members of the general public have become used to "scientists" (broadly including both bench researchers and public health officials) making claims about the importance of protecting public health.
 A values disclosure that might have been regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.
@@
 Rather than violating the value-free ideal, a scientist who discloses valuing public health might be seen as following the norms of the field of public health.
+
+All together, regarding the philosophical debates over values in science, the evidence from this replication study suggests that the effects of transparency on perceived trustworthiness might be context-dependent.
+Transparency about the influence of might be detrimental in some cases, beneficial in others, and neutral in still others.
+Notably, both of the papers cited as objecting to transparency [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021] focus on climate change, while many of Elliott's examples come from toxicology or environmental public health.
+The critics might be right that transparency has undermined trust in the climate context, while Elliott might be right that transparency can be neutral or even beneficial in some public health contexts.
+Because philosophers of science tend to work with a small number of rich case studies, methods such as qualitative comparative analysis [QCA\; @RaginComparativeMethodMoving2013] would be worth exploring.
+
+An obvious difference between climate and BPA is the public prominence and political polarization of the two issues.
+For decades, the fossil fuels industry has used a range of strategies to successfully attack the perceived trustworthiness of climate scientists, especially in the minds of conservative publics [@OreskesMerchantsDoubtHow2011; @SupranAssessingExxonMobilClimate2017].
+While the chemical industry has also had a substantial influence on the environmental safety regulatory system [@MichaelsDoubtTheirProduct2008; @VogelItSafeBPA2013], controversies over BPA have received much less public attention.
+A LexiNexis search indicates that uses of "bisphenol" in US newspapers rose and then almost immediately fell around 2008, shortly after a US consensus conference concluded that "BPA at concentrations found in the human body is associated with organizational changes" in a number of major physiological systems [@vomSaalChapelHillBisphenol2007].
+
+In more prominent and/or polarized cases than BPA, a number of other psychological mechanisms might be activated.
+Motivated reasoning or identity-protective cognition [@KahanPolarizingImpactScience2012] might cause members of the general public to rate a scientist as more or less trustworthy insofar as the scientist agrees with the respondent's "side" of the debate, regardless of whether and which values are disclosed.
+That is, the scientist conclusion might have a larger effect than scientist values.
+Alternatively, respondents might use values disclosures to identify which scientists are on "their side," and so scientist values would have a larger effect than scientist conclusion.
+Careful attention to experimental design will be needed to separate these potential multiway interactions between conclusions, disclosures, and public prominence.
 

### 2022-10-12 paper/paper.tex
finished revisions per reviewers

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 Elliott 2013). But some philosophers have objected to this transparency
-proposal, arguing that it might actually undermine trust in science.
-According to this objection, members of the public generally accept the
-ideal of value-free science; values disclosures would violate this
+proposal, arguing that it might undermine trust in science.
+
+Specifically, ethicists make a distinction between trust and
+trustworthiness (Baier 1986). Trust is a verb: placing one's trust in
+someone else, with respect to some activity or domain. Trustworthiness
+is an assessment, of whether or not that trust is appropriate. According
+to the objection to the ideal of transparency, transparency would not
+change how trustworthy science actually is --- the science is conducted
+in the same way --- but might reduce public perceptions of
+trustworthiness --- as members of the public see how the scientific
+sausage is made. For instance, members of the public generally accept
+the ideal of value-free science; values disclosures would violate this
 ideal, making scientists (incorrectly) appear biased and untrustworthy
@@
 Bromme 2015), which substantially overlaps but is not identical to the
-Elliott et al. (2017) scale. On the METI, participants rate the target
+Elliott et al. (2017) scale. Even though the ad hoc scale created by
+Elliot and colleagues had acceptable internal reliability, the METI was
+developed and psychometrically validated to measure the perceived
+trustworthiness of experts. On the METI, participants rate the target
 scientist (the fictional Dr.~Spence) on a 1-7 scale for 14 semantic
@@
 dimensions: competence, benevolence, and integrity. However, the three
-dimensions were strongly correlated in our sample. To avoid issues of
-multicollinearity in our analyses, we averaged participant scores into a
-single composite measure. In both Elliott et al. (2017) and our
-replication analysis, all items were averaged together to form a 1-7
-composite measure of trustworthiness. To aid interpretation, in the
-analysis the direction of the scale has been set so that increased
-values correspond to increased perceived trustworthiness.
-
-Ethicists make a key distinction between trust and trustworthiness
-(Baier 1986). Trust is a verb: placing one's trust in someone else, with
-respect to some activity or domain. Trustworthiness is an assessment, of
-whether or not that trust is appropriate. In the experiment,
-participants might trust Dr.~Spence's conclusion even if they judge
-Spence to be untrustworthy, or vice versa; though we would typically
-expect these two to go together. As its name indicates, METI assesses
-perceived trustworthiness, not trust (such as accepting Dr.~Spence's
-conclusion about BPA).
+dimensions were strongly correlated in our sample (Pearson r values
+ranging from .75 to .92). To avoid issues of multicollinearity in our
+analyses, we averaged participant scores into a single composite
+measure. In both Elliott et al. (2017) and our replication analysis, all
+items were averaged together to form a 1-7 composite measure of
+trustworthiness. To aid interpretation, in the analysis the direction of
+the scale has been set so that increased values correspond to increased
+perceived trustworthiness.
+
+As its name indicates, METI assesses perceived trustworthiness rather
+than trust (such as accepting Dr.~Spence's conclusion about BPA). It is
+conceivable that participants might trust Dr.~Spence's conclusion even
+if they judge Spence to be untrustworthy, or vice versa. But we would
+typically trust and trustworthiness to go together.
 
@@
 
-After filling in the METI, participants provided demographic information
-and other sections of the survey that are not examined here. Due to
+After filling in the METI, participants provided demographic
+information, including self-identifying their political ideology, and
+other sections of the survey that are not examined here. Due to
 researcher error, a question about the participants' values (whether
@@
 \item[H5. Variation in effects]
-The magnitude of the effects 2-4 vary depending on whether the
-participant prioritizes public health or economic growth.
+The magnitude of the effects for hypotheses 2-4 vary depending on
+whether the participant prioritizes public health or economic growth.
 \end{description}
@@
 
-Hypothesis 1 was analyzed using Spearman rank order correlation and
-visually. Hypotheses 2-5 were analyzed using linear regression models as
-a common framework, with a direct acyclic graph (DAG) constructed
-\emph{a priori} to identify appropriate adjustments (covariates) for H4
-and 5. Because Elliott et al. (2017) made their data publicly available,
-and our analytical approaches differ somewhat from theirs, we conducted
-parallel analyses of their data, and report these parallel results for
-H2 and 3. Exploratory data analysis was used throughout to support data
-validation and aid interpretation.
+Hypothesis 1 was analyzed using Spearman rank order correlation to test
+H1a and a visual inspection of descriptive statistics to test H1b.
+Hypotheses 2-5 were analyzed using linear regression models as a common
+framework, with a direct acyclic graph (DAG) constructed \emph{a priori}
+to identify appropriate adjustments (covariates) for H4 and 5. Because
+Elliott et al. (2017) made their data publicly available, and our
+analytical approaches differ somewhat from theirs, we conducted parallel
+analyses of their data, and report these parallel results for H2 and 3.
+Exploratory data analysis was used throughout to support data validation
+and aid interpretation.
 
@@
 to the Public Health + Does Not Cause Harm condition
-(Table~\ref{tbl-condition}). Of the full 988 participants, 844
-participants (85\%) responded to the followup question about their own
-values (participant prioritizes economic growth or public health).
-Consequently, subsamples for hypotheses 4 and 5 were substantially
-smaller than the full analysis sample.
+(Table~\ref{tbl-condition}). Due to researcher error a question about
+participants' values was not included in the original survey. Of the
+full 988 participants, 844 participants (85\%) responded to the followup
+question about their own values (participant prioritizes economic growth
+or public health). Consequently, subsamples for hypotheses 4 and 5 were
+substantially smaller than the full analysis sample.
 
@@
 The study was approved by the UC Merced IRB on August 17, 2021, and data
-collection ran October 18-20, 2021. As mentioned above, due to
-researcher error a question about participants' values was not included
-in the original survey. A followup survey asking this question of the
-same group of participants was conducted December 8, 2021 through March
-5, 2022.
+collection ran October 18-20, 2021. The followup survey asking the
+initial sample of participants about their own values regarding economic
+growth and public health was conducted December 8, 2021 through March 5,
+2022.
 
@@
 \emph{t}(335) = 2.48, \emph{p} = 0.014. It seems likely that claims of a
-shared values effect could have been the result of a less appropriate
+shared values effect could have been an artifact of a less appropriate
 analytic approach given the nature of the data.
@@
 course of the COVID-19 pandemic, members of the general public have
-become used to ``scientists'' (including both bench researchers and
-public health officials) making claims about the importance of
+become used to ``scientists'' (broadly including both bench researchers
+and public health officials) making claims about the importance of
 protecting public health. A values disclosure that might have been
@@
 of public health.
+
+All together, regarding the philosophical debates over values in
+science, the evidence from this replication study suggests that the
+effects of transparency on perceived trustworthiness might be
+context-dependent. Transparency about the influence of might be
+detrimental in some cases, beneficial in others, and neutral in still
+others. Notably, both of the papers cited as objecting to transparency
+(John 2017; Kovaka 2021) focus on climate change, while many of
+Elliott's examples come from toxicology or environmental public health.
+The critics might be right that transparency has undermined trust in the
+climate context, while Elliott might be right that transparency can be
+neutral or even beneficial in some public health contexts. Because
+philosophers of science tend to work with a small number of rich case
+studies, methods such as qualitative comparative analysis (QCA; Ragin
+2013) would be worth exploring.
+
+An obvious difference between climate and BPA is the public prominence
+and political polarization of the two issues. For decades, the fossil
+fuels industry has used a range of strategies to successfully attack the
+perceived trustworthiness of climate scientists, especially in the minds
+of conservative publics (Oreskes and Conway 2011; Supran and Oreskes
+2017). While the chemical industry has also had a substantial influence
+on the environmental safety regulatory system (Michaels 2008; Vogel
+2013), controversies over BPA have received much less public attention.
+A LexiNexis search indicates that uses of ``bisphenol'' in US newspapers
+rose and then almost immediately fell around 2008, shortly after a US
+consensus conference concluded that ``BPA at concentrations found in the
+human body is associated with organizational changes'' in a number of
+major physiological systems (vom Saal et al. 2007).
+
+In more prominent and/or polarized cases than BPA, a number of other
+psychological mechanisms might be activated. Motivated reasoning or
+identity-protective cognition (Kahan et al. 2012) might cause members of
+the general public to rate a scientist as more or less trustworthy
+insofar as the scientist agrees with the respondent's ``side'' of the
+debate, regardless of whether and which values are disclosed. That is,
+the scientist conclusion might have a larger effect than scientist
+values. Alternatively, respondents might use values disclosures to
+identify which scientists are on ``their side,'' and so scientist values
+would have a larger effect than scientist conclusion. Careful attention
+to experimental design will be needed to separate these potential
+multiway interactions between conclusions, disclosures, and public
+prominence.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-KahanPolarizingImpactScience2012}{}}%
+Kahan, Dan M, Ellen Peters, Maggie Wittlin, Paul Slovic, Lisa Larrimore
+Ouellette, Donald Braman, and Gregory Mandel. 2012. {``The Polarizing
+Impact of Science Literacy and Numeracy on Perceived Climate Change
+Risks.''} \emph{Nature Climate Change} 2 (10): 732--35.
+\url{https://doi.org/10.1038/nclimate1547}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-KennedyEvaluation2016Election2018}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-MichaelsDoubtTheirProduct2008}{}}%
+Michaels, David. 2008. \emph{Doubt is their product how industry's
+assault on science threatens your health}. Oxford: Oxford Univ. Press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-MildenbergerDistributionClimateChange2016}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-OreskesMerchantsDoubtHow2011}{}}%
+Oreskes, Naomi, and Erik M. Conway. 2011. \emph{Merchants of Doubt: How
+a Handful of Scientists Obscured the Truth on Issues from Tobacco Smoke
+to Global Warming}. New York, NY, USA: Bloomsbury.
+
 \leavevmode\vadjust pre{\hypertarget{ref-PeerDataQualityPlatforms2021}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-RaginComparativeMethodMoving2013}{}}%
+Ragin, Charles C. 2013. \emph{The comparative method: Moving beyond
+qualitative and quantitative strategies}. Oakland, California:
+University of California Press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-RepresentativeSamplesFAQ2022}{}}%
@@
 \url{https://web.archive.org/web/20220313050153/https://researcher-help.prolific.co/hc/en-gb/articles/360019238413-Representative-samples-FAQ}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vomSaalChapelHillBisphenol2007}{}}%
+Saal, Frederick S. vom, Benson T. Akingbemi, Scott M. Belcher, Linda S.
+Birnbaum, D. Andrew Crain, Marcus Eriksen, Francesca Farabollini, et al.
+2007. {``Chapel Hill Bisphenol A Expert Panel Consensus Statement:
+Integration of Mechanisms, Effects in Animals and Potential to Impact
+Human Health at Current Levels of Exposure.''} \emph{Reproductive
+Toxicology} 24 (2): 131--38.
+\url{https://doi.org/10.1016/j.reprotox.2007.07.005}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-SupranAssessingExxonMobilClimate2017}{}}%
+Supran, Geoffrey, and Naomi Oreskes. 2017. {``Assessing ExxonMobil's
+Climate Change Communications (1977--2014).''} \emph{Environmental
+Research Letters} 12 (8): 084019.
+\url{https://doi.org/10.1088/1748-9326/aa815f}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-VogelItSafeBPA2013}{}}%
+Vogel, Sarah A. 2013. \emph{Is It Safe?: BPA and the Struggle to Define
+the Safety of Chemicals}. Oakland, California: University of California
+Press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-WickhamWelcomeTidyverse2019}{}}%

### 2022-10-18 paper/paper.Qmd
2 grammar issues

@@
 Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa.
-But we would typically trust and trustworthiness to go together.
+But we would typically expect trust and trustworthiness to go together.
 
@@
 All together, regarding the philosophical debates over values in science, the evidence from this replication study suggests that the effects of transparency on perceived trustworthiness might be context-dependent.
-Transparency about the influence of might be detrimental in some cases, beneficial in others, and neutral in still others.
+Transparency about the influence of values might be detrimental in some cases, beneficial in others, and neutral in still others.
 Notably, both of the papers cited as objecting to transparency [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021] focus on climate change, while many of Elliott's examples come from toxicology or environmental public health.
@@
 Motivated reasoning or identity-protective cognition [@KahanPolarizingImpactScience2012] might cause members of the general public to rate a scientist as more or less trustworthy insofar as the scientist agrees with the respondent's "side" of the debate, regardless of whether and which values are disclosed.
-That is, the scientist conclusion might have a larger effect than scientist values.
-Alternatively, respondents might use values disclosures to identify which scientists are on "their side," and so scientist values would have a larger effect than scientist conclusion.
+That is, the scientist's conclusion might have a larger effect than scientist's disclosed values.
+Alternatively, respondents might use values disclosures to identify whether a scientist is on "their side," and so the scientist's disclosed values might have a larger effect than the scientist's conclusion.
 Careful attention to experimental design will be needed to separate these potential multiway interactions between conclusions, disclosures, and public prominence.

### 2022-10-18 paper/paper.tex
2 grammar issues

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, sharp corners, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 if they judge Spence to be untrustworthy, or vice versa. But we would
-typically trust and trustworthiness to go together.
+typically expect trust and trustworthiness to go together.
 
@@
 effects of transparency on perceived trustworthiness might be
-context-dependent. Transparency about the influence of might be
+context-dependent. Transparency about the influence of values might be
 detrimental in some cases, beneficial in others, and neutral in still
@@
 debate, regardless of whether and which values are disclosed. That is,
-the scientist conclusion might have a larger effect than scientist
-values. Alternatively, respondents might use values disclosures to
-identify which scientists are on ``their side,'' and so scientist values
-would have a larger effect than scientist conclusion. Careful attention
-to experimental design will be needed to separate these potential
-multiway interactions between conclusions, disclosures, and public
-prominence.
+the scientist's conclusion might have a larger effect than scientist's
+disclosed values. Alternatively, respondents might use values
+disclosures to identify whether a scientist is on ``their side,'' and so
+the scientist's disclosed values might have a larger effect than the
+scientist's conclusion. Careful attention to experimental design will be
+needed to separate these potential multiway interactions between
+conclusions, disclosures, and public prominence.
 
